### 2022-10-21 src/demo_data_wrangling.Rmd
population data complation

---
title: "Metropolitan Statistical Area Evaluation"
author: "Grace Bianchi"
date: "5/12/2022"
output:   html_document
---

This report provides an explanatory overview of demographic and real estate data for metropolitan statistical areas. The code wrangles and analyzes data collected from government agencies to create metrics to explore the relationships between criteria and compare across MSAs. Specifically, this data analysis includes population growth, household income, employment rates, house prices, and property taxes. Preliminary analysis of population change and size narrowed down potential areas which experienced population growth rate greater than 0.5% and have populations greater than 300,000 people. In addition, 28 MSAs were removed from analysis, as requested by the client, due to hyper-competitive real estate markets. In brief, there are 44 metropolitan statistical areas with potential for ZNE Capital aqcuire real estate.

## Preliminary analysis {.tabset}

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor) #clean_names
library(DT) # create data table
library(scales) 
library(lubridate)
library(gridExtra) 
library(ggrepel)
library(kableExtra)

library(tigris) # Census shape files
library(sf)
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)
```

### Population Growth
All states were considered with the exception of CA, HI, and AK. Initial analysis of Metropolitan Statistical Areas (MSAs) relied on Census data of annual population estimates, in which the population had to be greater than 300,000 individuals for the purpose of attracting investors. Population change from 2020 to 2021 was calculated by $$\frac{2021 estimate- 2020 estimate}{2021 estimate} * 100$$
MSAs with a population growth > 0.5%, were selected to gather further data on. The 0.5% growth was determined.

Dataset: Annual Resident Population Estimates and Estimated Components of Resident Population Change for Metropolitan and Micropolitan Statistical Areas and Their Geographic Components: April 1, 2020 to July 1, 2021 (CBSA-EST2021-ALLDATA)

Source: [U.S. Census Bureau, Population Division](https://www.census.gov/data/tables/time-series/demo/popest/2020s-total-metro-and-micro-statistical-areas.html)

```{r population growth data wrangling}
# read in the data MSA population data from Census
msa_pop_2021_data <- read_csv(here("data","raw", "msa_2021_alldata.csv")) %>% 
  clean_names() %>% # data wrangling to make column headers in lower_snake_case
  filter(lsad == "Metropolitan Statistical Area") %>% # select MSA cities
  select("cbsa", "name", "popestimate2020", "popestimate2021", "residual2020", "residual2021") %>% # select data 
  filter(popestimate2021 >= 300000) %>% 
  separate(col = "name",
           into = c("MSA", "state"), # split city and state data
           sep = ", ") %>% 
  mutate(pop_percent_change = (popestimate2021-popestimate2020)/popestimate2021 * 100)  # add population % change column

# create __ 
potential_MSAs <- msa_pop_2021_data %>% 
  filter(pop_percent_change > 0.5 ) %>% 
  mutate(pop_percent_change20_21 = round(pop_percent_change, digits = 3)) %>% 
  filter (!state %in% c("CA", "HI", "AK")) %>% 
  #separate(col = "state", into = c("state1", "state2", "state3"),    sep = "-") %>% 
  filter(!cbsa %in% c("10580", "10900", "12420", "12940", "14260", "15980", "18580", "19100", "19660", "19740", "23540", "25420", "26420", "27260", "29820", "34940", "36260", "36740", "38060", "39100", "39300", "39340", "41620", "45220", "45300", "46060", "46140", "39900")) %>%  # remove MSAs as instructed by Owen 
  select(cbsa, MSA, state, pop_percent_change20_21)

```

```{r population growth msa map, eval=FALSE}
# adding column for levels of population growth
pop_growth_pos <- potential_MSAs %>% 
  mutate(pop_pos = case_when(
    pop_percent_change20_21 > 2 ~ "high",
    pop_percent_change20_21 > 1 ~ "medium",
    pop_percent_change20_21 > 0 ~ "low"
  ))

# MSA shapefiles from Census
cbsa <- core_based_statistical_areas(cb = TRUE) %>% 
  mutate(CBSAFP = as.numeric(CBSAFP)) %>% 
  inner_join(pop_growth_pos, by = c("CBSAFP" = "cbsa"))

# state and U.S. map coordinates
states <- map_data("state")
usa<- map_data("usa")

# map of U.S. with MSAs highlighted
ggplot() +
  geom_polygon(data = states,
               aes(x = long, y = lat,
                   group = group)) + 
  geom_sf (data = cbsa, aes(fill = pop_pos)) +
  labs(title = "MSAs with the highest population growth in 2020")
```

Figure 1. Population growth in Metropolitan Statistical Areas (n=44) in 2020. Highlighted areas have a population > 300,000 individuals and a population growth rate greater than 0.5%. The colors of the MSAs represent incremental growth rates, in which low(green) is 0.5-1% increase, medium(blue) is 1-2% increase, and high(pink) is 2-3.5% increase. Data: U.S. Census Bureau, Population Division.

```{r fig.height = 7, fig.align='center', eval = FALSE}
ggplot(data = positive_MSA_pop_change, aes(x = pop_percent_change20_21, y = reorder(MSA, pop_percent_change20_21))) + 
  geom_point(size = 3, color = "darkgreen") + 
  theme_minimal(base_size = 12.5)

```

### Household Income
Household Median Income data was also obtained from Census data at the MSA level and amended to existing MSAs with a positive population increase. Median is a better indicator of the population than the mean income values due to the positive skew from the highest incomes.

Dataset: Median Household Income (USD) INCOME IN THE PAST 12 MONTHS (IN 2020 INFLATION-ADJUSTED DOLLARS)

Source: [U.S. Census Bureau](https://data.census.gov/cedsci/table?q=median%20household%20income&t=Income%20and%20Earnings&g=0100000US%243100000&tid=ACSST5Y2020.S1901)

```{r household income data wrangling}
MSA_income_data <- read_csv(here("data","raw", "MSA_income", "2021_data", "2021_income_data.csv")) %>% 
  row_to_names(1) %>% 
  clean_names() %>% 
  mutate(median_income = as.numeric(estimate_households_median_income_dollars)) %>% 
  mutate(income_moe_USD = as.numeric(margin_of_error_households_median_income_dollars)) %>% 
  separate(col = "geography",
           into = c("id", 'cbsa'),
           sep = 9) %>% 
  mutate(cbsa = as.numeric(cbsa)) # change column class from chr to dbl

# join population growth data set to household median income
MSA_house_income <- MSA_income_data %>% # rename row names
  inner_join(potential_MSAs) %>% 
  select(cbsa, MSA, state,median_income, income_moe_USD) 
#select(cbsa, MSA, state, pop_percent_change20_21, median_income, income_moe_USD)

```

```{r, household income visual, fig.height = 7, fig.align='center'}
ggplot(MSA_house_income, aes(x = median_income, y = reorder(MSA, median_income))) +
  geom_errorbarh(aes(xmin = median_income - income_moe_USD, xmax = median_income + income_moe_USD)) + 
  geom_point(size = 3, color = "darkgreen") + 
  theme_minimal(base_size = 12.5) + 
  labs(title = "Median household income in 2021", 
       subtitle = "", 
       x = "estimate (in USD)", 
       y = "") + 
  scale_x_continuous(labels = label_dollar())

```

Figure 2. Median Household income of Metropolitan Statistical Areas in 2021. The green points indicate the median income value for each MSA with the margin of error presented by the bars on either side of each point. Data: U.S. Census Bureau.

### Employment Rates
MSA monthly unemployment rates (seasonally-adjusted) were provided by The Bureau of Labor statistics in the civil labor force data at the MSA level. The average annual unemployment rates were calculated using the monthly rates from 2012 to 2020. Employment rates were calculated by dividing the number of employment / civilian labor force. The employment data for the month of January in years 2018-2021 was selected to calculate the change in employment rates over time. Year-over-Year employment growth was calculated by dividing the employment rate of the year by the previous year and subtracting one; for example, 2018 employment YoY growth = (2019 employment rate / 2018 employment rate) - 1.

Annual Average Employment Rates & Year-over-Year change civil labor force data from [The Bureau of Labor Statistics](https://www.bls.gov/lau/metrossa.htm)

Monthly unemployment rates (seasonally-adjusted) from [The Bureau of Labor Statistics](https://www.bls.gov/web/metro/metro_area_3mma.pdf)

```{r, employment data wrangling}
unemployment_data <- read_csv(here("data", "raw", "civil_labor_force_BLS.csv")) %>% #rename col headers
  clean_names() %>% 
  rename(cbsa = x3) %>% 
  rename(area = x4) %>% 
  rename(year = x5) %>% 
  rename(month = x6) %>% 
  rename(civilian_labor_force = x7) %>% 
  rename(employment = x8) %>% 
  rename(unemployment = x9) %>% 
  rename(unemployment_rate = x10) %>% 
  mutate(employment_rate = (employment/civilian_labor_force)*100) # calculate

# data wrangling, only selecting data for potential MSAs
MSA_employment <- unemployment_data %>% 
  filter(year >= 2015) %>% 
  full_join(MSA_house_income) %>% 
  drop_na() %>% 
  select(cbsa, MSA, year, month, employment_rate, median_income, income_moe_USD)

# calculate YoY growth 
MSA_employment_yoy <- MSA_employment %>% 
  filter(year >= 2016) %>% 
  filter(month == "01") %>% 
  arrange(MSA, month) %>% 
  group_by(MSA) %>% 
  mutate(YoY = (employment_rate-lag(employment_rate))/lag(employment_rate)*100) %>% 
  mutate(employment_YoY = round(YoY, digits = 3)) %>% 
  mutate(emp_pos = employment_YoY >= 0) %>% # function to create column for yoy value, true = positive YoY
  select(!c(YoY, month)) %>% 
  drop_na()

```

```{r eval = FALSE}
# average employment from 2000 to 2022
employment_summary_since2000 <- MSA_employment %>% 
  group_by(MSA) %>% 
  filter(year >= 2000) %>% 
  summarize(
    avg_employment = mean(employment_rate),
    sd_employment = sd(employment_rate)
  )

```

```{r employment visuals, fig.align = "center", fig.show="hold", fig.height= 7, fig.width=10}
# visual of MSA employment rates, shows major drop during covid
ggplot(MSA_employment_yoy, aes(x = year, y = employment_rate, color = MSA)) +
  geom_line() +
  xlim(2016, 2022) +
  theme(legend.position = "none") +

  labs(x = "Year",
       y = "Employment Rate",
      title = "Employment Rate Trend since 2016")

# year over year change in employment rate, split up by year
ggplot(MSA_employment_yoy, aes(x = employment_YoY, y = reorder(MSA, employment_YoY), fill = emp_pos)) +
  geom_col(position = "identity", color = "black", size = 0.25) +
  scale_fill_manual(values = c("red", "darkolivegreen3"), guide = FALSE) +
  facet_grid(~year) +
  labs(x = "Year-over_year change in employment rate",
       y = "")

```

```{r payroll data}
# payroll data has not been added to final table
payroll_data <- read_csv(here("data","raw", "employee_payroll_MSA21_22.csv")) %>% 
  row_to_names(4) %>% 
  clean_names() %>% 
  rename(MSA = na) %>% 
  rename(payroll_change_may21 = may_2021) %>% 
  rename(payroll_change_mar22 = mar_2022) %>% 
  rename(payroll_change_apr22 = apr_2022) %>% 
  rename(payroll_change_may22 = may_2022_p)

MSA_payroll <- payroll_data %>% 
  full_join(potential_MSAs) %>% 
  drop_na() %>% 
  pivot_longer(cols = 2:5,
               names_prefix = "payroll_change_",
               names_to = "year",
               values_to = "payroll_change") %>% 
  mutate(payroll_change = as.numeric(payroll_change)) %>% 
  group_by(MSA, year)  %>% 
  mutate(payroll_pos = payroll_change >= 0)

```
```{r payroll visual, fig.height = 7}
ggplot(MSA_payroll, aes(x = payroll_change, y = reorder(MSA, payroll_change), fill = payroll_pos)) +
  geom_col(position = "identity", color = "black", size = 0.25) +
  scale_fill_manual(values = c("red", "darkolivegreen3"), guide = FALSE) +
  labs(x = "percent change",
       y = "",
       title = "Quarterly change in employee payroll from 2021 to 2022") +
  facet_grid(~year)

```

<https://www.bls.gov/web/metro.supp.toc.htm> <https://www.bls.gov/web/metro/metro_oty_change.htm> - significant change

### House Prices

Median housing prices by MSA were obtained from the Association of Realtors for years 2019, 2020, and 2021. The percent difference in 2021 Q1 and 2022 Q1 was used to show the change in price in 2021. **Next quarterly release will be 11/10**

Source: [The National Association of Realtors](https://www.nar.realtor/research-and-statistics/housing-statistics/metropolitan-median-area-prices-and-affordability)

```{r house price data}
metro_house_cost_data <- read_csv(here("data","raw", "metro_home_costs", "metro_home_prices.csv")) %>% 
  row_to_names(3) %>% 
  clean_names() %>% 
  separate(col = "metropolitan_area",
           into = c("cbsa", 'metro_area'),
           sep = 5) %>% 
  mutate(cbsa = as.numeric(cbsa)) %>% 
  mutate(x2020 = as.numeric(x2020))

metro_house_prices <- MSA_house_income %>%
  left_join(metro_house_cost_data)  %>% 
  rename(house_price_2020 = x2020) %>% 
  rename(house_price_2021 = x2021_r) %>% 
  rename(houseprice_change21_22 = q1_q1) %>% 
  select(!c(metro_area, x2019, na)) %>% 
  # select(cbsa, houseprice_change21_22) %>% 
  separate(col = "houseprice_change21_22",
           into = "houseprice_change21_22",
           sep = "%") %>% 
  mutate(houseprice_change21_22 = as.numeric(houseprice_change21_22))
```

```{r house price exploratory, eval = FALSE}
ggplot(data = metro_house_cost_data, aes(x = log(x2020))) +
   geom_histogram(bins = 100)

ggplot(data = metro_house_cost_data, aes(sample = log(x2020))) +
  geom_qq()

```

The following MSAs were not included in the housing dataset:

- Augusta-Richmond County
- Clarksville, TN-KY
- Greeley, CO
- Killeen, TX
- Lubbock, TX
- McAllen-Edinburg-Mission, TX
- Salisbury, MD
- Savannah, GA

```{r, house price visual, fig.height = 7, fig.align='center'}
ggplot(metro_house_prices, aes(x = houseprice_change21_22, y = reorder(MSA, houseprice_change21_22))) +
  geom_col(fill = "dodgerblue4") +
  geom_text(aes(label = houseprice_change21_22),
            size = 4,
            hujust = 1, nudge_x = -1,
            color = "white") +
  theme_minimal()  +
  labs(x = "% change",
       y = "",
       title = "Change in House Prices in 2021")

```

### Property Taxes
It was difficult to find property tax information by city that was consistent across cities. Median real estate taxes in dollars are provided by the U.S. Census and data is collected at the MSA level. A metric of property tax and median home prices can be used to normalized.

Dataset: B25103  MORTGAGE STATUS BY MEDIAN REAL ESTATE TAXES PAID (USD) in 2019

Source:[U.S. Census Bureau](https://data.census.gov/cedsci/table?q=property%20tax&g=0100000US%2431000M1&d=ACS%201-Year%20Estimates%20Detailed%20Tables&tid=ACSDT1Y2019.B25103&tp=true)

```{r property taxes data wrangling}
property_tax_data <- read_csv(here("data","raw", "wide_property_tax_data.csv"))

# create new data frame with tax estimate values
median_tax_df <- property_tax_data %>% 
  select(ends_with("estimate")) %>% 
  slice(2) %>% 
  stack() %>% # flip data table to be tidy
  rename(MSA = ind) %>% 
  mutate(median_tax_USD = as.numeric(values)) %>% 
  separate(col = MSA,
           into = c("MSA", "state"),
           sep = ", ") %>% 
  separate(col = state,
           into = c("state"),
           sep = " ") %>% 
  select(!values)

# make data frame with margin of error values
moe_tax_df <- property_tax_data %>% 
  select(ends_with("error")) %>% 
  slice(2) %>% 
  stack() %>% # flip data table to be tidy
  rename(MSA = ind) %>% 
  separate(col = values,
           into = c("value", "tax_moe_USD"),
           sep = 1) %>% 
  mutate(tax_moe_USD = as.numeric(tax_moe_USD)) %>% 
  separate(col = MSA,
           into = c("MSA", "state"),
           sep = ", ") %>% 
  separate(col = state,
           into = c("state", "area"),
           sep = " ") %>% 
  select(!c(value, area))

MSA_property_tax <- metro_house_prices %>% 
  full_join(median_tax_df) %>% 
  left_join(moe_tax_df) %>% 
  drop_na(cbsa) # drop all MSA rows if NA in cbsa column

```

```{r property tax visual, fig.height = 7, fig.align='center'}
ggplot(MSA_property_tax, aes(x = median_tax_USD, y = reorder(MSA, median_tax_USD))) +
  geom_errorbarh(aes(xmin = median_tax_USD - tax_moe_USD, xmax = median_tax_USD + tax_moe_USD)) + 
  geom_point(size = 3, color = "darkgreen") + 
  theme_minimal(base_size = 12.5) + 
  labs(title = "Median property tax in 2019", 
       subtitle = "", 
       x = "Median Property Tax estimate (in USD)", 
       y = "") + 
  scale_x_continuous(labels = label_dollar())
```

```{r join population, unemployment, income, and house price data}
# join employment data with income, house prices, tax, and population data
MSApop_income_housecost_tax_emp <- MSA_employment_yoy %>%
  filter(year == "2022") %>% 
  mutate(employment_rate = round(employment_rate, digits = 3)) %>% 
  select(cbsa, year, employment_rate, employment_YoY) %>% 
  right_join(MSA_property_tax) %>% 
  select(!year)

```

### MSA data table

```{r data table}
# table seperated by MSAs
datatable(MSApop_income_housecost_tax_emp, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  )
)
```

### City Data Table
MSAs were split up by cities to collect utility, rent, and landlord data.

```{r MSAs seperate into cities}
#  split up MSAs into cities to collect utility, rent, and landlord data.
msa_seperated <- MSApop_income_housecost_tax_emp %>% 
  separate_rows(MSA, sep = "-") %>% 
  rename(city = MSA)# %>% # seperate MSA cities into their own columns
 # full_join(payroll_data, by = c("MSA" = "area")) %>% 
 # drop_na(pop_percent_change20_21)

# table seperated by city
datatable(msa_seperated, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  )
)

```

### 2022-10-21 src/real_estate.Rmd
reading files from googledrive urls

---
title: "Real Estate"
author: "Grace Bianchi"
date: '2022-10-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-10-21 src/real_estate.Rmd
reading google sheet urls

@@
 knitr::opts_chunk$set(echo = TRUE)
+
+library(googlesheets4)
+
+MSA_demo_data <- read_sheet("https://docs.google.com/spreadsheets/d/12do9sRCAcrIe0K-VcJWXNmmTm_jg5rKjZ0NeJPOm9Jw/edit#gid=0")
+city_realEstate_ata <- read_sheet("https://docs.google.com/spreadsheets/d/12do9sRCAcrIe0K-VcJWXNmmTm_jg5rKjZ0NeJPOm9Jw/edit#gid=359643482")
 ```
-
-## R Markdown
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-10-21 src/landlord.Rmd
made landlord file

---
title: "landlord"
author: "Virginia Pan"
date: "10/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-10-21 src/real_estate.Rmd

@@
 MSA_demo_data <- read_sheet("https://docs.google.com/spreadsheets/d/12do9sRCAcrIe0K-VcJWXNmmTm_jg5rKjZ0NeJPOm9Jw/edit#gid=0")
-city_realEstate_ata <- read_sheet("https://docs.google.com/spreadsheets/d/12do9sRCAcrIe0K-VcJWXNmmTm_jg5rKjZ0NeJPOm9Jw/edit#gid=359643482")
+city_realEstate_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990")
 ```

### 2022-10-22 src/landlord.Rmd
imported landlord data and cleaning it.

@@
 knitr::opts_chunk$set(echo = TRUE)
+
+library(googlesheets4)
+library(janitor)
+library(tidyverse)
+library(kableExtra)
+
+MSA_landlord_data <- read_sheet("https://docs.google.com/spreadsheets/d/1zCePYRyJ5lb1Z5jvCujg17EtwqpxJN1E1_HwJL66uLM/edit#gid=122953884") 
 ```
 
-## R Markdown
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
+```{r}
+MSA_landlord_data_binary <- MSA_landlord_data %>% 
+  select(cbsa, MSA, rent_control, eviction_notice_days, security_depopsit_limit_YN, notice_for_entry_days_YN)
 ```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-10-22 src/landlord.Rmd
calculating landlord score

@@
 ```{r}
-MSA_landlord_data_binary <- MSA_landlord_data %>% 
-  select(cbsa, MSA, rent_control, eviction_notice_days, security_depopsit_limit_YN, notice_for_entry_days_YN)
+# #only take Y/N columns
+# MSA_landlord_data_binary <- MSA_landlord_data %>% 
+#   select(cbsa, MSA, rent_control, eviction_notice_days, security_depopsit_limit_YN, notice_for_entry_days_YN)
 ```
+
+```{r}
+#mutate Y/N columns into 0/1
+
+#rent_control
+MSA_landlord_data <- MSA_landlord_data %>% 
+  mutate(
+    rent_control_binary = case_when(
+      rent_control == "N" ~ 1, 
+      rent_control == "Y" ~ 0))
+
+#security_deposit_limit
+MSA_landlord_data <- MSA_landlord_data %>% 
+  mutate(
+    security_depopsit_limit_binary = case_when(
+      security_depopsit_limit_YN == "N" ~ 1, 
+      security_depopsit_limit_YN == "Y" ~ 0))
+
+#notice_of_entry_days
+MSA_landlord_data <- MSA_landlord_data %>% 
+  mutate(
+    notice_for_entry_days_binary = case_when(
+      notice_for_entry_days_YN == "N" ~ 1, 
+      notice_for_entry_days_YN == "Y" ~ 0))
+```
+
+```{r}
+#normalize eviction notice (days)
+MSA_landlord_data <- MSA_landlord_data %>% 
+  mutate(
+    eviction_notice_days_normalized = (eviction_notice_days-min(eviction_notice_days))/(max(eviction_notice_days)-min(eviction_notice_days)) 
+  )
+
+#"invert" normalized score for eviction notice days
+#1 - normalized score (since more eviction notice is worse for landlords)
+MSA_landlord_data <- MSA_landlord_data %>% 
+  mutate(
+    eviction_notice_days_normalized_inverted = 1 - eviction_notice_days_normalized
+  )
+```
+
+```{r}
+#calculate landlord category score
+
+#input landlord weights for each indicator in the following order: rent control, security deposit limit, notice for entry, eviction notice days
+#Get weights from Owen
+#Weights must sum to 1
+landlord_weights <- c(0.5, 0.1, 0.1, 0.3)
+size <- dim(MSA_landlord_data)
+for (x in 1:size[1]) {
+  print(x)
+}
+x```

### 2022-10-22 src/landlord.Rmd
landlord scores calculated

@@
 #Weights must sum to 1
-landlord_weights <- c(0.5, 0.1, 0.1, 0.3)
+landlord_weights <- c(0.25, 0.25, 0.25, 0.25)
+
+#make landlord_score column
+MSA_landlord_data <- MSA_landlord_data %>% 
+  mutate(
+    landlord_score = MSA #this is just a place holder
+  )
 size <- dim(MSA_landlord_data)
-for (x in 1:size[1]) {
-  print(x)
+for (i in 1:size[1]) {
+  MSA_x_data <- c(MSA_landlord_data$rent_control_binary[i], 
+                  MSA_landlord_data$security_depopsit_limit_binary[i],
+                  MSA_landlord_data$notice_for_entry_days_binary[i],
+                  MSA_landlord_data$eviction_notice_days_normalized_inverted[i])
+  MSA_landlord_data$landlord_score[i] = weighted.mean(MSA_x_data, landlord_weights)
 }
-x```
+```

### 2022-10-25 src/real_estate.Rmd
psuedocode

--- 2022-10-21 src/real_estate.Rmd
@@
 MSA_demo_data <- read_sheet("https://docs.google.com/spreadsheets/d/12do9sRCAcrIe0K-VcJWXNmmTm_jg5rKjZ0NeJPOm9Jw/edit#gid=0")
-city_realEstate_ata <- read_sheet("https://docs.google.com/spreadsheets/d/12do9sRCAcrIe0K-VcJWXNmmTm_jg5rKjZ0NeJPOm9Jw/edit#gid=359643482")
+
+city_realEstate_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990")
 ```
+
+```{r}
+# aggregate yardi data into MSAs by cbsa
+MSA_realEstate_data <- city_realEstate_data %>%
+
+```
+
+```{r}
+# calculate rent to income ratio
+
+# YoY rent change
+```
+
+```{r}
+# calculate income to home price
+```
+
+```{r}
+
+```
+
+```{r}
+# normalize everything into new data table
+
+normalized <-
+
+```
+
+```{r}
+# invert normalized rent to income ratio so lower values (more favorable) have higher scores
+
+= 1- normalized score
+```
+
+```{r}
+# weights from owen
+
+# place holder column
+MSA_real_estate_score <- 
+  mutate
+
+weights <- c()
+
+```

### 2022-10-25 src/real_estate.Rmd
complied real estate and demographic data

@@
 
+library(tidyverse)
+library(janitor)
 library(googlesheets4)
@@
 MSA_realEstate_data <- city_realEstate_data %>% 
+  clean_names() %>% 
+  group_by(cbsa) %>% 
+  summarize(avg_rent_2021 = mean(average_rent_2021, na.rm = TRUE),
+            avg_rent_2022 = mean(x2022_rent, na.rm = TRUE),
+            avg_occupancy_21 = mean(percent_occupancy_2021, na.rm = TRUE))
 
@@
 ```{r}
+# combine real estate and demographic data by cbsa
+MSA_demo_data <- MSA_demo_data %>% 
+  left_join(MSA_realEstate_data)
+```
+
+```{r}
 # calculate rent to income ratio
+compliedMSA_realEstate <- MSA_demo_data %>% 
+  mutate(income_to_rent = median_income_USD_2021/avg_rent_2021)
 

### 2022-10-26 src/landlord.Rmd
landlord catagory scores calculated
Update landlord.Rmd
test-from VA github desktop2
virginia new laptop github test
testing github token on new laptop
still testing github
wends AM

@@
 ```{r}
-# #only take Y/N columns
-# MSA_landlord_data_binary <- MSA_landlord_data %>% 
-#   select(cbsa, MSA, rent_control, eviction_notice_days, security_depopsit_limit_YN, notice_for_entry_days_YN)
-```
-
-```{r}
-#mutate Y/N columns into 0/1
-
-#rent_control
-MSA_landlord_data <- MSA_landlord_data %>% 
-  mutate(
-    rent_control_binary = case_when(
-      rent_control == "N" ~ 1, 
-      rent_control == "Y" ~ 0))
+# #rent_control
+# MSA_landlord_data <- MSA_landlord_data %>% 
+#   mutate(
+#     rent_control_binary = case_when(
+#       rent_control == "N" ~ 1, 
+#       rent_control == "Y" ~ 0))
 
@@
   mutate(
-    security_depopsit_limit_binary = case_when(
-      security_depopsit_limit_YN == "N" ~ 1, 
-      security_depopsit_limit_YN == "Y" ~ 0))
+    security_depopsit_limit_num = case_when(
+      security_deposit_limit == "N" ~ 1.0, 
+      security_deposit_limit == "1_month_rent" ~ 0.5,
+      security_deposit_limit == "2_months_rent" ~ 0.0))
 
@@
   mutate(
-    notice_for_entry_days_binary = case_when(
-      notice_for_entry_days_YN == "N" ~ 1, 
-      notice_for_entry_days_YN == "Y" ~ 0))
+    notice_for_entry_days_num = case_when(
+      notice_for_entry_days == "N" ~ 1.0, 
+      notice_for_entry_days == "0" ~ 1.0,
+      notice_for_entry_days == "0.5" ~ 0.75,
+      notice_for_entry_days == "1" ~ 0.5,
+      notice_for_entry_days == "2" ~ 0.0,))
 ```
@@
 ```
+
+```{r}
+#test Virginia new laptop on github
+#testing github token setupf
+```

### 2022-10-26 src/real_estate.Rmd

@@
 ```{r}
+# select columns in demographic data
+MSA_demo_data <- MSA_demo_data %>% 
+  clean_names() %>% 
+  select(cbsa, msa, state, employment_yo_y21_22, pop_percent_change20_21, median_income_usd_2021, house_price_2021, median_tax_usd)
+```
+
+```{r}
 # aggregate yardi data into MSAs by cbsa
@@
 # combine real estate and demographic data by cbsa
-MSA_demo_data <- MSA_demo_data %>% 
+MSA_realEstate_data <- MSA_demo_data %>% 
   left_join(MSA_realEstate_data)

### 2022-10-26 src/landlord.Rmd
pick up on line 61
double checked scoring. Corrected security deposit limit

@@
       security_deposit_limit == "N" ~ 1.0, 
-      security_deposit_limit == "1_month_rent" ~ 0.5,
-      security_deposit_limit == "2_months_rent" ~ 0.0))
+      security_deposit_limit == "1_month_rent" ~ 0.0,
+      security_deposit_limit == "2_months_rent" ~ 0.5))
 
@@
 ```{r}
+#PICK UP HERE!
 #calculate landlord category score

### 2022-10-27 src/landlord.Rmd
through line 84 dilter rent control

@@
 ```{r}
-#PICK UP HERE!
 #calculate landlord category score
 
-#input landlord weights for each indicator in the following order: rent control, security deposit limit, notice for entry, eviction notice days
-#Get weights from Owen
+#input landlord weights for each indicator in the following order: eviction notice days, security deposit limit, notice for entry
+
 #Weights must sum to 1
-landlord_weights <- c(0.25, 0.25, 0.25, 0.25)
+landlord_weights <- c(0.9, 0.05, 0.05)
 
@@
 for (i in 1:size[1]) {
-  MSA_x_data <- c(MSA_landlord_data$rent_control_binary[i], 
-                  MSA_landlord_data$security_depopsit_limit_binary[i],
-                  MSA_landlord_data$notice_for_entry_days_binary[i],
-                  MSA_landlord_data$eviction_notice_days_normalized_inverted[i])
+  MSA_x_data <- c(MSA_landlord_data$eviction_notice_days_normalized_inverted[i],
+                  MSA_landlord_data$security_depopsit_limit_num[i],
+                  MSA_landlord_data$notice_for_entry_days_num[i]
+                  )
   MSA_landlord_data$landlord_score[i] = weighted.mean(MSA_x_data, landlord_weights)
@@
 ```{r}
-#test Virginia new laptop on github
-#testing github token setupf
+#PICK UP HERE
+#If there is rent control, immediately disqualify location 
 ```

### 2022-10-28 src/real_estate.Rmd
normalized data

--- 2022-10-25 src/real_estate.Rmd
@@
 ```{r}
+# select columns in demographic data
+MSA_demo_data <- MSA_demo_data %>% 
+  clean_names() %>% 
+    # calculate income to home price
+  mutate(income_to_homeprice = median_income_usd_2021/house_price_2021) %>% 
+  select(cbsa, msa, state, employment_yo_y21_22, pop_percent_change20_21, median_income_usd_2021, income_to_homeprice)
+```
+
+```{r}
 # aggregate yardi data into MSAs by cbsa
 MSA_realEstate_data <- city_realEstate_data %>% 
+  clean_names() %>% 
+  group_by(cbsa) %>% 
+  # calculate average values for MSAs with more than one city
+  summarize(avg_rent_2021 = mean(average_rent_2021, na.rm = TRUE),
+            avg_rent_2022 = mean(x2022_rent, na.rm = TRUE),
+            avg_occupancy_21 = mean(percent_occupancy_2021, na.rm = TRUE)) %>% 
+  # calculate change in rent
+  mutate(rent_change21_22 = (avg_rent_2022- avg_rent_2021)/avg_rent_2021)
 
@@
 ```{r}
+# combine real estate and demographic data by cbsa
+MSA_realEstate_data <- MSA_demo_data %>% 
+  left_join(MSA_realEstate_data) %>% 
   # calculate rent to income ratio
+  mutate(rent_to_income = avg_rent_2021/median_income_usd_2021)
 
-# YoY rent change
+```
+
+```{r}
 ```
@@
 ```{r}
-
-```
-
-```{r}
-# normalize everything into new data table
-
-normalized <-
+# normalize columns into new data table
+n_MSA_realEstate_data <- MSA_realEstate_data %>% 
+  # normalize population data
+  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>% 
+  # normalize employment growth
+  mutate(n_employ_change = (employment_yo_y21_22 - min(employment_yo_y21_22)) / (max(employment_yo_y21_22) - min(employment_yo_y21_22))) %>% 
+  # normalize rent to income
+  mutate(n_rent_to_income = (rent_to_income- min(rent_to_income)) / (max(rent_to_income) - min(rent_to_income))) %>% 
+  # normalize income to home price
+  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice)))
 

### 2022-10-31 src/demo_data_wrangling.Rmd
calculating and adding portland employment data

@@
   drop_na()
+
+```
+
+```{r}
+# adding portland employment rates from https://data.bls.gov/pdq/SurveyOutputServlet
+portland_employment_data <- read_csv(here("data", "raw", "portland_me_employment.csv")) %>% 
+  row_to_names(row_number = 10) %>% 
+  clean_names() %>% 
+  select(year, jan) %>% 
+  rename(civilian_labor = jan)
+
+employment_data_jan_2021 <- 100-4.9
+employment_data_jan_2022 <- 100- 3.6
+
+yoy <- (employment_data_jan_2022 - employment_data_jan_2021)/ employment_data_jan_2022 *100
 

### 2022-10-31 src/real_estate.Rmd
normalized data

@@
 ```{r}
-```
-
-```{r}
-# calculate income to home price
-```
-
-```{r}
 # normalize columns into new data table
-n_MSA_realEstate_data <- MSA_realEstate_data %>% 
+MSA_realEstate_data <- MSA_realEstate_data %>% 
   # normalize population data
@@
 # invert normalized rent to income ratio so lower values (more favorable) have higher scores
+MSA_realEstate_data <- MSA_realEstate_data %>% 
+  mutate(income_to_rent_inverted = 1 - n_income_to_rent)
 
-= 1- normalized score
 ```
@@
 # weights from owen
+real_estate_weights <- c()
 
-# place holder column
-MSA_real_estate_score <- 
-  mutate
+#make real_estate_score column
+MSA_realEstate_data <- MSA_realEstate_data %>% 
+  mutate(real_estate_score = MSA) #this is just a place holder
+```
 
-weights <- c()
+```{r}
+size <- dim(MSA_realEstate_data)
 
+for (i in 1:size[1]) {
+  MSA_x_data <- c(MSA_realEstate_data$eviction_notice_days_normalized_inverted[i],
+                  MSA_realEstate_data$security_depopsit_limit_num[i],
+                  MSA_realEstate_data$notice_for_entry_days_num[i]
+                  )
+  MSA_realEstate_data$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
+  }
 ```

### 2022-11-01 src/demo_data_wrangling.Rmd
added missing employment data and normalized values
median rent for MSAs

@@
 
-yoy <- (employment_data_jan_2022 - employment_data_jan_2021)/ employment_data_jan_2022 *100
+yoy <- (employment_data_jan_2022 - employment_data_jan_2021)/ employment_data_jan_2021 *100
 
@@
 <https://www.bls.gov/web/metro.supp.toc.htm> <https://www.bls.gov/web/metro/metro_oty_change.htm> - significant change
+
+2021 Median Rent from https://www.huduser.gov/portal/datasets/50per.html#2021
+
+```{r}
+# median rents from hud
+MSA_median_rent2021_data <- read_csv(here("data", "raw", "2021MSA_medianrent.csv")) %>% 
+  separate(col = cbsasub21,
+           into = c("metro", "cbsa"),
+           sep = -5) %>% 
+  mutate(cbsa = as.double(cbsa)) %>% 
+  select(cbsa, rent50_0, rent50_1, rent50_2, rent50_3, rent50_4) %>% 
+  right_join(potential_MSAs) 
+```
 

### 2022-11-01 src/demo_data_wrangling.Rmd
added median rent data and weights for data

--- 2022-10-31 src/demo_data_wrangling.Rmd
@@
 <https://www.bls.gov/web/metro.supp.toc.htm> <https://www.bls.gov/web/metro/metro_oty_change.htm> - significant change
+
+2021 Median Rent from https://www.huduser.gov/portal/datasets/50per.html#2021
+
+```{r}
+# median rents from hud
+MSA_median_rent2021_data <- read_csv(here("data", "raw", "2021MSA_medianrent.csv")) %>% 
+  separate(col = cbsasub21,
+           into = c("metro", "cbsa"),
+           sep = 5) %>% 
+  separate(col = cbsa,
+           into = c("cbsa", "unknown"),
+           sep = -5) %>% 
+  mutate(cbsa = as.double(cbsa)) %>% 
+  select(cbsa, rent50_0, rent50_1, rent50_2, rent50_3, rent50_4) %>% 
+  right_join(potential_MSAs) 
+```
 

### 2022-11-01 src/real_estate.Rmd
added median rent data and weights for data

--- 2022-10-28 src/real_estate.Rmd
@@
 city_realEstate_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990")
+
+MSA_median_rent_data <- read_sheet("https://docs.google.com/spreadsheets/d/1oxNxWQRWpANbgb603ShBlw84QkyvsRaH4TzHKKVWhxw/edit#gid=0")
 ```
@@
 # select columns in demographic data
-MSA_demo_data <- MSA_demo_data %>% 
+MSA_demographics <- MSA_demo_data %>% 
   clean_names() %>% 
@@
 ```{r}
+# MSA_median rent data
+MSA_realEstate_data<- MSA_median_rent_data %>% 
+  select(cbsa, rent50_1) %>%
+  mutate(rent50_1 = as.numeric(rent50_1)) %>% 
+  left_join(MSA_realEstate_data)
+```
+
+```{r}
 # combine real estate and demographic data by cbsa
-MSA_realEstate_data <- MSA_demo_data %>% 
+MSA_realEstate_data <- MSA_demographics %>% 
   left_join(MSA_realEstate_data) %>% 
   # calculate rent to income ratio
-  mutate(rent_to_income = avg_rent_2021/median_income_usd_2021)
+  mutate(avg_rent_to_income = avg_rent_2021/median_income_usd_2021) %>% 
+  mutate(med_rent_to_income = rent50_1/median_income_usd_2021)
 
@@
   # normalize rent to income
-  mutate(n_rent_to_income = (rent_to_income- min(rent_to_income)) / (max(rent_to_income) - min(rent_to_income))) %>% 
+  mutate(n_avg_rent_to_income = (avg_rent_to_income- min(avg_rent_to_income)) / (max(avg_rent_to_income) - min(avg_rent_to_income))) %>% 
+  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
   # normalize income to home price
-  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice)))
+  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
+  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21))/(max(avg_occupancy_21) - min(avg_occupancy_21)))
 
@@
 # invert normalized rent to income ratio so lower values (more favorable) have higher scores
+MSA_realEstate_data <- MSA_realEstate_data %>% 
+  mutate(med_rent_to_income_inverted = 1 - n_med_rent_to_income) %>% 
+  mutate(avg_rent_to_income_inverted = 1 - n_avg_rent_to_income)
 
-= 1- normalized score
 ```
@@
 # weights from owen
+real_estate_weights <- c(0.35, 0.20, 0.15, 0.10, 0.10, 0.10)
 
-# place holder column
-MSA_real_estate_score <- 
-  mutate
+#make real_estate_score column
+MSA_realEstate_data <- MSA_realEstate_data %>% 
+  mutate(real_estate_score = MSA) #this is just a place holder
+```
 
-weights <- c()
+```{r}
+size <- dim(MSA_realEstate_data)
 
+for (i in 1:size[1]) {
+  MSA_x_data <- c(MSA_realEstate_data$n_pop_change[i],
+                  MSA_realEstate_data$n_employ_change[i],
+                  MSA_realEstate_data$n_avg_occupancy[i],
+                  MSA_realEstate_data$n_rent_to_income[i],
+                  MSA_realEstate_data$n_income_to_homeprice[i]
+                  )
+  MSA_realEstate_data$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
+  }
 ```

### 2022-11-01 src/real_estate.Rmd
updating to remove NA values

@@
   # normalize rent to income
-  mutate(n_avg_rent_to_income = (avg_rent_to_income- min(avg_rent_to_income)) / (max(avg_rent_to_income) - min(avg_rent_to_income))) %>% 
+  mutate(n_avg_rent_to_income = (avg_rent_to_income- min(avg_rent_to_income, na.rm = TRUE)) / (max(avg_rent_to_income, na.rm = TRUE) - min(avg_rent_to_income, na.rm = TRUE))) %>% 
   mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
@@
   mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
-  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21))/(max(avg_occupancy_21) - min(avg_occupancy_21)))
+  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE))/(max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE)))
 
@@
 MSA_realEstate_data <- MSA_realEstate_data %>% 
-  mutate(med_rent_to_income_inverted = 1 - n_med_rent_to_income) %>% 
-  mutate(avg_rent_to_income_inverted = 1 - n_avg_rent_to_income)
+  mutate(n_med_rent_to_income_inverted = 1 - n_med_rent_to_income) %>% 
+  mutate(n_avg_rent_to_income_inverted = 1 - n_avg_rent_to_income)
 

### 2022-11-02 src/real_estate.Rmd
calculating real estate score

--- 2022-10-31 src/real_estate.Rmd
@@
 # select columns in demographic data
-MSA_demo_data <- MSA_demo_data %>% 
+MSA_demographics <- MSA_demo_data %>% 
   clean_names() %>% 
-    # calculate income to home price
-  mutate(income_to_homeprice = median_income_usd_2021/house_price_2021) %>% 
-  select(cbsa, msa, state, employment_yo_y21_22, pop_percent_change20_21, median_income_usd_2021, income_to_homeprice)
+  # convert house prices to be in thousands
+  mutate(house_price_2021 = house_price_2021 * 1000) %>% 
+  select(cbsa, msa, state, employment_yo_y21_22, pop_percent_change20_21, median_income_usd_2021, house_price_2021)
 ```
@@
   # calculate change in rent
-  mutate(rent_change21_22 = (avg_rent_2022- avg_rent_2021)/avg_rent_2021)
+  mutate(rent_change21_22 = ((avg_rent_2022- avg_rent_2021)/avg_rent_2021)*100)
 
@@
 ```{r}
+# add MSA_median rent data
+MSA_realEstate_compiled<- MSA_median_rent_data %>% 
+  select(cbsa, rent50_1) %>%
+  mutate(rent50_1 = as.numeric(rent50_1)) %>% 
+  left_join(MSA_realEstate_data) %>% 
+  left_join(MSA_demographics)
+```
+
+```{r}
 # combine real estate and demographic data by cbsa
-MSA_realEstate_data <- MSA_demo_data %>% 
-  left_join(MSA_realEstate_data) %>% 
+MSA_RE_ratios <- MSA_realEstate_compiled %>% 
+      # calculate income to home price
+  mutate(income_to_homeprice = median_income_usd_2021/house_price_2021) %>% 
   # calculate rent to income ratio
-  mutate(rent_to_income = avg_rent_2021/median_income_usd_2021)
+  mutate(avg_rent_to_income = avg_rent_2021/median_income_usd_2021) %>% 
+  mutate(med_rent_to_income = rent50_1/median_income_usd_2021)
 
@@
 # normalize columns into new data table
-MSA_realEstate_data <- MSA_realEstate_data %>% 
+n_MSA_realEstate <- MSA_RE_ratios %>% 
   # normalize population data
@@
   # normalize rent to income
-  mutate(n_rent_to_income = (rent_to_income- min(rent_to_income)) / (max(rent_to_income) - min(rent_to_income))) %>% 
+  mutate(n_avg_rent_to_income = (avg_rent_to_income- min(avg_rent_to_income, na.rm = TRUE)) / (max(avg_rent_to_income, na.rm = TRUE) - min(avg_rent_to_income, na.rm = TRUE))) %>% 
+  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
   # normalize income to home price
-  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice)))
+  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
+  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
+  mutate(n_rent_change = (rent_change21_22 - min(rent_change21_22, na.rm = TRUE)) / (max(rent_change21_22, na.rm = TRUE) - min(rent_change21_22, na.rm = TRUE)))
 
@@
 # invert normalized rent to income ratio so lower values (more favorable) have higher scores
-MSA_realEstate_data <- MSA_realEstate_data %>% 
-  mutate(income_to_rent_inverted = 1 - n_income_to_rent)
+n_MSA_realEstate <- n_MSA_realEstate %>% 
+  mutate(n_med_rent_to_income_inverted = 1 - n_med_rent_to_income) %>% 
+  mutate(n_avg_rent_to_income_inverted = 1 - n_avg_rent_to_income)
 
@@
 # weights from owen
-real_estate_weights <- c()
+real_estate_weights <- c(0.35, 0.20, 0.15, 0.10, 0.10, 0.10, 0.10)
 
 #make real_estate_score column
-MSA_realEstate_data <- MSA_realEstate_data %>% 
-  mutate(real_estate_score = MSA) #this is just a place holder
+n_MSA_realEstate <- n_MSA_realEstate %>% 
+  select(cbsa, msa, starts_with("n_")) %>% 
+  mutate(real_estate_score = msa) #this is just a place holder
 ```
@@
 ```{r}
-size <- dim(MSA_realEstate_data)
+size <- dim(n_MSA_realEstate)
 
-for (i in 1:size[1]) {
-  MSA_x_data <- c(MSA_realEstate_data$eviction_notice_days_normalized_inverted[i],
-                  MSA_realEstate_data$security_depopsit_limit_num[i],
-                  MSA_realEstate_data$notice_for_entry_days_num[i]
+weighted_RE_data <- n_MSA_realEstate %>% 
+  mutate(weighted_pop_change = 0.3 * n_pop_change) %>% 
+  mutate(weight_emp = 0.25 * n_employ_change)
+
+for (i in size[1]) {
+  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
+                  n_MSA_realEstate$n_employ_change[i],
+                  n_MSA_realEstate$n_avg_occupancy[i],
+                  n_MSA_realEstate$n_rent_change[i],
+                  n_MSA_realEstate$n_med_rent_to_income_inverted[i],
+                  n_MSA_realEstate$n_avg_rent_to_income_inverted[i], # compare avg rent and median rent
+                  n_MSA_realEstate$n_income_to_homeprice[i]
                   )
-  MSA_realEstate_data$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
+  n_MSA_realEstate$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
 }
+
 ```

### 2022-11-02 src/real_estate.Rmd
calculated separate real estate scores for median rent and avg rent

@@
 ```{r}
+# manually check weighted score
+n_MSA_realEstate<- n_MSA_realEstate %>% 
+  mutate(weighted_pop_change = 0.35 * n_pop_change) %>% 
+  mutate(weight_emp = 0.20 * n_employ_change) %>% 
+  mutate(weighted_occ = 0.15 * n_avg_occupancy) %>% 
+  mutate(weighted_rent_change = 0.10 * n_rent_change) %>% 
+  mutate(weighted_med_rent_income = 0.10 * n_med_rent_to_income_inverted) %>% 
+  mutate(weight_income_homeprice = 0.10 * n_income_to_homeprice) %>% 
+  mutate(total_score = weighted_pop_change + weight_emp + weighted_occ + weighted_rent_change + weighted_med_rent_income + weight_income_homeprice)
+```
+
+```{r}
 # weights from owen
-real_estate_weights <- c(0.35, 0.20, 0.15, 0.10, 0.10, 0.10, 0.10)
+real_estate_weights <- c(0.35, 0.20, 0.15, 0.10, 0.10, 0.10)
 
@@
 n_MSA_realEstate <- n_MSA_realEstate %>% 
-  select(cbsa, msa, starts_with("n_")) %>% 
-  mutate(real_estate_score = msa) #this is just a place holder
+  select(cbsa, msa, starts_with("n_"), total_score) %>% 
+  mutate(real_estate_score = msa) %>% #this is just a place holder
+  mutate(RE_score_avg_rent = msa)
 ```
@@
 
-weighted_RE_data <- n_MSA_realEstate %>% 
-  mutate(weighted_pop_change = 0.3 * n_pop_change) %>% 
-  mutate(weight_emp = 0.25 * n_employ_change)
-
-for (i in size[1]) {
+for (i in 1:size[1]) {
   MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
@@
                   n_MSA_realEstate$n_med_rent_to_income_inverted[i], 
-                  n_MSA_realEstate$n_avg_rent_to_income_inverted[i], # compare avg rent and median rent
                   n_MSA_realEstate$n_income_to_homeprice[i]
@@
 ```
+
+```{r}
+
+# score using average rent
+for (i in 1:size[1]) {
+  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
+                  n_MSA_realEstate$n_employ_change[i],
+                  n_MSA_realEstate$n_avg_occupancy[i],
+                  n_MSA_realEstate$n_rent_change[i],
+                  n_MSA_realEstate$n_avg_rent_to_income_inverted[i], # compare avg rent and median rent
+                  n_MSA_realEstate$n_income_to_homeprice[i]
+                  )
+  n_MSA_realEstate$RE_score_avg_rent[i] = weighted.mean(MSA_x_data, real_estate_weights)
+}
+```

### 2022-11-08 src/real_estate.Rmd
recalculated rent change

@@
   clean_names() %>% 
+   # calculate change in rent
+  mutate(rent_change = (x2022_rent - average_rent_2021)/average_rent_2021 * 100) %>% 
   group_by(cbsa) %>% 
@@
             avg_rent_2022 = mean(x2022_rent, na.rm = TRUE),
-            avg_occupancy_21 = mean(percent_occupancy_2021, na.rm = TRUE)) %>% 
-  # calculate change in rent
-  mutate(rent_change21_22 = ((avg_rent_2022- avg_rent_2021)/avg_rent_2021)*100)
+            avg_occupancy_21 = mean(percent_occupancy_2021, na.rm = TRUE),
+            avg_rent_change = mean(rent_change, na.rm = TRUE))
 

### 2022-11-08 src/real_estate.Rmd
cleaning up code

@@
   mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
-  mutate(n_rent_change = (rent_change21_22 - min(rent_change21_22, na.rm = TRUE)) / (max(rent_change21_22, na.rm = TRUE) - min(rent_change21_22, na.rm = TRUE)))
+  mutate(n_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))
 

### 2022-11-08 src/real_estate.Rmd
add column for median rent

@@
 MSA_realEstate_compiled<- MSA_median_rent_data %>% 
-  select(cbsa, rent50_1) %>%
-  mutate(rent50_1 = as.numeric(rent50_1)) %>% 
+  select(cbsa, rent50_0) %>%
+  mutate(median_rent = as.numeric(rent50_0)) %>% 
   left_join(MSA_realEstate_data) %>% 
@@
   mutate(avg_rent_to_income = avg_rent_2021/median_income_usd_2021) %>% 
-  mutate(med_rent_to_income = rent50_1/median_income_usd_2021)
+  mutate(med_rent_to_income = median_rent/median_income_usd_2021)
 

### 2022-11-10 src/demo_data_wrangling.Rmd
new list of MSAs/Cities with updated population sizeof 500,000

@@
   select("cbsa", "name", "popestimate2020", "popestimate2021", "residual2020", "residual2021") %>% # select data 
-  filter(popestimate2021 >= 300000) %>% 
+  filter(popestimate2021 >= 500000) %>% 
   separate(col = "name",

### 2022-11-11 src/demo_data_wrangling.Rmd
calculating normalized and employment rates

@@
   filter(!cbsa %in% c("10580", "10900", "12420", "12940", "14260", "15980", "18580", "19100", "19660", "19740", "23540", "25420", "26420", "27260", "29820", "34940", "36260", "36740", "38060", "39100", "39300", "39340", "41620", "45220", "45300", "46060", "46140", "39900")) %>%  # remove MSAs as instructed by Owen
-  select(cbsa, MSA, state, pop_percent_change20_21)
+  mutate(normalized_pop2021 = (popestimate2021-min(popestimate2021))/ (max(popestimate2021)-min(popestimate2021))) %>% 
+  select(cbsa, MSA, state, pop_percent_change20_21, normalized_pop2021)
 
@@
 # join population growth data set to household median income
-MSA_house_income <- MSA_income_data %>% # rename row names
+income_potential_MSAs <- MSA_income_data %>% # rename row names
+  select(cbsa, median_income) %>% 
   inner_join(potential_MSAs) %>% 
-  select(cbsa, MSA, state,median_income, income_moe_USD) 
+  select(cbsa, MSA, state, median_income, pop_percent_change20_21, normalized_pop2021) #reorder columns
 #select(cbsa, MSA, state, pop_percent_change20_21, median_income, income_moe_USD)
@@
 
-```{r, household income visual, fig.height = 7, fig.align='center'}
+```{r, household income visual, fig.height = 7, fig.align='center', eval = FALSE}
 ggplot(MSA_house_income, aes(x = median_income, y = reorder(MSA, median_income))) +
@@
 MSA_employment <- unemployment_data %>% 
-  filter(year >= 2015) %>% 
-  full_join(MSA_house_income) %>% 
+  filter(year >= 2020) %>% 
+  full_join(income_potential_MSAs) %>% 
   drop_na() %>% 
-  select(cbsa, MSA, year, month, employment_rate, median_income, income_moe_USD)
+  select(cbsa, MSA, year, month, employment_rate)
+
+avg_emp_yoy <- MSA_employment %>% 
+  group_by(MSA, year) %>% 
+  summarize(avg_employment_rate = mean(employment_rate)) %>% 
+  mutate(avg_emp_YoY = (avg_employment_rate-lag(avg_employment_rate))/lag(avg_employment_rate)*100) %>% 
+  drop_na()
 
@@
 MSA_employment_yoy <- MSA_employment %>% 
-  filter(year >= 2016) %>% 
+  filter(year >= 2020) %>% 
   filter(month == "01") %>% 
@@
   mutate(employment_YoY = round(YoY, digits = 3)) %>% 
-  mutate(emp_pos = employment_YoY >= 0) %>% # function to create column for yoy value, true = positive YoY
-  select(!c(YoY, month)) %>% 
-  drop_na()
+  mutate(emp_pos = employment_YoY >= 0) %>%# function to create column for yoy value, true = positive YoY
+  inner_join(avg_emp_yoy) %>% 
+  filter(year == "2022") %>% 
+  select(!c(YoY, month, year))
 
@@
 yoy <- (employment_data_jan_2022 - employment_data_jan_2021)/ employment_data_jan_2021 *100
+
+potential_MSAs <- potential_MSAs %>%
 
@@
   geom_line() +
-  xlim(2016, 2022) +
+  xlim(2021, 2022) +
   theme(legend.position = "none") +
@@
 
-```{r payroll data}
+```{r payroll data, eval = FALSE}
 # payroll data has not been added to final table
@@
 ```
-```{r payroll visual, fig.height = 7}
+```{r payroll visual, fig.height = 7, eval = FALSE}
 ggplot(MSA_payroll, aes(x = payroll_change, y = reorder(MSA, payroll_change), fill = payroll_pos)) +
@@
 
-```{r}
+```{r median rent}
 # median rents from hud
@@
   separate(col = cbsa,
-           into = c("cbsa", "unknown"),
+           into = c("unknown", "cbsa"),
            sep = -5) %>% 
   mutate(cbsa = as.double(cbsa)) %>% 
-  select(cbsa, rent50_0, rent50_1, rent50_2, rent50_3, rent50_4) %>% 
+  mutate(median_rent2021 = as.numeric(rent50_0)) %>% 
+  select(cbsa, median_rent2021) %>% 
   right_join(potential_MSAs)
+
+MSA_median_rent2022_data <- read_csv(here("data", "raw", "2022MSA_median_rent.csv")) %>% 
+  separate(col = cbsasub22,
+           into = c("metro", "cbsa"),
+           sep = 5) %>% 
+  separate(col = cbsa,
+           into = c("unknown", "cbsa"),
+           sep = -5) %>% 
+  mutate(cbsa = as.double(cbsa)) %>% 
+  mutate(median_rent2022 = as.numeric(rent_50_0)) %>% 
+  select(cbsa, median_rent2022)
+
+potential_MSAs <- potential_MSAs %>% 
+  right_join(MSA_median_rent2022_data) %>% 
+  drop_na()
+
 ```
@@
 
-```{r, house price visual, fig.height = 7, fig.align='center'}
+```{r, house price visual, fig.height = 7, fig.align='center', eval = FALSE}
 ggplot(metro_house_prices, aes(x = houseprice_change21_22, y = reorder(MSA, houseprice_change21_22))) +
@@
 
-```{r property taxes data wrangling}
+```{r property taxes data wrangling, eval = FALSE}
 property_tax_data <- read_csv(here("data","raw", "wide_property_tax_data.csv"))
@@
 
-```{r property tax visual, fig.height = 7, fig.align='center'}
+```{r property tax visual, fig.height = 7, fig.align='center', eval = FALSE}
 ggplot(MSA_property_tax, aes(x = median_tax_USD, y = reorder(MSA, median_tax_USD))) +
@@
   select(cbsa, year, employment_rate, employment_YoY) %>% 
-  right_join(MSA_property_tax) %>% 
+  # right_join(MSA_property_tax) %>% 
   select(!year)

### 2022-11-11 src/demo_data_wrangling.Rmd
updating population sixe, including normalized population and population estimate

@@
 
-# create __ 
+# create object with viable MSAs 
 potential_MSAs <- msa_pop_2021_data %>% 
@@
   mutate(normalized_pop2021 = (popestimate2021-min(popestimate2021))/ (max(popestimate2021)-min(popestimate2021))) %>% 
-  select(cbsa, MSA, state, pop_percent_change20_21, normalized_pop2021)
+  select(cbsa, MSA, state, pop_percent_change20_21, normalized_pop2021, popestimate2021)
 
@@
   inner_join(potential_MSAs) %>% 
-  select(cbsa, MSA, state, median_income, pop_percent_change20_21, normalized_pop2021) #reorder columns
-#select(cbsa, MSA, state, pop_percent_change20_21, median_income, income_moe_USD)
+  select(cbsa, MSA, state, median_income, pop_percent_change20_21, popestimate2021, normalized_pop2021) #reorder columns
 
@@
 MSA_employment <- unemployment_data %>% 
-  filter(year >= 2020) %>% 
+  filter(year >= 2019) %>% 
   full_join(income_potential_MSAs) %>% 
@@
   mutate(avg_emp_YoY = (avg_employment_rate-lag(avg_employment_rate))/lag(avg_employment_rate)*100) %>% 
+  mutate(avg_emp_YoY = round(avg_emp_YoY, digits = 3)) %>% 
   drop_na()
 
-# calculate YoY growth 
-MSA_employment_yoy <- MSA_employment %>% 
-  filter(year >= 2020) %>% 
-  filter(month == "01") %>% 
-  arrange(MSA, month) %>% 
-  group_by(MSA) %>% 
-  mutate(YoY = (employment_rate-lag(employment_rate))/lag(employment_rate)*100) %>% 
-  mutate(employment_YoY = round(YoY, digits = 3)) %>% 
-  mutate(emp_pos = employment_YoY >= 0) %>%# function to create column for yoy value, true = positive YoY
-  inner_join(avg_emp_yoy) %>% 
-  filter(year == "2022") %>% 
-  select(!c(YoY, month, year))
+MSA_emp_income <- income_potential_MSAs %>% 
+  right_join(avg_emp_yoy) %>% 
+  filter(year == "2021") %>% 
+  select(!c(avg_employment_rate, year))
 
@@
 
-potential_MSAs <- potential_MSAs %>%
-
 ```
 
-```{r eval = FALSE}
+```{r employment visuals, fig.align = "center", fig.show="hold", fig.height= 7, fig.width=10, eval = FALSE}
 # average employment from 2000 to 2022
@@
 
-```
-
-```{r employment visuals, fig.align = "center", fig.show="hold", fig.height= 7, fig.width=10}
 # visual of MSA employment rates, shows major drop during covid
@@
   labs(x = "Year-over_year change in employment rate",
-       y = "")
+       y = " ")
 
@@
 ```
+
 ```{r payroll visual, fig.height = 7, eval = FALSE}
@@
   mutate(median_rent2022 = as.numeric(rent_50_0)) %>% 
-  select(cbsa, median_rent2022)
+  select(cbsa, median_rent2022) %>% 
+  right_join(MSA_median_rent2021_data)
 
-potential_MSAs <- potential_MSAs %>% 
-  right_join(MSA_median_rent2022_data) %>% 
-  drop_na()
+potential_MSAs <- MSA_median_rent2022_data %>% 
+  left_join(MSA_emp_income)
 
@@
 ```{r house price data}
-metro_house_cost_data <- read_csv(here("data","raw", "metro_home_costs", "metro_home_prices.csv")) %>% 
+metro_house_cost_data <- read_csv(here("data","raw", "metro_home_costs", "metro-home-prices-2022-11-10.csv")) %>% 
   row_to_names(3) %>% 
@@
   mutate(cbsa = as.numeric(cbsa)) %>% 
-  mutate(x2020 = as.numeric(x2020))
+  mutate(x2021 = as.numeric(x2021)) %>% 
+  mutate(x2021 = x2021 * 1000) %>% 
+  rename(house_price_2021 = x2021)
 
-metro_house_prices <- MSA_house_income %>%
-  left_join(metro_house_cost_data)  %>% 
-  rename(house_price_2020 = x2020) %>% 
-  rename(house_price_2021 = x2021_r) %>% 
-  rename(houseprice_change21_22 = q1_q1) %>% 
-  select(!c(metro_area, x2019, na)) %>% 
-  # select(cbsa, houseprice_change21_22) %>% 
-  separate(col = "houseprice_change21_22",
-           into = "houseprice_change21_22",
-           sep = "%") %>% 
-  mutate(houseprice_change21_22 = as.numeric(houseprice_change21_22))
+metro_house_prices <- metro_house_cost_data %>%
+  select(cbsa, house_price_2021) %>% 
+  right_join(potential_MSAs) %>% 
+  select(cbsa, MSA, state, pop_percent_change20_21, normalized_pop2021, popestimate2021, median_income, avg_emp_YoY, median_rent2021, median_rent2022, house_price_2021)
 ```
@@
 
-```{r join population, unemployment, income, and house price data}
-# join employment data with income, house prices, tax, and population data
-MSApop_income_housecost_tax_emp <- MSA_employment_yoy %>%
-  filter(year == "2022") %>% 
-  mutate(employment_rate = round(employment_rate, digits = 3)) %>% 
-  select(cbsa, year, employment_rate, employment_YoY) %>% 
-  # right_join(MSA_property_tax) %>% 
-  select(!year)
-
-```
-
 ### MSA data table
@@
 # table seperated by MSAs
-datatable(MSApop_income_housecost_tax_emp, extensions = 'Buttons', options = list(
+datatable(metro_house_prices, extensions = 'Buttons', options = list(
     dom = 'Bfrtip',
@@
 #  split up MSAs into cities to collect utility, rent, and landlord data.
-msa_seperated <- MSApop_income_housecost_tax_emp %>% 
+msa_seperated <- metro_house_prices %>% 
   separate_rows(MSA, sep = "-") %>% 

### 2022-11-13 src/landlord.Rmd
check scoring calculation
FILTERED TO ONLY INCLUDE NEW MSA LIST (POP >500,000) 29 MSAS TOTAL. CHECKED LANDLORD SCORE CALCULATION.

@@
 MSA_landlord_data <- read_sheet("https://docs.google.com/spreadsheets/d/1zCePYRyJ5lb1Z5jvCujg17EtwqpxJN1E1_HwJL66uLM/edit#gid=122953884")
+
+select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")
 ```
@@
 ```{r}
-# #rent_control
-# MSA_landlord_data <- MSA_landlord_data %>% 
-#   mutate(
-#     rent_control_binary = case_when(
-#       rent_control == "N" ~ 1, 
-#       rent_control == "Y" ~ 0))
+#filter by new MSA list
+new_MSA_list <- unique(select_MSAs$cbsa)
 
+MSA_landlord_data <- MSA_landlord_data %>% 
+  filter(cbsa %in% c(new_MSA_list))
+```
+
+```{r}
 #security_deposit_limit
@@
 ```{r}
-#PICK UP HERE
-#If there is rent control, immediately disqualify location 
+#If there is rent control, immediately disqualify location (set score to -999)
+for (j in 1:size[1]) {
+  if(MSA_landlord_data$rent_control[j] == "Y"){
+    MSA_landlord_data$landlord_score[j] = -999
+  }
+}
 ```

### 2022-11-13 src/real_estate.Rmd
multiplied monthly rents by 12

@@
   # calculate rent to income ratio
-  mutate(avg_rent_to_income = avg_rent_2021/median_income_usd_2021) %>% 
-  mutate(med_rent_to_income = median_rent/median_income_usd_2021)
+  # multiply monthly rent x 12 for annual rent!
+  mutate(avg_rent_to_income = (avg_rent_2021*12)/median_income_usd_2021) %>% 
+  mutate(med_rent_to_income = (median_rent*12)/median_income_usd_2021)
 
@@
   mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
+  # normalize occupancy data
   mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
+  # normalize rent change
   mutate(n_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))

### 2022-11-13 src/real_estate.Rmd
Filtered by MSAs with pop>500,000 at the end (put 29 MSAs data in new table).

@@
 MSA_median_rent_data <- read_sheet("https://docs.google.com/spreadsheets/d/1oxNxWQRWpANbgb603ShBlw84QkyvsRaH4TzHKKVWhxw/edit#gid=0")
+
+select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")
 ```
@@
 ```
+
+```{r}
+#FILTER BY NEW MSA LIST
+new_MSA_list <- unique(select_MSAs$cbsa)
+
+n_MSA_realEstate_select_MSAs <- n_MSA_realEstate %>% 
+  filter(cbsa %in% c(new_MSA_list))
+```

### 2022-11-13 src/real_estate.Rmd
updating scores for updated list of MSAs

--- 2022-11-08 src/real_estate.Rmd
@@
 
-MSA_demo_data <- read_sheet("https://docs.google.com/spreadsheets/d/12do9sRCAcrIe0K-VcJWXNmmTm_jg5rKjZ0NeJPOm9Jw/edit#gid=0")
-
 city_realEstate_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990")
 
-MSA_median_rent_data <- read_sheet("https://docs.google.com/spreadsheets/d/1oxNxWQRWpANbgb603ShBlw84QkyvsRaH4TzHKKVWhxw/edit#gid=0")
-```
+select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")
 
-```{r}
-# select columns in demographic data
-MSA_demographics <- MSA_demo_data %>% 
-  clean_names() %>% 
-  # convert house prices to be in thousands
-  mutate(house_price_2021 = house_price_2021 * 1000) %>% 
-  select(cbsa, msa, state, employment_yo_y21_22, pop_percent_change20_21, median_income_usd_2021, house_price_2021)
 ```
@@
   clean_names() %>% 
+  rename(avg_rent2022 = x2022_rent) %>% 
    # calculate change in rent
-  mutate(rent_change = (x2022_rent - average_rent_2021)/average_rent_2021 * 100) %>% 
+  mutate(rent_change = (avg_rent2022 - average_rent_2021)/average_rent_2021 * 100) %>% 
   group_by(cbsa) %>% 
@@
   summarize(avg_rent_2021 = mean(average_rent_2021, na.rm = TRUE),
-            avg_rent_2022 = mean(x2022_rent, na.rm = TRUE),
+            avg_rent_2022 = mean(avg_rent2022, na.rm = TRUE),
             avg_occupancy_21 = mean(percent_occupancy_2021, na.rm = TRUE),
@@
 # add MSA_median rent data
-MSA_realEstate_compiled<- MSA_median_rent_data %>% 
-  select(cbsa, rent50_0) %>%
-  mutate(median_rent = as.numeric(rent50_0)) %>% 
+compliedMSA_realEstate<- select_MSAs %>% 
   left_join(MSA_realEstate_data) %>% 
-  left_join(MSA_demographics)
-```
-
-```{r}
-# combine real estate and demographic data by cbsa
-MSA_RE_ratios <- MSA_realEstate_compiled %>% 
-      # calculate income to home price
-  mutate(income_to_homeprice = median_income_usd_2021/house_price_2021) %>% 
+    # calculate rent change
+  mutate(med_rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100) %>% 
+  # calculate income to homeprice
+  mutate(income_to_homeprice = median_income/house_price_2021) %>% 
   # calculate rent to income ratio
-  mutate(avg_rent_to_income = avg_rent_2021/median_income_usd_2021) %>% 
-  mutate(med_rent_to_income = median_rent/median_income_usd_2021)
+  # multiply monthly rent x 12 for annual rent!
+  mutate(avg_rent_to_income = (avg_rent_2021*12)/median_income) %>% 
+  mutate(med_rent_to_income = (median_rent2021*12)/median_income)
 
@@
 # normalize columns into new data table
-n_MSA_realEstate <- MSA_RE_ratios %>% 
+n_MSA_realEstate <- compliedMSA_realEstate %>% 
   # normalize population data
@@
   # normalize employment growth
-  mutate(n_employ_change = (employment_yo_y21_22 - min(employment_yo_y21_22)) / (max(employment_yo_y21_22) - min(employment_yo_y21_22))) %>% 
+  mutate(n_employ_change = (avg_emp_YoY - min(avg_emp_YoY)) / (max(avg_emp_YoY) - min(avg_emp_YoY))) %>% 
   # normalize rent to income
-  mutate(n_avg_rent_to_income = (avg_rent_to_income- min(avg_rent_to_income, na.rm = TRUE)) / (max(avg_rent_to_income, na.rm = TRUE) - min(avg_rent_to_income, na.rm = TRUE))) %>% 
   mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
   # normalize income to home price
+  mutate(n_med_rent_change = (med_rent_change - min(med_rent_change)) / (max(med_rent_change) - min(med_rent_change))) %>% 
   mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
+  mutate(n_avg_rent_to_income = (avg_rent_to_income- min(avg_rent_to_income, na.rm = TRUE)) / (max(avg_rent_to_income, na.rm = TRUE) - min(avg_rent_to_income, na.rm = TRUE))) %>% 
   mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
-  mutate(n_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))
+  # normalize rent change
+  mutate(n_avg_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))
 
@@
   mutate(weighted_occ = 0.15 * n_avg_occupancy) %>% 
-  mutate(weighted_rent_change = 0.10 * n_rent_change) %>% 
+  mutate(weighted_avg_rent_change = 0.10 * n_avg_rent_change) %>% 
+  mutate(weighted_med_rent_change = 10* n_med_rent_change) %>% 
   mutate(weighted_med_rent_income = 0.10 * n_med_rent_to_income_inverted) %>% 
   mutate(weight_income_homeprice = 0.10 * n_income_to_homeprice) %>% 
-  mutate(total_score = weighted_pop_change + weight_emp + weighted_occ + weighted_rent_change + weighted_med_rent_income + weight_income_homeprice)
+  mutate(med_total_score = weighted_pop_change + weight_emp + weighted_occ + weighted_med_rent_change + weighted_med_rent_income + weight_income_homeprice) %>% 
+  mutate(avg_total_score = weighted_pop_change + weight_emp + weighted_occ + weighted_avg_rent_change + weighted_avg_rent_income + weight_income_homeprice)
 ```
@@
 n_MSA_realEstate <- n_MSA_realEstate %>% 
-  select(cbsa, msa, starts_with("n_"), total_score) %>% 
-  mutate(real_estate_score = msa) %>% #this is just a place holder
+  select(cbsa, MSA, starts_with("n_"), total_score) %>% 
+  mutate(real_estate_score_med = msa) %>% #this is just a place holder
   mutate(RE_score_avg_rent = msa)
@@
                   )
-  n_MSA_realEstate$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
+  n_MSA_realEstate$real_estate_score_med[i] = weighted.mean(MSA_x_data, real_estate_weights)
 }

### 2022-11-13 src/real_estate.Rmd
created total scores columns to avg and median rent

@@
 ```{r}
-# manually check weighted score
-n_MSA_realEstate<- n_MSA_realEstate %>% 
-  mutate(weighted_pop_change = 0.35 * n_pop_change) %>% 
-  mutate(weight_emp = 0.20 * n_employ_change) %>% 
-  mutate(weighted_occ = 0.15 * n_avg_occupancy) %>% 
-  mutate(weighted_avg_rent_change = 0.10 * n_avg_rent_change) %>% 
-  mutate(weighted_med_rent_change = 10* n_med_rent_change) %>% 
-  mutate(weighted_med_rent_income = 0.10 * n_med_rent_to_income_inverted) %>% 
-  mutate(weight_income_homeprice = 0.10 * n_income_to_homeprice) %>% 
-  mutate(med_total_score = weighted_pop_change + weight_emp + weighted_occ + weighted_med_rent_change + weighted_med_rent_income + weight_income_homeprice) %>% 
-  mutate(avg_total_score = weighted_pop_change + weight_emp + weighted_occ + weighted_avg_rent_change + weighted_avg_rent_income + weight_income_homeprice)
+MSA_n_data_avg_rent <- n_MSA_realEstate %>% 
+  select(cbsa, MSA, n_pop_change, n_employ_change, n_avg_occupancy, n_avg_rent_change, n_avg_rent_to_income_inverted, n_income_to_homeprice)
+
+MSA_n_data_med_rent <- n_MSA_realEstate %>% 
+  select(cbsa, MSA, n_pop_change, n_employ_change, n_avg_occupancy, n_med_rent_change, n_med_rent_to_income_inverted, n_income_to_homeprice)
 ```
@@
 n_MSA_realEstate <- n_MSA_realEstate %>% 
-  select(cbsa, MSA, starts_with("n_"), total_score) %>% 
-  mutate(real_estate_score_med = msa) %>% #this is just a place holder
-  mutate(RE_score_avg_rent = msa)
+  select(cbsa, MSA, starts_with("n_")) %>% 
+  mutate(real_estate_score = MSA) %>% #this is just a place holder
+  mutate(RE_score_avg_rent = MSA)
 ```
@@
                   n_MSA_realEstate$n_avg_occupancy[i],
-                  n_MSA_realEstate$n_rent_change[i],
+                  n_MSA_realEstate$n_med_rent_change[i],
                   n_MSA_realEstate$n_med_rent_to_income_inverted[i], 
@@
                   )
-  n_MSA_realEstate$real_estate_score_med[i] = weighted.mean(MSA_x_data, real_estate_weights)
+  n_MSA_realEstate$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
 }
@@
                   n_MSA_realEstate$n_avg_occupancy[i],
-                  n_MSA_realEstate$n_rent_change[i],
+                  n_MSA_realEstate$n_avg_rent_change[i],
                   n_MSA_realEstate$n_avg_rent_to_income_inverted[i], # compare avg rent and median rent

### 2022-11-13 src/demo_data_wrangling.Rmd
calculating portland ME employment data

@@
 MSA_emp_income <- income_potential_MSAs %>% 
-  right_join(avg_emp_yoy) %>% 
+  full_join(avg_emp_yoy) %>% 
   filter(year == "2021") %>% 
@@
 # adding portland employment rates from https://data.bls.gov/pdq/SurveyOutputServlet
-portland_employment_data <- read_csv(here("data", "raw", "portland_me_employment.csv")) %>% 
-  row_to_names(row_number = 10) %>% 
-  clean_names() %>% 
-  select(year, jan) %>% 
-  rename(civilian_labor = jan)
+portland_employment_data <- unemployment_data %>% 
+  filter(year >= 2019) %>% 
+  filter(cbsa == "76750") %>% 
+   group_by(year) %>% 
+  summarize(avg_employment_rate = mean(employment_rate)) %>% 
+  mutate(avg_emp_YoY = (avg_employment_rate-lag(avg_employment_rate))/lag(avg_employment_rate)*100) %>% 
+  mutate(avg_emp_YoY = round(avg_emp_YoY, digits = 3))
 

### 2022-11-14 src/MSAtotal_scores.Rmd
added function to print datatable. Double checked landlord policy

---
title: "Total Score"
author: "Grace Bianchi"
date: '2022-11-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(googlesheets4)

```

```{r}
# read in the data
category_score_data <- read_sheet(https://docs.google.com/spreadsheets/d/1zu6ta_gI5lF3l2RC5eE3BuUsywYg8gd6xcMfuTKb9Bk/edit#gid=852316062)

```

### 2022-11-14 src/landlord.Rmd
saved data to google sheets in r folder on google drive
added function to print datatable. Double checked landlord policy

@@
 library(kableExtra)
+library(DT)
 
@@
 }
+
+datatable(MSA_landlord_data, extensions = 'Buttons', options = list(
+    dom = 'Bfrtip',
+    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
+  )
+)
 ```

### 2022-11-14 src/MSAtotal_scores.Rmd
cr3eating rmd for total score; reading in real estate and landlord score

@@
 # read in the data
-category_score_data <- read_sheet(https://docs.google.com/spreadsheets/d/1zu6ta_gI5lF3l2RC5eE3BuUsywYg8gd6xcMfuTKb9Bk/edit#gid=852316062)
+real_estate_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit#gid=2007178520")
+
+landlord_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1nLpDhM88a3rEBDVK5RZlk2B6QuwVIwvsfE5HeBz0mLo/edit#gid=983090810")
 

### 2022-11-14 src/demo_data_wrangling.Rmd
collected demographic data for MSAs with populations greater than 300.000

@@
   select("cbsa", "name", "popestimate2020", "popestimate2021", "residual2020", "residual2021") %>% # select data
-  filter(popestimate2021 >= 500000) %>% 
+
   separate(col = "name",
@@
   right_join(potential_MSAs) %>% 
-  select(cbsa, MSA, state, pop_percent_change20_21, normalized_pop2021, popestimate2021, median_income, avg_emp_YoY, median_rent2021, median_rent2022, house_price_2021)
+  select(cbsa, MSA, state, pop_percent_change20_21, normalized_pop2021, popestimate2021, median_income, avg_emp_YoY, median_rent2021, median_rent2022, house_price_2021) %>% 
+  filter(popestimate2021 >= 300000)
 ```

### 2022-11-15 src/demo_data_wrangling.Rmd
cleaning up data (removed pay roll and tax info)

@@
 Initial analysis of Metropolitan Statistical Areas (MSAs) relied on Census data of annual population estimates, in which the population had to be greater than 300,000 individuals for the purpose of attracting investors.
-Population change from 2020 to 2021 was calculated by $$\frac{2021 estimate- 2020 estimate}{2021 estimate} * 100$$
+Population change from 2020 to 2021 was calculated by $$\frac{2021 estimate- 2020 estimate}{2020 estimate} * 100$$
 MSAs with a population growth > 0.5%, were selected to gather further data on. The 0.5% growth was determined.
@@
 # read in the data MSA population data from Census
-msa_pop_2021_data <- read_csv(here("data","raw", "msa_2021_alldata.csv")) %>% 
+msa_pop_2021_data <- read_csv(here("data","raw", "msa_2021_pop_data.csv")) %>% 
   clean_names() %>% # data wrangling to make column headers in lower_snake_case
@@
            sep = ", ") %>% 
-  mutate(pop_percent_change = (popestimate2021-popestimate2020)/popestimate2021 * 100)  # add population % change column
+  mutate(pop_percent_change = (popestimate2021-popestimate2020)/popestimate2020 * 100)  # add population % change column
 
@@
   filter(pop_percent_change > 0.5 ) %>% 
+  filter(popestimate2021 >= 300000) %>% 
   mutate(pop_percent_change20_21 = round(pop_percent_change, digits = 3)) %>% 
@@
 
-```{r fig.height = 7, fig.align='center', eval = FALSE}
+```{r population change visual, fig.height = 7, fig.align='center', eval = FALSE}
 ggplot(data = positive_MSA_pop_change, aes(x = pop_percent_change20_21, y = reorder(MSA, pop_percent_change20_21))) + 
@@
 
-Dataset: Median Household Income (USD) INCOME IN THE PAST 12 MONTHS (IN 2020 INFLATION-ADJUSTED DOLLARS)
+Dataset: Median Household Income (USD) INCOME IN THE PAST 12 MONTHS (IN 2021 INFLATION-ADJUSTED DOLLARS)
 
@@
 ```{r household income data wrangling}
-MSA_income_data <- read_csv(here("data","raw", "MSA_income", "2021_data", "2021_income_data.csv")) %>% 
+MSA_income_data <- read_csv(here("data","raw", "2021_income_data.csv")) %>% 
   row_to_names(1) %>% 
@@
   mutate(avg_emp_YoY = (avg_employment_rate-lag(avg_employment_rate))/lag(avg_employment_rate)*100) %>% 
-  mutate(avg_emp_YoY = round(avg_emp_YoY, digits = 3)) %>% 
+  mutate(avg_emp_YoY = round(avg_emp_YoY, digits = 4)) %>% 
   drop_na()
+
+yoy_test <- MSA_employment %>% 
+  group_by(MSA, year) %>% 
+  summarize(avg_employment_rate = mean(employment_rate)) %>% 
+  filter(year >= "2020") %>% 
+  pivot_wider(names_from = year,
+              values_from = avg_employment_rate,
+              names_prefix = "avg_employment_rate") %>% 
+  mutate(avg_emp_yoy = (avg_employment_rate2021 - avg_employment_rate2020)/avg_employment_rate2020 *100)
 
@@
   mutate(avg_emp_YoY = (avg_employment_rate-lag(avg_employment_rate))/lag(avg_employment_rate)*100) %>% 
-  mutate(avg_emp_YoY = round(avg_emp_YoY, digits = 3))
+  mutate(avg_emp_YoY = round(avg_emp_YoY, digits = 4))
 
@@
 ```{r employment visuals, fig.align = "center", fig.show="hold", fig.height= 7, fig.width=10, eval = FALSE}
-# average employment from 2000 to 2022
-employment_summary_since2000 <- MSA_employment %>% 
-  group_by(MSA) %>% 
-  filter(year >= 2000) %>% 
-  summarize(
-    avg_employment = mean(employment_rate),
-    sd_employment = sd(employment_rate)
-  )
-
 # visual of MSA employment rates, shows major drop during covid
@@
        y = " ")
-
-```
-
-```{r payroll data, eval = FALSE}
-# payroll data has not been added to final table
-payroll_data <- read_csv(here("data","raw", "employee_payroll_MSA21_22.csv")) %>% 
-  row_to_names(4) %>% 
-  clean_names() %>% 
-  rename(MSA = na) %>% 
-  rename(payroll_change_may21 = may_2021) %>% 
-  rename(payroll_change_mar22 = mar_2022) %>% 
-  rename(payroll_change_apr22 = apr_2022) %>% 
-  rename(payroll_change_may22 = may_2022_p)
-
-MSA_payroll <- payroll_data %>% 
-  full_join(potential_MSAs) %>% 
-  drop_na() %>% 
-  pivot_longer(cols = 2:5,
-               names_prefix = "payroll_change_",
-               names_to = "year",
-               values_to = "payroll_change") %>% 
-  mutate(payroll_change = as.numeric(payroll_change)) %>% 
-  group_by(MSA, year)  %>% 
-  mutate(payroll_pos = payroll_change >= 0)
-
-```
-
-```{r payroll visual, fig.height = 7, eval = FALSE}
-ggplot(MSA_payroll, aes(x = payroll_change, y = reorder(MSA, payroll_change), fill = payroll_pos)) +
-  geom_col(position = "identity", color = "black", size = 0.25) +
-  scale_fill_manual(values = c("red", "darkolivegreen3"), guide = FALSE) +
-  labs(x = "percent change",
-       y = "",
-       title = "Quarterly change in employee payroll from 2021 to 2022") +
-  facet_grid(~year)
 
@@
 
-Median housing prices by MSA were obtained from the Association of Realtors for years 2019, 2020, and 2021.
-The percent difference in 2021 Q1 and 2022 Q1 was used to show the change in price in 2021.
-**Next quarterly release will be 11/10**
+Median housing prices by MSA were obtained from the Association of Realtors for 2021.
 
@@
   select(cbsa, MSA, state, pop_percent_change20_21, normalized_pop2021, popestimate2021, median_income, avg_emp_YoY, median_rent2021, median_rent2022, house_price_2021) %>% 
-  filter(popestimate2021 >= 300000)
+  filter(popestimate2021 >= 500000)
 ```

### 2022-11-18 src/demo_data_wrangling.Rmd
output code directly to google sheet that is used for the scoring Rmds

@@
 library(kableExtra)
-
+library(googlesheets4)
 library(tigris) # Census shape files
@@
   filter(year >= 2019) %>% 
+  mutate(cbsa = ifelse(cbsa == '76750', "38860", cbsa)) %>% # change south portland, ME to corresponding cbsa
+  mutate(cbsa = as.numeric(cbsa)) %>% 
   full_join(income_potential_MSAs) %>% 
@@
   drop_na()
-
-yoy_test <- MSA_employment %>% 
-  group_by(MSA, year) %>% 
-  summarize(avg_employment_rate = mean(employment_rate)) %>% 
-  filter(year >= "2020") %>% 
-  pivot_wider(names_from = year,
-              values_from = avg_employment_rate,
-              names_prefix = "avg_employment_rate") %>% 
-  mutate(avg_emp_yoy = (avg_employment_rate2021 - avg_employment_rate2020)/avg_employment_rate2020 *100)
 
@@
            sep = -5) %>% 
-  mutate(cbsa = as.double(cbsa)) %>% 
+  mutate(cbsa = ifelse(cbsa == 'M6400', "38860", cbsa)) %>% 
+  mutate(cbsa = as.numeric(cbsa)) %>% 
   mutate(median_rent2021 = as.numeric(rent50_0)) %>% 
@@
            sep = -5) %>% 
-  mutate(cbsa = as.double(cbsa)) %>% 
+  mutate(cbsa = ifelse(unknown == '38860M', "38860", cbsa)) %>% 
   mutate(median_rent2022 = as.numeric(rent_50_0)) %>% 
   select(cbsa, median_rent2022) %>% 
+  mutate(cbsa = as.numeric(cbsa)) %>% 
   right_join(MSA_median_rent2021_data)
@@
   select(cbsa, MSA, state, pop_percent_change20_21, normalized_pop2021, popestimate2021, median_income, avg_emp_YoY, median_rent2021, median_rent2022, house_price_2021) %>% 
+  mutate(house_price_2021 = ifelse(cbsa == "12260", "284900", house_price_2021)) %>% 
+  mutate(house_price_2021 = ifelse(cbsa == "32580", "259000", house_price_2021))
+
+MSAs <-metro_house_prices %>% 
   filter(popestimate2021 >= 500000)
-```
-
-```{r house price exploratory, eval = FALSE}
-ggplot(data = metro_house_cost_data, aes(x = log(x2020))) +
-   geom_histogram(bins = 100)
-
-ggplot(data = metro_house_cost_data, aes(sample = log(x2020))) +
-  geom_qq()
-
 ```
@@
 - Salisbury, MD
-- Savannah, GA
 
@@
 
-### Property Taxes
-It was difficult to find property tax information by city that was consistent across cities.
-Median real estate taxes in dollars are provided by the U.S.
-Census and data is collected at the MSA level.
-A metric of property tax and median home prices can be used to normalized.
-
-Dataset: B25103  MORTGAGE STATUS BY MEDIAN REAL ESTATE TAXES PAID (USD) in 2019
-
-Source:[U.S. Census Bureau](https://data.census.gov/cedsci/table?q=property%20tax&g=0100000US%2431000M1&d=ACS%201-Year%20Estimates%20Detailed%20Tables&tid=ACSDT1Y2019.B25103&tp=true)
-
-```{r property taxes data wrangling, eval = FALSE}
-property_tax_data <- read_csv(here("data","raw", "wide_property_tax_data.csv"))
-
-# create new data frame with tax estimate values
-median_tax_df <- property_tax_data %>% 
-  select(ends_with("estimate")) %>% 
-  slice(2) %>% 
-  stack() %>% # flip data table to be tidy
-  rename(MSA = ind) %>% 
-  mutate(median_tax_USD = as.numeric(values)) %>% 
-  separate(col = MSA,
-           into = c("MSA", "state"),
-           sep = ", ") %>% 
-  separate(col = state,
-           into = c("state"),
-           sep = " ") %>% 
-  select(!values)
-
-# make data frame with margin of error values
-moe_tax_df <- property_tax_data %>% 
-  select(ends_with("error")) %>% 
-  slice(2) %>% 
-  stack() %>% # flip data table to be tidy
-  rename(MSA = ind) %>% 
-  separate(col = values,
-           into = c("value", "tax_moe_USD"),
-           sep = 1) %>% 
-  mutate(tax_moe_USD = as.numeric(tax_moe_USD)) %>% 
-  separate(col = MSA,
-           into = c("MSA", "state"),
-           sep = ", ") %>% 
-  separate(col = state,
-           into = c("state", "area"),
-           sep = " ") %>% 
-  select(!c(value, area))
-
-MSA_property_tax <- metro_house_prices %>% 
-  full_join(median_tax_df) %>% 
-  left_join(moe_tax_df) %>% 
-  drop_na(cbsa) # drop all MSA rows if NA in cbsa column
-
-```
-
-```{r property tax visual, fig.height = 7, fig.align='center', eval = FALSE}
-ggplot(MSA_property_tax, aes(x = median_tax_USD, y = reorder(MSA, median_tax_USD))) +
-  geom_errorbarh(aes(xmin = median_tax_USD - tax_moe_USD, xmax = median_tax_USD + tax_moe_USD)) + 
-  geom_point(size = 3, color = "darkgreen") + 
-  theme_minimal(base_size = 12.5) + 
-  labs(title = "Median property tax in 2019", 
-       subtitle = "", 
-       x = "Median Property Tax estimate (in USD)", 
-       y = "") + 
-  scale_x_continuous(labels = label_dollar())
-```
-
 ### MSA data table
@@
 # table seperated by MSAs
-datatable(metro_house_prices, extensions = 'Buttons', options = list(
+datatable(MSAs, extensions = 'Buttons', options = list(
     dom = 'Bfrtip',
@@
 )
+
+gs4_create("updated_MSAs")
+sheet_write(MSAs, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "MSAs")
 ```
@@
   separate_rows(MSA, sep = "-") %>% 
-  rename(city = MSA)# %>% # seperate MSA cities into their own columns
- # full_join(payroll_data, by = c("MSA" = "area")) %>% 
- # drop_na(pop_percent_change20_21)
+  rename(city = MSA) %>% # seperate MSA cities into their own columns
+ drop_na(city)
+
+gs4_create("updated_MSAs")
+sheet_write(msa_seperated, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "cities")
 

### 2022-11-18 src/demo_data_wrangling.Rmd
updating score for MSAs without occupancy rates

@@
 MSAs <-metro_house_prices %>% 
-  filter(popestimate2021 >= 500000)
+  filter(popestimate2021 >= 500000) %>% 
+  distinct(cbsa, .keep_all = TRUE) # remove duplicate portland row
 ```
@@
 
-gs4_create("updated_MSAs")
+#gs4_create("updated_MSAs")
 sheet_write(MSAs, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "MSAs")
@@
 #  split up MSAs into cities to collect utility, rent, and landlord data.
-msa_seperated <- metro_house_prices %>% 
+msa_seperated <- MSAs %>% 
   separate_rows(MSA, sep = "-") %>% 
-  rename(city = MSA) %>% # seperate MSA cities into their own columns
- drop_na(city)
+  rename(city = MSA) %>% # separate MSA cities into their own columns
+  filter(! city == "")
 
-gs4_create("updated_MSAs")
+#gs4_create("updated_MSAs")
 sheet_write(msa_seperated, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "cities")

### 2022-11-18 src/real_estate.Rmd
saved data to google sheets in r folder on google drive
added function to print datatable. Double checked landlord policy
updating score for MSAs without occupancy rates

@@
 library(googlesheets4)
+library(DT)
 
@@
   # calculate income to homeprice
+  mutate(house_price_2021 = as.numeric(house_price_2021)) %>% 
   mutate(income_to_homeprice = median_income/house_price_2021) %>% 
@@
 ```{r}
-MSA_n_data_avg_rent <- n_MSA_realEstate %>% 
-  select(cbsa, MSA, n_pop_change, n_employ_change, n_avg_occupancy, n_avg_rent_change, n_avg_rent_to_income_inverted, n_income_to_homeprice)
-
-MSA_n_data_med_rent <- n_MSA_realEstate %>% 
-  select(cbsa, MSA, n_pop_change, n_employ_change, n_avg_occupancy, n_med_rent_change, n_med_rent_to_income_inverted, n_income_to_homeprice)
-```
-
-```{r}
 # weights from owen
@@
 n_MSA_realEstate <- n_MSA_realEstate %>% 
-  select(cbsa, MSA, starts_with("n_")) %>% 
+  select(cbsa, MSA, starts_with("n_"), ) %>% 
   mutate(real_estate_score = MSA) %>% #this is just a place holder
@@
 ```
+
+```{r}
+# create downloadable data table
+datatable(n_MSA_realEstate, extensions = 'Buttons', options = list(
+    dom = 'Bfrtip',
+    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
+  )
+)
+```
+
+```{r}
+#write result to google sheet
+# gs4_create("MSA_real_estate_score")
+sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "n_MSA_realEstate")
+```

### 2022-11-21 src/demo_data_wrangling.Rmd
Use median rent of 1 bedroom apartment

@@
   mutate(cbsa = as.numeric(cbsa)) %>% 
-  mutate(median_rent2021 = as.numeric(rent50_0)) %>% 
+  mutate(median_rent2021 = as.numeric(rent50_1)) %>% 
   select(cbsa, median_rent2021) %>% 
@@
   mutate(cbsa = ifelse(unknown == '38860M', "38860", cbsa)) %>% 
-  mutate(median_rent2022 = as.numeric(rent_50_0)) %>% 
+  mutate(median_rent2022 = as.numeric(rent_50_1)) %>% 
   select(cbsa, median_rent2022) %>% 

### 2022-11-29 src/real_estate.Rmd
inverted income to home price column

@@
   # multiply monthly rent x 12 for annual rent!
-  mutate(avg_rent_to_income = (avg_rent_2021*12)/median_income) %>% 
   mutate(med_rent_to_income = (median_rent2021*12)/median_income)
@@
   mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
-  mutate(n_avg_rent_to_income = (avg_rent_to_income- min(avg_rent_to_income, na.rm = TRUE)) / (max(avg_rent_to_income, na.rm = TRUE) - min(avg_rent_to_income, na.rm = TRUE))) %>% 
   mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
@@
   mutate(n_med_rent_to_income_inverted = 1 - n_med_rent_to_income) %>% 
-  mutate(n_avg_rent_to_income_inverted = 1 - n_avg_rent_to_income)
+  mutate(n_income_to_homeprice_inverted = 1 - n_income_to_homeprice)
 
@@
   select(cbsa, MSA, starts_with("n_"), ) %>% 
-  mutate(real_estate_score = MSA) %>% #this is just a place holder
-  mutate(RE_score_avg_rent = MSA)
+  mutate(real_estate_score = MSA) 
 ```
@@
                   n_MSA_realEstate$n_med_rent_to_income_inverted[i], 
-                  n_MSA_realEstate$n_income_to_homeprice[i]
+                  n_MSA_realEstate$n_income_to_homeprice_inverted[i]
                   )
@@
 
-```
-
-```{r}
-
-# score using average rent
-for (i in 1:size[1]) {
-  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
-                  n_MSA_realEstate$n_employ_change[i],
-                  n_MSA_realEstate$n_avg_occupancy[i],
-                  n_MSA_realEstate$n_avg_rent_change[i],
-                  n_MSA_realEstate$n_avg_rent_to_income_inverted[i], # compare avg rent and median rent
-                  n_MSA_realEstate$n_income_to_homeprice[i]
-                  )
-  n_MSA_realEstate$RE_score_avg_rent[i] = weighted.mean(MSA_x_data, real_estate_weights)
-}
 ```

### 2022-11-29 src/demo_data_wrangling.Rmd
double checking change in emp rate

@@
 
+avg_emp_change <- MSA_employment %>% 
+  group_by(MSA, year) %>% 
+  summarize(avg_employment_rate = mean(employment_rate)) %>% 
+  pivot_wider(names_from = year,
+              values_from = avg_employment_rate,
+              names_prefix = "emp_rate") %>% 
+  mutate(emp_change20_21 = (emp_rate2021 - emp_rate2020) /emp_rate2020 *100) %>% 
+  full_join(avg_emp_yoy) %>% 
+  filter(year == "2021")
+
 MSA_emp_income <- income_potential_MSAs %>% 

### 2022-11-30 src/MSAtotal_scores.Rmd
read in real estate and landlord score data

@@
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, warning=FALSE, message=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
@@
 ```
+
+```{r}
+# create table with category scores
+# isolating score from each category data sheet
+real_estate_score <- real_estate_score_data %>% 
+  select(cbsa, real_estate_score)
+
+category_score <- landlord_score_data %>% 
+  select(MSA, cbsa, landlord_score) %>% 
+  inner_join(real_estate_score)
+
+```
+
+```{r}
+category_weights <- c(0.4, 0.25, 0.15, 0.10, 0.10)
+```
+
+```{r}
+size <- dim(category_score)
+
+for (i in 1:size[1]) {
+  MSA_x_data <- c(category_score$real_estate_score[i],
+                  category_score$financial_score[i],
+                  category_score$landlord_score[i],
+                  category_score$climate_impact_score[i], 
+                  category_score$energy_score[i]
+                  )
+  category_score$total_score[i] = weighted.mean(MSA_x_data, category_weights)
+}
+
+```

### 2022-12-05 src/anchor_city_real_estate.Rmd
calculate real estate score using anchor city occupancy rate

--- 2022-11-29 src/real_estate.Rmd
@@
 
+anchor_cities <- read_sheet("https://docs.google.com/spreadsheets/d/1leWRliw7cGGQTGIdBJzakjWZ4skQzSh04lFw_LfbtLU/edit#gid=0") %>%  clean_names()
+
 ```
@@
 # aggregate yardi data into MSAs by cbsa
-MSA_realEstate_data <- city_realEstate_data %>% 
+city_realEstate_data <- city_realEstate_data %>% 
   clean_names() %>% 
+  select(!x7) %>% 
   rename(avg_rent2022 = x2022_rent) %>% 
-   # calculate change in rent
-  mutate(rent_change = (avg_rent2022 - average_rent_2021)/average_rent_2021 * 100) %>% 
-  group_by(cbsa) %>% 
-  # calculate average values for MSAs with more than one city
-  summarize(avg_rent_2021 = mean(average_rent_2021, na.rm = TRUE),
-            avg_rent_2022 = mean(avg_rent2022, na.rm = TRUE),
-            avg_occupancy_21 = mean(percent_occupancy_2021, na.rm = TRUE),
-            avg_rent_change = mean(rent_change, na.rm = TRUE))
+  rename(occ_rate2021 = percent_occupancy_2021)
+
+anchor_cities_list <- unique(anchor_cities$city_msa)
+
+anchor_city_data <- city_realEstate_data %>% 
+  filter(city %in% c(anchor_cities_list))
 
@@
 compliedMSA_realEstate<- select_MSAs %>% 
-  left_join(MSA_realEstate_data) %>% 
+  left_join(anchor_city_data, by = "cbsa") %>% 
     # calculate rent change
@@
   mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
-  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
-  # normalize rent change
-  mutate(n_avg_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))
+  mutate(n_avg_occupancy = (occ_rate2021 - min(occ_rate2021)) / (max(occ_rate2021) - min(occ_rate2021)))
 
@@
 ```{r}
-# create downloadable data table
-datatable(n_MSA_realEstate, extensions = 'Buttons', options = list(
-    dom = 'Bfrtip',
-    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
-  )
-)
+#write result to google sheet
+#gs4_create("anchor_real_estate_score")
+sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "anchor_realEstate")
 ```
-
-```{r}
-#write result to google sheet
-# gs4_create("MSA_real_estate_score")
-sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "n_MSA_realEstate")
-```

### 2022-12-06 src/landlord.Rmd
changed landlord score calc 80: eviction notice, 20: security deposit, no entry notice

@@
 
-select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")
+# select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")
 ```
@@
 #filter by new MSA list
-new_MSA_list <- unique(select_MSAs$cbsa)
-
-MSA_landlord_data <- MSA_landlord_data %>% 
-  filter(cbsa %in% c(new_MSA_list))
+# new_MSA_list <- unique(select_MSAs$cbsa)
+# 
+# MSA_landlord_data <- MSA_landlord_data %>% 
+#   filter(cbsa %in% c(new_MSA_list))
 ```
@@
 
-#notice_of_entry_days
-MSA_landlord_data <- MSA_landlord_data %>% 
-  mutate(
-    notice_for_entry_days_num = case_when(
-      notice_for_entry_days == "N" ~ 1.0, 
-      notice_for_entry_days == "0" ~ 1.0,
-      notice_for_entry_days == "0.5" ~ 0.75,
-      notice_for_entry_days == "1" ~ 0.5,
-      notice_for_entry_days == "2" ~ 0.0,))
+# #notice_of_entry_days
+# MSA_landlord_data <- MSA_landlord_data %>% 
+#   mutate(
+#     notice_for_entry_days_num = case_when(
+#       notice_for_entry_days == "N" ~ 1.0, 
+#       notice_for_entry_days == "0" ~ 1.0,
+#       notice_for_entry_days == "0.5" ~ 0.75,
+#       notice_for_entry_days == "1" ~ 0.5,
+#       notice_for_entry_days == "2" ~ 0.0,))
 ```
@@
 #Weights must sum to 1
-landlord_weights <- c(0.9, 0.05, 0.05)
+landlord_weights <- c(0.8, 0.2)
 
@@
   MSA_x_data <- c(MSA_landlord_data$eviction_notice_days_normalized_inverted[i],
-                  MSA_landlord_data$security_depopsit_limit_num[i],
-                  MSA_landlord_data$notice_for_entry_days_num[i]
+                  MSA_landlord_data$security_depopsit_limit_num[i]
                   )
@@
 ```
+
+```{r}
+#write result to google sheet
+gs4_create("MSA_landlord_score")
+sheet_write(MSA_landlord_data, ss = "https://docs.google.com/spreadsheets/d/1nLpDhM88a3rEBDVK5RZlk2B6QuwVIwvsfE5HeBz0mLo/edit?usp=sharing", sheet = "MSA_landlord_data")
+```

### 2022-12-06 src/landlord.Rmd
don't make new sheet to write in

@@
 #write result to google sheet
-gs4_create("MSA_landlord_score")
+#gs4_create("MSA_landlord_score")
 sheet_write(MSA_landlord_data, ss = "https://docs.google.com/spreadsheets/d/1nLpDhM88a3rEBDVK5RZlk2B6QuwVIwvsfE5HeBz0mLo/edit?usp=sharing", sheet = "MSA_landlord_data")

### 2023-01-17 src/reopt.Rmd
read in reopt data

---
title: "reOpt Data"
author: "Grace Bianchi"
date: "2023-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(googlesheets4)
```

```{r}
# read in data
reopt_data <- read_sheet("https://docs.google.com/spreadsheets/d/1VROhOTWY4jSeI_zNgFui4ie6QiIz7uzzli8rF72fAyg/edit#gid=0") %>% 
  clean_names()

```

```{r}
Annual Renewable Electricity (%)
Lifecycle Total tons CO2
Internal Rate of Return (IRR) (%)
Lifecycle Costs of Health Emissions ($)

```

### 2023-01-18 src/reopt.Rmd
normalized annual energy production metric

@@
 reopt_data <- read_sheet("https://docs.google.com/spreadsheets/d/1VROhOTWY4jSeI_zNgFui4ie6QiIz7uzzli8rF72fAyg/edit#gid=0") %>% 
-  clean_names()
+  clean_names() %>% 
+  drop_na()
 
@@
 ```{r}
-Annual Renewable Electricity (%)
-Lifecycle Total tons CO2
-Internal Rate of Return (IRR) (%)
-Lifecycle Costs of Health Emissions ($)
+annual_renewable_electricity <- reopt_data %>% 
+  select(cbsa, city_msa, county, annual_renewable_electricity_percent_of_total) %>% 
+  rename(annual_renewable_percent = annual_renewable_electricity_percent_of_total) %>% 
+  mutate(annual_renewable_percent = annual_renewable_percent*100) %>% 
+  # normalize data 
+  mutate(normalized_annual_renewable = (annual_renewable_percent - min(annual_renewable_percent))/(max(annual_renewable_percent)-min(annual_renewable_percent)))
+
+total_Co2_tons <- reopt_data %>% 
+  select(cbsa, city_msa, county, lifecycle_total_t_co2) %>%
+
+irr_percent <- reopt_data %>% 
+   select(cbsa, city_msa, county, irr)
+
+health_emissions_costs <- reopt_data  %>% 
+   select(cbsa, city_msa, county, lifecycle_costs_of_health_emissions_difference_from_bau
 

### 2023-01-18 src/reopt.Rmd
selecting metric for each criteria

--- 2023-01-17 src/reopt.Rmd
@@
 ```{r}
-Annual Renewable Electricity (%)
-Lifecycle Total tons CO2
-Internal Rate of Return (IRR) (%)
-Lifecycle Costs of Health Emissions ($)
+energy_production_metric <- reopt_data %>% 
+  select(cbsa, city_msa, county, annual_renewable_electricity_percent_of_total) %>% 
+  rename(annual_renewable_percent = annual_renewable_electricity_percent_of_total) %>% 
+  mutate(annual_renewable_percent = annual_renewable_percent*100) %>% 
+  # normalize data 
+  mutate(normalized_annual_renewable = (annual_renewable_percent - min(annual_renewable_percent))/(max(annual_renewable_percent)-min(annual_renewable_percent)))
+
+climate_impacts_metric <- reopt_data %>% 
+  select(cbsa, city_msa, county, lifecycle_total_t_co2) %>%
+
+financial_metric <- reopt_data %>% 
+   select(cbsa, city_msa, county, irr)
+
+health_emissions_costs <- reopt_data  %>% 
+   select(cbsa, city_msa, county, lifecycle_costs_of_health_emissions_difference_from_bau
 

### 2023-01-18 src/reopt.Rmd
normalized health impact metric

--- 2023-01-17 src/reopt.Rmd
@@
 ```{r}
-Annual Renewable Electricity (%)
-Lifecycle Total tons CO2
-Internal Rate of Return (IRR) (%)
-Lifecycle Costs of Health Emissions ($)
+energy_production_metric <- reopt_data %>% 
+  select(cbsa, city_msa, county, annual_renewable_electricity_percent_of_total) %>% 
+  rename(annual_renewable_percent = annual_renewable_electricity_percent_of_total) %>% 
+  # normalize data 
+  mutate(normalized_annual_renewable = (annual_renewable_percent - min(annual_renewable_percent))/(max(annual_renewable_percent)-min(annual_renewable_percent)))
+
+climate_impacts_metric <- reopt_data %>% 
+  select(cbsa, city_msa, county, lifecycle_total_t_co2) %>% 
+  # normalize total co2 
+  mutate(normalized_total_co2 = (lifecycle_total_t_co2 - min(lifecycle_total_t_co2)) / (max(lifecycle_total_t_co2) - min(lifecycle_total_t_co2)))
+
+financial_metric <- reopt_data %>% 
+   select(cbsa, city_msa, county, irr) %>% 
+  # normalize IRR
+  mutate(normalized_irr = (irr - min(irr)) / (max(irr) - min(irr)))
+
+health_impact_metric <- reopt_data  %>% 
+   select(cbsa, city_msa, county, lifecycle_costs_of_health_emissions_difference_from_bau) %>% 
+  rename(health_emissions_cost = lifecycle_costs_of_health_emissions_difference_from_bau) %>% 
+  mutate(normalized_health_emission_cost = (health_emissions_cost - min(health_emissions_cost)) / (max(health_emissions_cost) - min(health_emissions_cost)))
 

### 2023-01-18 src/reopt.Rmd
combined metrics into one sheet

--- 2023-01-18 src/reopt.Rmd
@@
 ```{r}
-annual_renewable_electricity <- reopt_data %>% 
-  select(cbsa, city_msa, county, annual_renewable_electricity_percent_of_total) %>% 
+energy_production_metric <- reopt_data %>% 
+  select(cbsa, city_msa, state, annual_renewable_electricity_percent_of_total) %>% 
   rename(annual_renewable_percent = annual_renewable_electricity_percent_of_total) %>% 
-  mutate(annual_renewable_percent = annual_renewable_percent*100) %>% 
   # normalize data 
-  mutate(normalized_annual_renewable = (annual_renewable_percent - min(annual_renewable_percent))/(max(annual_renewable_percent)-min(annual_renewable_percent)))
+  mutate(normalized_annual_renewable = (annual_renewable_percent - min(annual_renewable_percent))/(max(annual_renewable_percent)-min(annual_renewable_percent))) %>% 
+  select(cbsa, city_msa, state, normalized_annual_renewable)
 
-total_Co2_tons <- reopt_data %>% 
-  select(cbsa, city_msa, county, lifecycle_total_t_co2) %>%
+climate_impacts_metric <- reopt_data %>% 
+  # normalize total co2 
+  mutate(normalized_total_co2 = (lifecycle_total_t_co2 - min(lifecycle_total_t_co2)) / (max(lifecycle_total_t_co2) - min(lifecycle_total_t_co2))) %>% 
+  # invert co2
+  mutate(inverted_n_total_co2 = 1 - normalized_total_co2) %>% 
+  select(cbsa, inverted_n_total_co2)
 
-irr_percent <- reopt_data %>% 
-   select(cbsa, city_msa, county, irr)
+financial_metric <- reopt_data %>% 
+  # normalize IRR
+  mutate(normalized_irr = (irr - min(irr)) / (max(irr) - min(irr))) %>% 
+  select(cbsa, normalized_irr)
 
-health_emissions_costs <- reopt_data  %>% 
-   select(cbsa, city_msa, county, lifecycle_costs_of_health_emissions_difference_from_bau
+health_impact_metric <- reopt_data  %>% 
+  rename(health_emissions_cost = lifecycle_costs_of_health_emissions_difference_from_bau) %>% 
+  mutate(normalized_health_emission_cost = (health_emissions_cost - min(health_emissions_cost)) / (max(health_emissions_cost) - min(health_emissions_cost))) %>% 
+  # invert 
+  mutate(inverted_n_health_emissions_cost = 1 - normalized_health_emission_cost) %>% 
+   select(cbsa, inverted_n_health_emissions_cost) 
+```
 
+```{r}
+# reopt normalized metrics
+reopt_metrics_normalized <- energy_production_metric %>% 
+  left_join(climate_impacts_metric) %>% 
+  left_join(financial_metric) %>% 
+  left_join(health_impact_metric)
 ```

### 2023-01-18 src/climate_risk.Rmd
new NRI climate data and climate risk rmd

---
title: "Climate Risk"
author: "Julia Bickford"
date: '2023-01-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
```
```{r}
#read in National Risk Index (NRI) county data

climate_risk <- read.csv(here("data", "raw", "NRI_Table_Counties.csv")) %>% 
  clean_names()

#filter data by county

```

### 2023-01-19 src/climate_risk.Rmd

@@
 knitr::opts_chunk$set(echo = TRUE)
-library(here)
+
 library(tidyverse)
 library(janitor)
+library(googlesheets4)
 ```
+
 ```{r}
+# read in reopt data (for STCOFIPS codes)
+reopt_data <- read_sheet("https://docs.google.com/spreadsheets/d/1VROhOTWY4jSeI_zNgFui4ie6QiIz7uzzli8rF72fAyg/edit#gid=0") %>% 
+  clean_names() %>% 
+  drop_na()
+
 #read in National Risk Index (NRI) county data
 
-climate_risk <- read.csv(here("data", "raw", "NRI_Table_Counties.csv")) %>% 
-  clean_names()
-
-#filter data by county
+climate_risk <- read_sheet("https://docs.google.com/spreadsheets/d/1J8-PrtmSASmcbLp7Z2TDYsqx1_sSM6naNYPzyMeElGM/edit?usp=sharing")
 

### 2023-01-20 src/climate_risk.Rmd
NRI scores calculated with stcofips id and cbsa number

@@
 ```
+
+```{r}
+#pull stcofips numbers for anchor cities
+anchor_stcofips <- reopt_data %>% pull(stcofips_5_digit)
+```
+
+```{r}
+#filter NRI data to only include anchor cities
+anchor_climate_risk <- climate_risk %>% filter(STCOFIPS %in% anchor_stcofips)
+```
+
+```{r}
+#select stcopfips number and overall risk score from NRI data 
+anchor_climate_risk_score <- anchor_climate_risk %>% select(STCOFIPS, RISK_SCORE)
+```
+
+```{r}
+#select stcofips number and cbsa number from from reopt anchor city data
+reopt_data_cbsa_stcofips <- reopt_data %>% select(cbsa, stcofips_5_digit)
+
+#rename stcofips col to merge data tables by stcofips number
+reopt_data_cbsa_stcofips <- reopt_data_cbsa_stcofips %>% rename("STCOFIPS" =  "stcofips_5_digit")
+```
+
+```{r}
+#merge data to create dable with: CBSA, stcofips, and risk score
+climate_risk_cbsa_stcofips <- merge(reopt_data_cbsa_stcofips, anchor_climate_risk_score, by="STCOFIPS")
+```
+
+```{r}
+#normalize climate risk scores
+climate_risk_cbsa_stcofips <- climate_risk_cbsa_stcofips %>% 
+  # normalize climate risk
+  mutate(normalized_climate_risk = (RISK_SCORE - min(RISK_SCORE)) / (max(RISK_SCORE) - min(RISK_SCORE)))
+
+#invert normalized scores so that places with less climate risk have a better score
+climate_risk_cbsa_stcofips <- climate_risk_cbsa_stcofips %>% 
+  # normalize climate risk
+  mutate(normalized_inverted_climate_risk = 1-normalized_climate_risk)
+
+```
+
+```{r}
+#write result to google sheet
+sheet_write(climate_risk_cbsa_stcofips, ss = "https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit?usp=sharing")
+```

### 2023-01-22 src/climate_risk.Rmd
sensitivity analysis data (mean, 95th percentile and distribution) in reopt.Rmd and climate_risk.Rmd

@@
 library(googlesheets4)
+library(ggplot2)
 ```
@@
 ```{r}
+#mean ("mean_metric"), 95th percentile ("perc_metric"), and distribution for sensitivity analysis
+
+mean_climate_risk <- mean(climate_risk_cbsa_stcofips$normalized_inverted_climate_risk)
+mean_climate_risk
+
+perc_climate_risk <- quantile(climate_risk_cbsa_stcofips$normalized_inverted_climate_risk, probs = 0.95)
+perc_climate_risk
+
+climate_risk_distribution <- density(climate_risk_cbsa_stcofips$normalized_inverted_climate_risk)
+plot(climate_risk_distribution)
+```
+
+```{r}
 #write result to google sheet

### 2023-01-22 src/reopt.Rmd
output normalized metrics to google sheets
sensitivity analysis data (mean, 95th percentile and distribution) in reopt.Rmd and climate_risk.Rmd

@@
 ```
+
+```{r}
+sheet_write(reopt_metrics_normalized, ss = "https://docs.google.com/spreadsheets/d/1VROhOTWY4jSeI_zNgFui4ie6QiIz7uzzli8rF72fAyg/edit#gid=498103540", sheet = "normalized_reopt_metrics")
+```
+
+```{r}
+# mean ("mean_metric"), 95th percentile ("perc_metric"), and distribution for sensitivity analysis
+
+#energy production 
+mean_energy_production <- mean(energy_production_metric$normalized_annual_renewable)
+mean_energy_production
+
+perc_energy_production <- quantile(energy_production_metric$normalized_annual_renewable, probs = 0.95)
+perc_energy_production
+
+energy_distribution <- density(energy_production_metric$normalized_annual_renewable)
+plot(energy_distribution)
+
+#climate impacts
+mean_total_co2 <- mean(climate_impacts_metric$inverted_n_total_co2)
+mean_total_co2
+
+perc_total_co2 <- quantile(climate_impacts_metric$inverted_n_total_co2, probs = 0.95)
+perc_total_co2
+
+climate_impacts_distribution <- density(climate_impacts_metric$inverted_n_total_co2)
+plot(climate_impacts_distribution)
+
+#financial
+mean_irr <- mean(financial_metric$normalized_irr)
+mean_irr
+
+perc_irr <- quantile(financial_metric$normalized_irr, probs = 0.95)
+perc_irr
+
+irr_distribution <- density(financial_metric$normalized_irr)
+plot(irr_distribution)
+
+#health impacts
+mean_health_emissions <- mean(health_impact_metric$inverted_n_health_emissions_cost)
+mean_health_emissions
+
+perc_health <- quantile(health_impact_metric$inverted_n_health_emissions_cost, probs = 0.95)
+perc_health
+
+health_distribution <- density(health_impact_metric$inverted_n_health_emissions_cost)
+plot(health_distribution)
+```

### 2023-01-23 src/MSAtotal_scores.Rmd
calculated total scores using weights from Owen and student weights

--- 2022-11-14 src/MSAtotal_scores.Rmd
@@
 # read in the data
-category_score_data <- read_sheet(https://docs.google.com/spreadsheets/d/1zu6ta_gI5lF3l2RC5eE3BuUsywYg8gd6xcMfuTKb9Bk/edit#gid=852316062)
+real_estate_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit#gid=1038418521") %>% 
+  select(cbsa, real_estate_score)
+
+rental_policy_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1nLpDhM88a3rEBDVK5RZlk2B6QuwVIwvsfE5HeBz0mLo/edit#gid=983090810") %>% 
+  select(cbsa, landlord_score)
+
+reopt_data <- read_sheet("https://docs.google.com/spreadsheets/d/1tPqpmS8JOeRyunEmSyFYPYwshKDGx54Z-MISgYO3wZE/edit#gid=1874679054") %>% 
+  select(cbsa, city_msa, state, normalized_irr, inverted_n_total_co2, inverted_n_health_emissions_cost, normalized_annual_renewable)
+
+climate_risk_data <- read_sheet("https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit#gid=2045475301") %>% 
+  select(cbsa, normalized_inverted_climate_risk)
 
 ```
+
+```{r}
+criteria_all <- real_estate_score_data %>% 
+  inner_join(rental_policy_score_data)  %>% 
+  inner_join(reopt_data) %>% 
+  inner_join(climate_risk_data)
+```
+
+```{r}
+# weight criteria for Owen
+criteria_weights_owen <- c(0.37, 0.12, 0.22, 0.07, 0.0, 0.07, 0.15)
+
+#make landlord_score column
+criteria_all <- criteria_all %>% 
+  mutate(total_score_owen = cbsa, #this is just a place holder
+         real_estate_score = as.numeric(real_estate_score),
+         landlord_score = as.numeric(landlord_score)
+  )
+
+size <- dim(criteria_all)
+for (i in 1:size[1]) {
+  cbsa_data <- c(criteria_all$real_estate_score[i],
+                criteria_all$landlord_score[i],
+                 criteria_all$normalized_irr[i],
+                 criteria_all$inverted_n_total_co2[i],
+                criteria_all$inverted_n_health_emissions_cost[i],
+                criteria_all$normalized_annual_renewable[i],
+                criteria_all$normalized_inverted_climate_risk[i]
+                  )
+  criteria_all$total_score_owen[i] = weighted.mean(cbsa_data, criteria_weights_owen)
+}
+
+```
+
+```{r}
+# weight criteria for students
+criteria_weights_students <- c(0.21, 0.0, 0.12, 0.17, 0.16, 0.17, 0.17)
+
+#make landlord_score column
+criteria_all <- criteria_all %>% 
+  mutate(total_score_students = cbsa, #this is just a place holder
+         real_estate_score = as.numeric(real_estate_score),
+         landlord_score = as.numeric(landlord_score)
+  )
+
+size <- dim(criteria_all)
+for (i in 1:size[1]) {
+  cbsa_data <- c(criteria_all$real_estate_score[i],
+                criteria_all$landlord_score[i],
+                 criteria_all$normalized_irr[i],
+                 criteria_all$inverted_n_total_co2[i],
+                criteria_all$inverted_n_health_emissions_cost[i],
+                criteria_all$normalized_annual_renewable[i],
+                criteria_all$normalized_inverted_climate_risk[i]
+                  )
+  criteria_all$total_score_students[i] = weighted.mean(cbsa_data, criteria_weights_students)
+}
+
+```
+
+```{r}
+sheet_write(criteria_all, ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0")
+```

### 2023-01-23 src/landlord.Rmd

@@
 library(DT)
+library(ggplot2)
 
@@
 ```
+
+```{r}
+# mean ("mean_metric"), 95th percentile ("perc_metric"), and distribution for sensitivity analysis
+
+#Eviction Notice (normalized, inverted)
+mean_eviction <- mean(MSA_landlord_data$eviction_notice_days_normalized_inverted)
+mean_eviction
+
+perc_eviction <- quantile(MSA_landlord_data$eviction_notice_days_normalized_inverted, probs = 0.95)
+perc_eviction
+
+eviction_distribution <- density(MSA_landlord_data$eviction_notice_days_normalized_inverted)
+plot(eviction_distribution)
+
+#Security Deposit Limit
+mean_deposit <- mean(MSA_landlord_data$security_depopsit_limit_num)
+mean_deposit
+
+perc_deposit <- quantile(MSA_landlord_data$security_depopsit_limit_num, probs = 0.95)
+perc_deposit
+
+deposit_distribution <- density(MSA_landlord_data$security_depopsit_limit_num)
+plot(deposit_distribution)
+
+```

### 2023-01-27 src/data_viz.Rmd
read in all data

---
title: "Data Visualization"
author: "Grace Bianchi"
date: "2023-01-27"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(googlesheets4)
library(tidyverse)
library(janitor)

```

```{r}
total_scores_data <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202")

real_estate_data <- read_sheet("https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit#gid=1038418521")

rental_policy_data <- read_sheet("https://docs.google.com/spreadsheets/d/1nLpDhM88a3rEBDVK5RZlk2B6QuwVIwvsfE5HeBz0mLo/edit#gid=983090810")

climate_risk_data <- read_sheet("https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit")
```

### 2023-01-27 src/data_viz.Rmd
started bar chart

@@
 library(janitor)
+library(ggplot2)
 
@@
 ```{r}
-total_scores_data <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202")
+total_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202") %>% 
+  rename(financial_score = normalized_irr,
+         climate_impact = inverted_n_total_co2,
+         health_impact = inverted_n_health_emissions_cost,
+         energy_production = normalized_annual_renewable,
+         climate_risk = normalized_inverted_climate_risk)
 
@@
 ```
+
+```{r}
+score_tidy <- total_score_data %>%
+
+```
+
+```{r}
+
+ggplot(data = total_score_data, aes(y = city_msa, x = total_score_owen)) + 
+    geom_bar(position="stack", stat="identity")
+```

### 2023-01-30 src/MSAtotal_scores.Rmd
tested different criteria weights, weighting climate risk as the highest

@@
 
-```{r}
+```{r original weights}
 # weight criteria for Owen
@@
 
-```{r}
+```{r student weights}
 # weight criteria for students
@@
 
+```{r climate risk weighted 50 and other 0.1}
+# weight criteria for Owen
+weights_climate50 <- c(0.1, 0.1, 0.1, 0.1, 0.0, 0.1, 0.50)
+
+#make landlord_score column
+criteria_all <- criteria_all %>% 
+  mutate(score_climate50 = cbsa, #this is just a place holder
+         real_estate_score = as.numeric(real_estate_score),
+         landlord_score = as.numeric(landlord_score)
+  )
+
+size <- dim(criteria_all)
+for (i in 1:size[1]) {
+  cbsa_data <- c(criteria_all$real_estate_score[i],
+                criteria_all$landlord_score[i],
+                 criteria_all$normalized_irr[i],
+                 criteria_all$inverted_n_total_co2[i],
+                criteria_all$inverted_n_health_emissions_cost[i],
+                criteria_all$normalized_annual_renewable[i],
+                criteria_all$normalized_inverted_climate_risk[i]
+                  )
+  criteria_all$score_climate50[i] = weighted.mean(cbsa_data, weights_climate50)
+}
+
+```
+
+```{r climate risk 40 }
+# weight criteria for Owen
+weights_climate40 <- c(0.15, 0.1, 0.15, 0.1, 0.0, 0.1, 0.40)
+
+#make landlord_score column
+criteria_all <- criteria_all %>% 
+  mutate(score_climate40 = cbsa, #this is just a place holder
+         real_estate_score = as.numeric(real_estate_score),
+         landlord_score = as.numeric(landlord_score)
+  )
+
+size <- dim(criteria_all)
+for (i in 1:size[1]) {
+  cbsa_data <- c(criteria_all$real_estate_score[i],
+                criteria_all$landlord_score[i],
+                 criteria_all$normalized_irr[i],
+                 criteria_all$inverted_n_total_co2[i],
+                criteria_all$inverted_n_health_emissions_cost[i],
+                criteria_all$normalized_annual_renewable[i],
+                criteria_all$normalized_inverted_climate_risk[i]
+                  )
+  criteria_all$score_climate40[i] = weighted.mean(cbsa_data, weights_climate40)
+}
+```
+
+```{r climate risk 50 and proportional weights for others}
+weights_climate50_prop <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)
+
+#make landlord_score column
+criteria_all <- criteria_all %>% 
+  mutate(score_climate50_prop = cbsa, #this is just a place holder
+         real_estate_score = as.numeric(real_estate_score),
+         landlord_score = as.numeric(landlord_score)
+  )
+
+size <- dim(criteria_all)
+for (i in 1:size[1]) {
+  cbsa_data <- c(criteria_all$real_estate_score[i],
+                criteria_all$landlord_score[i],
+                 criteria_all$normalized_irr[i],
+                 criteria_all$inverted_n_total_co2[i],
+                criteria_all$inverted_n_health_emissions_cost[i],
+                criteria_all$normalized_annual_renewable[i],
+                criteria_all$normalized_inverted_climate_risk[i]
+                  )
+  criteria_all$score_climate50_prop[i] = weighted.mean(cbsa_data, weights_climate50_prop)
+}
+```
+
 ```{r}

### 2023-01-30 src/data_viz.Rmd
started bar chart

--- 2023-01-27 src/data_viz.Rmd
@@
 ```{r}
-total_scores_data <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202")
+total_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202") %>% 
+  rename(financial_score = normalized_irr,
+         climate_impact = inverted_n_total_co2,
+         health_impact = inverted_n_health_emissions_cost,
+         electricity = normalized_annual_renewable,
+         climate_risk = normalized_inverted_climate_risk)
 
@@
 ```
+
+```{r}
+score_tidy <- total_score_data %>% 
+  pivot_longer(cols = c(real_estate_score, landlord_score, financial_score:climate_risk),
+               names_to = "criteria",
+               values_to = "criteria_score") %>% 
+  filter(city_msa != "Portland") %>% 
+  group_by(city_msa)
+
+```
+
+```{r}
+ggplot(data = score_tidy, aes(y = city_msa, x = total_score_owen, fill = criteria)) + 
+    geom_bar()
+
+ggplot(data = score_tidy, aes(y = city_msa, x = total_score_owen, fill = criteria)) + 
+    geom_col(position="stacked")
+```

### 2023-01-31 src/MSAtotal_scores.Rmd
renaming columns

@@
 ```{r climate risk 50 and proportional weights for others}
+
 weights_climate50_prop <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)
@@
 criteria_all <- criteria_all %>% 
-  mutate(score_climate50_prop = cbsa, #this is just a place holder
-         real_estate_score = as.numeric(real_estate_score),
-         landlord_score = as.numeric(landlord_score)
-  )
+  mutate(score_climate50_prop = cbsa)#this is just a place holder
 
@@
 
-```{r}
+```{r rename columns and output to google sheet}
+criteria_all <- criteria_all %>% 
+  rename(financial_score = normalized_irr,
+         climate_impact = inverted_n_total_co2,
+         health_impact = inverted_n_health_emissions_cost,
+         electricity = normalized_annual_renewable,
+         climate_risk = normalized_inverted_climate_risk)
+
 sheet_write(criteria_all, ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0")
 ```
+
+```{r}
+column_order <- c("real_estate_weighted", "rental_weighted", "financial_weighted", "electricity_weighted", "climate_impact_weighted", "climate_risk_weighted", "health_weighted", "cbsa", "city_msa", "state")
+
+weighted_criteria <- criteria_all %>% 
+  rename(real_estate_weighted = real_estate_score,
+        rental_weighted = landlord_score,
+        financial_weighted = normalized_irr,
+        electricity_weighted = normalized_annual_renewable,
+        climate_impact_weighted = inverted_n_total_co2,
+        climate_risk_weighted = normalized_inverted_climate_risk,
+        health_weighted = inverted_n_health_emissions_cost)
+
+weighted_criteria <- weighted_criteria[, column_order]
+
+```
+
+```{r data viz}
+
+score_tidy <- criteria_all %>% 
+  pivot_longer(cols = c(real_estate_score, landlord_score, financial_score:climate_risk),
+               names_to = "criteria",
+               values_to = "criteria_score") %>% 
+  filter(city_msa != "Portland") %>% 
+  group_by(city_msa)
+
+```

### 2023-01-31 src/MSAtotal_scores.Rmd
added owen's emailed weight calculations Jan 31.

@@
 
+```{r climate risk 35 and Jan 31 weights from owen}
+
+weights_climate35_owen <- c(0.25, 0.1, 0.15, 0.075, 0.0, 0.075, 0.35)
+
+#make landlord_score column
+criteria_all <- criteria_all %>% 
+  mutate(score_climate35_owen = cbsa)#this is just a place holder
+
+size <- dim(criteria_all)
+for (i in 1:size[1]) {
+  cbsa_data <- c(criteria_all$real_estate_score[i],
+                criteria_all$landlord_score[i],
+                 criteria_all$normalized_irr[i],
+                 criteria_all$inverted_n_total_co2[i],
+                criteria_all$inverted_n_health_emissions_cost[i],
+                criteria_all$normalized_annual_renewable[i],
+                criteria_all$normalized_inverted_climate_risk[i]
+                  )
+  criteria_all$score_climate35_owen[i] = weighted.mean(cbsa_data, weights_climate35_owen) 
+}
+```
+
 ```{r rename columns and output to google sheet}

### 2023-01-31 src/anchor_city_real_estate.Rmd
calculated weighted sum

@@
 library(DT)
+library(MCDA)
 
@@
 
-anchor_cities_list <- unique(anchor_cities$city_msa)
+anchor_cities_list <- unique(anchor_cities$anchor_city)
 
@@
   mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
-  mutate(n_avg_occupancy = (occ_rate2021 - min(occ_rate2021)) / (max(occ_rate2021) - min(occ_rate2021)))
+  mutate(n_avg_occupancy = (occ_rate2021 - min(occ_rate2021)) / (max(occ_rate2021) - min(occ_rate2021)), na.rm = FALSE)
 
@@
                   )
-  n_MSA_realEstate$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
+  n_MSA_realEstate$real_estate_score[i] = sum(MSA_x_data * real_estate_weights)
 }
+
+```
+
+```{r}
+
+weighted_table <-n_MSA_realEstate %>% 
+  mutate(weighted_pop_change = 0.35*n_pop_change,
+                  weighted_emp_change = 0.20*n_employ_change,
+                  weighted_occ = 0.15 * n_avg_occupancy,
+                  weighted_rent_change = 0.10*n_med_rent_change,
+                  weighted_rent_income = 0.10 * n_med_rent_to_income_inverted, 
+                  weighted_income_homeprice = 0.10 * n_income_to_homeprice_inverted) %>% 
+  mutate(real_estate_score = weighted_pop_change + weighted_emp_change + weighted_occ + weighted_rent_change + weighted_rent_income + weighted_income_homeprice)
 

### 2023-02-02 src/MSAtotal_scores.Rmd
rounded final scores and reordered columns

@@
 ```{r setup, include=FALSE, warning=FALSE, message=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
 
@@
 
-```{r}
+```{r read in data}
 criteria_all <- real_estate_score_data %>% 
@@
   inner_join(reopt_data) %>% 
-  inner_join(climate_risk_data)
+  inner_join(climate_risk_data) # combine all criteria scores into one dataframe
+
+# reorder columns
+col_order <-c("cbsa", "city_msa", "state", "real_estate_score", "landlord_score", "normalized_annual_renewable", "normalized_irr", "inverted_n_total_co2", "normalized_inverted_climate_risk", "inverted_n_health_emissions_cost")
+
+criteria_all <- criteria_all[, col_order]
+
 ```
@@
 criteria_all <- criteria_all %>% 
-  mutate(total_score_students = cbsa, #this is just a place holder
-         real_estate_score = as.numeric(real_estate_score),
-         landlord_score = as.numeric(landlord_score)
-  )
+  mutate(total_score_students = cbsa) #this is just a place holder
 
@@
 criteria_all <- criteria_all %>% 
-  mutate(score_climate50 = cbsa, #this is just a place holder
-         real_estate_score = as.numeric(real_estate_score),
-         landlord_score = as.numeric(landlord_score)
-  )
+  mutate(score_climate50 = cbsa) #this is just a place holder
 
@@
 criteria_all <- criteria_all %>% 
-  mutate(score_climate40 = cbsa, #this is just a place holder
-         real_estate_score = as.numeric(real_estate_score),
-         landlord_score = as.numeric(landlord_score)
-  )
+  mutate(score_climate40 = cbsa) #this is just a place holder
 
@@
 ```{r rename columns and output to google sheet}
+
 criteria_all <- criteria_all %>% 
@@
          electricity = normalized_annual_renewable,
-         climate_risk = normalized_inverted_climate_risk)
+         climate_risk = normalized_inverted_climate_risk) %>% 
+  mutate_if(is.numeric, ~round(., 5))
 
-sheet_write(criteria_all, ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0")
-```
-
-```{r}
-column_order <- c("real_estate_weighted", "rental_weighted", "financial_weighted", "electricity_weighted", "climate_impact_weighted", "climate_risk_weighted", "health_weighted", "cbsa", "city_msa", "state")
-
-weighted_criteria <- criteria_all %>% 
-  rename(real_estate_weighted = real_estate_score,
-        rental_weighted = landlord_score,
-        financial_weighted = normalized_irr,
-        electricity_weighted = normalized_annual_renewable,
-        climate_impact_weighted = inverted_n_total_co2,
-        climate_risk_weighted = normalized_inverted_climate_risk,
-        health_weighted = inverted_n_health_emissions_cost)
-
-weighted_criteria <- weighted_criteria[, column_order]
-
+sheet_write(criteria_all, ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "criteria_all")
 ```
@@
 
+ggplot(data = score_tidy, aes(y = city_msa, x = total_score_owen, fill = criteria)) + 
+    geom_col()
+
 ```

### 2023-02-02 src/climate_risk.Rmd
Read in EAL Score from NRI data; normalized and inverted EAL score

--- 2023-01-19 src/climate_risk.Rmd
@@
 ```
+
+```{r}
+#pull stcofips numbers for anchor cities
+anchor_stcofips <- reopt_data %>% pull(stcofips_5_digit)
+```
+
+```{r}
+#filter NRI data to only include anchor cities
+anchor_climate_risk <- climate_risk %>% filter(STCOFIPS %in% anchor_stcofips)
+```
+
+```{r}
+#select stcopfips number and overall risk score and EAL score from NRI data 
+climate_risk_score <- anchor_climate_risk %>% select(STCOFIPS, RISK_SCORE, EAL_SCORE)
+
+```
+
+```{r}
+#select stcofips number and cbsa number from from reopt anchor city data
+reopt_data_cbsa_stcofips <- reopt_data %>% select(cbsa, stcofips_5_digit)
+
+#rename stcofips col to merge data tables by stcofips number
+reopt_data_cbsa_stcofips <- reopt_data_cbsa_stcofips %>% rename("STCOFIPS" =  "stcofips_5_digit")
+```
+
+```{r}
+#merge data to create dable with: CBSA, stcofips, and risk score
+climate_risk_cbsa_stcofips <- merge(reopt_data_cbsa_stcofips, climate_risk_score, by="STCOFIPS")
+```
+
+```{r}
+#normalize climate risk scores
+climate_risk_EAL <- climate_risk_cbsa_stcofips %>% 
+  # normalize climate risk
+  mutate(normalized_climate_risk = (RISK_SCORE - min(RISK_SCORE)) / (max(RISK_SCORE) - min(RISK_SCORE)),
+         normalized_eal = (EAL_SCORE - min(EAL_SCORE)) / (max(EAL_SCORE) - min(EAL_SCORE)), #normalize EAL score
+         #invert normalized scores so that places with less climate risk have a better score
+        normalized_inverted_climate_risk = 1-normalized_climate_risk,
+        inverted_n_eal_score = 1 - normalized_eal)
+
+```
+
+```{r}
+#mean ("mean_metric"), 95th percentile ("perc_metric"), and distribution for sensitivity analysis
+
+mean_climate_risk <- mean(climate_risk_EAL$normalized_inverted_climate_risk)
+mean_climate_risk
+
+perc_climate_risk <- quantile(climate_risk_EAL$normalized_inverted_climate_risk, probs = 0.95)
+perc_climate_risk
+
+climate_risk_distribution <- density(climate_risk_EAL$normalized_inverted_climate_risk)
+plot(climate_risk_distribution)
+```
+
+```{r}
+#write result to google sheet
+sheet_write(climate_risk_EAL, ss = "https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit?usp=sharing", sheet = "climate_risk+EAL")
+```

### 2023-02-02 src/MSAtotal_scores.Rmd
calculated total scores using EAL data

@@
 
-climate_risk_data <- read_sheet("https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit#gid=2045475301") %>% 
-  select(cbsa, normalized_inverted_climate_risk)
+climate_risk_data <- read_sheet("https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit#gid=2045475301", sheet = "climate_risk+EAL") %>% 
+  select(cbsa, normalized_inverted_climate_risk, inverted_n_eal_score)
 
@@
 # reorder columns
-col_order <-c("cbsa", "city_msa", "state", "real_estate_score", "landlord_score", "normalized_annual_renewable", "normalized_irr", "inverted_n_total_co2", "normalized_inverted_climate_risk", "inverted_n_health_emissions_cost")
+col_order <-c("cbsa", "city_msa", "state", "real_estate_score", "landlord_score", "normalized_annual_renewable", "normalized_irr", "inverted_n_total_co2", "normalized_inverted_climate_risk", "inverted_n_eal_score", "inverted_n_health_emissions_cost")
 
@@
          real_estate_score = as.numeric(real_estate_score),
-         landlord_score = as.numeric(landlord_score)
+         landlord_score = as.numeric(landlord_score),
+         score_og_weights_EAL = cbsa
   )
@@
   criteria_all$total_score_owen[i] = weighted.mean(cbsa_data, criteria_weights_owen)
+}
+
+##### original weights with EAL score
+
+for (i in 1:size[1]) {
+  cbsa_data <- c(criteria_all$real_estate_score[i],
+                criteria_all$landlord_score[i],
+                 criteria_all$normalized_irr[i],
+                 criteria_all$inverted_n_total_co2[i],
+                criteria_all$inverted_n_health_emissions_cost[i],
+                criteria_all$normalized_annual_renewable[i],
+                criteria_all$inverted_n_eal_score[i] ## EAL score
+                  )
+  criteria_all$score_og_weights_EAL[i] = weighted.mean(cbsa_data, criteria_weights_owen)
 }
@@
 criteria_all <- criteria_all %>% 
-  mutate(score_climate50_prop = cbsa)#this is just a place holder
+  mutate(score_climate50_prop = cbsa, #this is just a place holder
+         score_climate50_prop_EAL = cbsa)
 
@@
 }
+
+### Using climate weight 50 and proportional others with EAL
+for (i in 1:size[1]) {
+  cbsa_data <- c(criteria_all$real_estate_score[i],
+                criteria_all$landlord_score[i],
+                 criteria_all$normalized_irr[i],
+                 criteria_all$inverted_n_total_co2[i],
+                criteria_all$inverted_n_health_emissions_cost[i],
+                criteria_all$normalized_annual_renewable[i],
+                criteria_all$inverted_n_eal_score[i]
+                  )
+  criteria_all$score_climate50_prop_EAL[i] = weighted.mean(cbsa_data, weights_climate50_prop) 
+}
+
 ```
@@
          electricity = normalized_annual_renewable,
-         climate_risk = normalized_inverted_climate_risk) %>% 
+         climate_risk = normalized_inverted_climate_risk,
+         climate_EAL_risk = inverted_n_eal_score) %>% 
   mutate_if(is.numeric, ~round(., 5))
 
-sheet_write(criteria_all, ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "criteria_all")
+sheet_write(criteria_all, ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "scores+EAL_2_2_23")
 ```

### 2023-02-03 src/MSAtotal_scores.Rmd
working on barplot

@@
 score_tidy <- criteria_all %>% 
-  pivot_longer(cols = c(real_estate_score, landlord_score, financial_score:climate_risk),
+  pivot_longer(cols = c(4:18),
                names_to = "criteria",
@@
   filter(city_msa != "Portland") %>% 
-  group_by(city_msa)
+  group_by(city_msa) %>% 
+  filter(criteria %in% c("real_estate_score", "landlord_score", "financial_score", "climate_impact", "electricity", "climate_risk", "climate_EAL_risk", "health_impact", "electricity"))
 
-ggplot(data = score_tidy, aes(y = city_msa, x = total_score_owen, fill = criteria)) + 
+ggplot(data = score_tidy, aes(y = city_msa, x = criteria_score, fill = criteria)) + 
     geom_col()

### 2023-02-03 src/MSAtotal_scores.Rmd
calculated total score w finalized weights

@@
          real_estate_score = as.numeric(real_estate_score),
-         landlord_score = as.numeric(landlord_score),
-         score_og_weights_EAL = cbsa
+         landlord_score = as.numeric(landlord_score)
   )
@@
                 criteria_all$normalized_annual_renewable[i],
-                criteria_all$normalized_inverted_climate_risk[i]
-                  )
+                criteria_all$normalized_inverted_climate_risk[i])
   criteria_all$total_score_owen[i] = weighted.mean(cbsa_data, criteria_weights_owen)
-}
-
-##### original weights with EAL score
-
-for (i in 1:size[1]) {
-  cbsa_data <- c(criteria_all$real_estate_score[i],
-                criteria_all$landlord_score[i],
-                 criteria_all$normalized_irr[i],
-                 criteria_all$inverted_n_total_co2[i],
-                criteria_all$inverted_n_health_emissions_cost[i],
-                criteria_all$normalized_annual_renewable[i],
-                criteria_all$inverted_n_eal_score[i] ## EAL score
-                  )
-  criteria_all$score_og_weights_EAL[i] = weighted.mean(cbsa_data, criteria_weights_owen)
 }
@@
 # weight criteria for students
-criteria_weights_students <- c(0.21, 0.0, 0.12, 0.17, 0.16, 0.17, 0.17)
+criteria_weights_students <- c(0.18, 0.0, 0.10, 0.19, 0.17, 0.14, 0.22)
 
@@
 criteria_all <- criteria_all %>% 
-  mutate(total_score_students = cbsa) #this is just a place holder
+  mutate(total_score_students = cbsa,
+         stud_EAL_total = cbsa) #this is just a place holder
 
@@
 
-```
-
-```{r climate risk weighted 50 and other 0.1}
-# weight criteria for Owen
-weights_climate50 <- c(0.1, 0.1, 0.1, 0.1, 0.0, 0.1, 0.50)
-
-#make landlord_score column
-criteria_all <- criteria_all %>% 
-  mutate(score_climate50 = cbsa) #this is just a place holder
-
 size <- dim(criteria_all)
@@
                 criteria_all$normalized_annual_renewable[i],
-                criteria_all$normalized_inverted_climate_risk[i]
+                criteria_all$inverted_n_eal_score[i]
                   )
-  criteria_all$score_climate50[i] = weighted.mean(cbsa_data, weights_climate50)
+  criteria_all$stud_EAL_total[i] = weighted.mean(cbsa_data, criteria_weights_students)
 }
@@
 
-```{r climate risk 40 }
-# weight criteria for Owen
-weights_climate40 <- c(0.15, 0.1, 0.15, 0.1, 0.0, 0.1, 0.40)
-
-#make landlord_score column
-criteria_all <- criteria_all %>% 
-  mutate(score_climate40 = cbsa) #this is just a place holder
-
-size <- dim(criteria_all)
-for (i in 1:size[1]) {
-  cbsa_data <- c(criteria_all$real_estate_score[i],
-                criteria_all$landlord_score[i],
-                 criteria_all$normalized_irr[i],
-                 criteria_all$inverted_n_total_co2[i],
-                criteria_all$inverted_n_health_emissions_cost[i],
-                criteria_all$normalized_annual_renewable[i],
-                criteria_all$normalized_inverted_climate_risk[i]
-                  )
-  criteria_all$score_climate40[i] = weighted.mean(cbsa_data, weights_climate40)
-}
-```
-
-```{r climate risk 50 and proportional weights for others}
+```{r OWEN FINAL WEIGHTS climate risk 50 and proportional weights for others}
 
@@
 
-```{r climate risk 35 and Jan 31 weights from owen}
+```{r climate risk 35 and Jan 31 weights from owen, eval = FALSE}
 
@@
   rename(financial_score = normalized_irr,
-         climate_impact = inverted_n_total_co2,
+         co2_e = inverted_n_total_co2,
          health_impact = inverted_n_health_emissions_cost,
@@
 
-sheet_write(criteria_all, ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "scores+EAL_2_2_23")
+sheet_write(criteria_all, ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "FINAL_total_score")
 ```
 
-```{r data viz}
+```{r data viz, eval = FALSE}
 

### 2023-02-03 src/MSAtotal_scores.Rmd
renamed columns and updated student weights

@@
 
-```{r original weights}
+```{r original weights, eval = FALSE}
 # weight criteria for Owen
@@
 
-```{r student weights}
+```{r FINAL student weights}
 # weight criteria for students
@@
 criteria_all <- criteria_all %>% 
-  mutate(total_score_students = cbsa,
-         stud_EAL_total = cbsa) #this is just a place holder
+  mutate(final_tot_students = cbsa,
+         stud_EAL_total = cbsa,
+         real_estate_score = as.numeric(real_estate_score),
+         landlord_score = as.numeric(landlord_score)) #this is just a place holder
 
@@
                   )
-  criteria_all$total_score_students[i] = weighted.mean(cbsa_data, criteria_weights_students)
+  criteria_all$final_tot_students[i] = weighted.mean(cbsa_data, criteria_weights_students)
 }
@@
 criteria_all <- criteria_all %>% 
-  mutate(score_climate50_prop = cbsa, #this is just a place holder
+  mutate(owen_FINAL_score = cbsa, #this is just a place holder
          score_climate50_prop_EAL = cbsa)
@@
                   )
-  criteria_all$score_climate50_prop[i] = weighted.mean(cbsa_data, weights_climate50_prop) 
+  criteria_all$owen_FINAL_score[i] = weighted.mean(cbsa_data, weights_climate50_prop) 
 }

### 2023-02-04 src/MSAtotal_scores.Rmd
working on bar chart

@@
 
+```
+
+```{r student weights with EAL, eval = FALSE}
 size <- dim(criteria_all)
@@
 }
-
 ```
@@
 
+```
+
+```{r Climate with 50 & proportional others with EAL, eval=FALSE}
 ### Using climate weight 50 and proportional others with EAL
@@
 
-sheet_write(criteria_all, ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "FINAL_total_score")
+#sheet_write(criteria_all, ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "FINAL_total_score")
 ```
 
-```{r data viz, eval = FALSE}
+```{r data viz}
+## read in weighted criteria sheet
+weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_Owen")
 
-score_tidy <- criteria_all %>% 
-  pivot_longer(cols = c(4:18),
+score_tidy <- weighted_criteria %>% 
+  pivot_longer(cols = c(4:9), # selecting criteria cols
                names_to = "criteria",
                values_to = "criteria_score") %>% 
-  filter(city_msa != "Portland") %>% 
-  group_by(city_msa) %>% 
-  filter(criteria %in% c("real_estate_score", "landlord_score", "financial_score", "climate_impact", "electricity", "climate_risk", "climate_EAL_risk", "health_impact", "electricity"))
+ filter(city_msa != "Portland") # removing portland (negative number)
 
-ggplot(data = score_tidy, aes(y = city_msa, x = criteria_score, fill = criteria)) + 
-    geom_col()
+ggplot(data = score_tidy, aes(y = reorder(cbsa, FINAL_owen_SCORE), x = criteria_score, fill = criteria_labels)) + 
+    geom_col(aes(fill = criteria)) +
+  labs(x = "Total Score (Client)",
+       y = " ") +
+  guides(fill=guide_legend(title="Criteria (weighted)")) +
+scale_fill_discrete(labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Finances", "Real Estate", "Landlord")) +
+   scale_fill_manual(values = c("#FDB462", "gold", "palevioletred1",  "#BC80BD", "#6699CC", "#44AA99"))
 

### 2023-02-06 src/MSAtotal_scores.Rmd
rename legend names

@@
 
-ggplot(data = score_tidy, aes(y = reorder(cbsa, FINAL_owen_SCORE), x = criteria_score, fill = criteria_labels)) + 
+ggplot(data = score_tidy, aes(y = reorder(cbsa, FINAL_owen_SCORE), x = criteria_score)) + 
     geom_col(aes(fill = criteria)) +
@@
   guides(fill=guide_legend(title="Criteria (weighted)")) +
-scale_fill_discrete(labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Finances", "Real Estate", "Landlord")) +
-   scale_fill_manual(values = c("#FDB462", "gold", "palevioletred1",  "#BC80BD", "#6699CC", "#44AA99"))
+   scale_fill_manual(values = c("#FDB462", "gold", "palevioletred1",  "#BC80BD", "#6699CC", "#44AA99"),
+                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Finances", "Real Estate", "Landlord")) +
+  scale_y_discrete((labels = score_tidy$city_msa))
 

### 2023-02-06 src/MSAtotal_scores.Rmd
Changing axis labels on bar chart almost finalized!

--- 2023-02-04 src/MSAtotal_scores.Rmd
@@
 ## read in weighted criteria sheet
-weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_Owen")
+weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_Owen") %>% 
+  unite("msa_state", city_msa:state, sep = ", ")
 
+# read in cbsa data with full names
 score_tidy <- weighted_criteria %>% 
-  pivot_longer(cols = c(4:9), # selecting criteria cols
+  pivot_longer(cols = c("WT_real_estate":"WT_climRisk"), # selecting criteria cols
                names_to = "criteria",
                values_to = "criteria_score") %>% 
- filter(city_msa != "Portland") # removing portland (negative number)
+ filter(msa_state != "Portland, ME") %>%  # removing portland (negative number)
+  group_by(cbsa)
 
-ggplot(data = score_tidy, aes(y = reorder(cbsa, FINAL_owen_SCORE), x = criteria_score, fill = criteria_labels)) + 
+ggplot(data = score_tidy, aes(y = reorder(msa_state, FINAL_owen_SCORE), x = criteria_score)) + 
     geom_col(aes(fill = criteria)) +
@@
        y = " ") +
+  theme_light() +
   guides(fill=guide_legend(title="Criteria (weighted)")) +
-scale_fill_discrete(labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Finances", "Real Estate", "Landlord")) +
-   scale_fill_manual(values = c("#FDB462", "gold", "palevioletred1",  "#BC80BD", "#6699CC", "#44AA99"))
+   scale_fill_manual(values = c("palevioletred1","#FDB462",  "gold", "#44AA99", "dodgerblue4", "#BC80BD"),
+                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Finances", "Real Estate", "Landlord Policy"))
 

### 2023-02-06 src/MSAtotal_scores.Rmd
exported chart to jpeg

@@
 
+jpeg(filename = "wt_criteria_client_bar_chart")
 ggplot(data = score_tidy, aes(y = reorder(msa_state, FINAL_owen_SCORE), x = criteria_score)) + 
@@
        y = " ") +
-  theme_light() +
+  theme_bw() +
   guides(fill=guide_legend(title="Criteria (weighted)")) +
    scale_fill_manual(values = c("palevioletred1","#FDB462",  "gold", "#44AA99", "dodgerblue4", "#BC80BD"),
-                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Finances", "Real Estate", "Landlord Policy"))
-
+                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Finances", "Real Estate", "Landlord Policy")) +
+  scale_x_continuous(limits = c(0, 0.7), expand = c(0,0), breaks=seq(0, 0.7, 0.1))
+dev.off()
 ```

### 2023-02-06 src/MSAtotal_scores.Rmd
export tiff

--- 2023-02-06 src/MSAtotal_scores.Rmd
@@
 
+tiff(filename = "wt_criteria_client_chart.tiff",  width = 6, height = 4, units = "in", res = 200)
 ggplot(data = score_tidy, aes(y = reorder(msa_state, FINAL_owen_SCORE), x = criteria_score)) + 
@@
        y = " ") +
-  theme_light() +
+  theme_bw() +
   guides(fill=guide_legend(title="Criteria (weighted)")) +
    scale_fill_manual(values = c("palevioletred1","#FDB462",  "gold", "#44AA99", "dodgerblue4", "#BC80BD"),
-                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Finances", "Real Estate", "Landlord Policy"))
-
+                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Solar IRR", "Real Estate", "Landlord Policy")) +
+  scale_x_continuous(limits = c(0, 0.7), expand = c(0,0), breaks=seq(0, 0.7, 0.1))
+dev.off()
 ```

### 2023-02-08 src/MSAtotal_scores.Rmd
finalized bar chart

--- 2023-02-06 src/MSAtotal_scores.Rmd
@@
 
-```{r data viz}
+```{r client data viz}
 ## read in weighted criteria sheet
-weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_Owen") %>% 
+client_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_Owen") %>% 
   unite("msa_state", city_msa:state, sep = ", ")
@@
 # read in cbsa data with full names
-score_tidy <- weighted_criteria %>% 
+score_tidy <- client_weighted_criteria %>% 
   pivot_longer(cols = c("WT_real_estate":"WT_climRisk"), # selecting criteria cols
@@
 
+tiff(filename = "wt_criteria_client_chart.tiff", width = 8, height = 6, units = "in", res = 300)
 ggplot(data = score_tidy, aes(y = reorder(msa_state, FINAL_owen_SCORE), x = criteria_score)) +
     geom_col(aes(fill = criteria)) +
-  labs(x = "Total Score (Client)",
-       y = " ") +
-  theme_light() +
+  labs(x = "Total Score",
+       y = " ",
+       title = "Client") +
+  theme_bw() +
   guides(fill=guide_legend(title="Criteria (weighted)")) +
    scale_fill_manual(values = c("palevioletred1","#FDB462",  "gold", "#44AA99", "dodgerblue4", "#BC80BD"),
-                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Finances", "Real Estate", "Landlord Policy"))
+                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Solar IRR", "Real Estate", "Landlord Policy")) +
+  scale_x_continuous(limits = c(0, 0.7), expand = c(0,0), breaks=seq(0, 0.7, 0.1))
+dev.off()
+```
 
+```{r}
+## read in weighted criteria sheet
+student_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_stud") %>% 
+  unite("msa_state", city_msa:state, sep = ", ")
+
+# read in cbsa data with full names
+score_tidy2 <- student_weighted_criteria %>% 
+  pivot_longer(cols = c("WT_real_estate":"WT_climRisk"), # selecting criteria cols
+               names_to = "criteria",
+               values_to = "criteria_score") %>% 
+# filter(msa_state != "Portland, ME") %>%  # removing portland (negative number)
+  group_by(cbsa)
+
+tiff(filename = "wt_criteria_student_chart.tiff", width = 8, height = 6, units = "in", res = 300)
+ggplot(data = score_tidy2, aes(y = reorder(msa_state, FINAL_st_score), x = criteria_score)) + 
+    geom_col(aes(fill = criteria)) +
+  labs(x = "Total Score",
+       y = " ",
+       title = "Student") +
+  theme_bw() +
+  guides(fill=guide_legend(title="Criteria (weighted)")) +
+   scale_fill_manual(values = c("palevioletred1","#FDB462",  "gold", "#44AA99", "dodgerblue4", "#BC80BD"),
+                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Solar IRR", "Real Estate", "Health Impacts")) +
+  scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1))
+dev.off()
 ```

### 2023-02-14 src/MSAtotal_scores.Rmd
output real estate metrics to google sheet

@@
 
-tiff(filename = "wt_criteria_client_chart.tiff", width = 8, height = 6, units = "in", res = 300)
+tiff(filename = "wt_criteria_client_chart.tiff",width = 8, height = 6, units = "in", res = 300, pointsize = 12)
 ggplot(data = score_tidy, aes(y = reorder(msa_state, FINAL_owen_SCORE), x = criteria_score)) +
@@
   guides(fill=guide_legend(title="Criteria (weighted)")) +
-   scale_fill_manual(values = c("palevioletred1","#FDB462",  "gold", "#44AA99", "dodgerblue4", "#BC80BD"),
+   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD"),
                      labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Solar IRR", "Real Estate", "Landlord Policy")) +
@@
   guides(fill=guide_legend(title="Criteria (weighted)")) +
-   scale_fill_manual(values = c("palevioletred1","#FDB462",  "gold", "#44AA99", "dodgerblue4", "#BC80BD"),
+   scale_fill_manual(values = c("palevioletred1", "orange",  "gold", "#44AA99", "dodgerblue4", "#6495ED"),
                      labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Solar IRR", "Real Estate", "Health Impacts")) +

### 2023-02-15 src/demo_data_wrangling.Rmd
added copyright line for house price data
cleaning up data, added pseudocode

--- 2022-11-21 src/demo_data_wrangling.Rmd
@@
 date: "5/12/2022"
-output:   html_document
+output:   
+  html_document: 
+  code_folding: hide
 ---
@@
 Preliminary analysis of population change and size narrowed down potential areas which experienced population growth rate greater than 0.5% and have populations greater than 300,000 people.
-In addition, 28 MSAs were removed from analysis, as requested by the client, due to hyper-competitive real estate markets.
-In brief, there are 44 metropolitan statistical areas with potential for ZNE Capital aqcuire real estate.
+In addition, 27 MSAs were removed from analysis, as requested by the client, due to hyper-competitive real estate markets.
+In brief, there are 29 metropolitan statistical areas with potential for ZNE Capital aqcuire real estate.
 
@@
 
-```{r setup, include=FALSE, warning = FALSE, message = FALSE}
+```{r setup, include=TRUE, warning = FALSE, message = FALSE}
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
@@
 library(DT) # create data table
-library(scales) 
+# library(scales) 
 library(lubridate)
@@
 library(kableExtra)
-library(googlesheets4)
-library(tigris) # Census shape files
+library(googlesheets4) # output tables to google sheet
+library(tigris) # imports Census shape files
 library(sf)
@@
 All states were considered with the exception of CA, HI, and AK.
-Initial analysis of Metropolitan Statistical Areas (MSAs) relied on Census data of annual population estimates, in which the population had to be greater than 300,000 individuals for the purpose of attracting investors.
+Initial analysis of Metropolitan Statistical Areas (MSAs) relied on Census data of annual population estimates, in which the population had to be greater than 500,000 individuals for the purpose of attracting investors.
 Population change from 2020 to 2021 was calculated by $$\frac{2021 estimate- 2020 estimate}{2020 estimate} * 100$$
@@
   select("cbsa", "name", "popestimate2020", "popestimate2021", "residual2020", "residual2021") %>% # select data 
-
   separate(col = "name",
@@
            sep = ", ") %>% 
-  mutate(pop_percent_change = (popestimate2021-popestimate2020)/popestimate2020 * 100)  # add population % change column
+  mutate(pop_percent_change = (popestimate2021-popestimate2020)/popestimate2020 * 100)  # calculate population % change
 
@@
 potential_MSAs <- msa_pop_2021_data %>% 
-  filter(pop_percent_change > 0.5 ) %>% 
-  filter(popestimate2021 >= 300000) %>% 
+  filter(pop_percent_change > 0.5 ) %>% # select areas with population growth greater than .5%
+  filter(popestimate2021 >= 500000) %>% 
   mutate(pop_percent_change20_21 = round(pop_percent_change, digits = 3)) %>% 
-  filter (!state %in% c("CA", "HI", "AK")) %>% 
-  #separate(col = "state", into = c("state1", "state2", "state3"),    sep = "-") %>% 
-  filter(!cbsa %in% c("10580", "10900", "12420", "12940", "14260", "15980", "18580", "19100", "19660", "19740", "23540", "25420", "26420", "27260", "29820", "34940", "36260", "36740", "38060", "39100", "39300", "39340", "41620", "45220", "45300", "46060", "46140", "39900")) %>%  # remove MSAs as instructed by Owen
-  mutate(normalized_pop2021 = (popestimate2021-min(popestimate2021))/ (max(popestimate2021)-min(popestimate2021))) %>% 
-  select(cbsa, MSA, state, pop_percent_change20_21, normalized_pop2021, popestimate2021)
+  filter (!state %in% c("CA", "HI", "AK")) %>% # remove areas within these states
+  filter(!cbsa %in% c("10580", "10900", "12420", "12940", "14260", "15980", "18580", "19100", "19660", "19740", "23540", "25420", "26420", "27260", "29820", "34940", "36260", "36740", "38060", "39100", "39300", "39340", "41620", "45220", "45300", "46060", "46140", "39900")) %>%  # remove MSAs as instructed by client
+  select(cbsa, MSA, state, pop_percent_change20_21, popestimate2021)
 
@@
 
-```{r population growth msa map, eval=FALSE}
-# adding column for levels of population growth
-pop_growth_pos <- potential_MSAs %>% 
-  mutate(pop_pos = case_when(
-    pop_percent_change20_21 > 2 ~ "high",
-    pop_percent_change20_21 > 1 ~ "medium",
-    pop_percent_change20_21 > 0 ~ "low"
-  ))
-
-# MSA shapefiles from Census
+```{r population growth map, eval=FALSE}
+## optional visualization
+# read in MSA shapefiles from Census
 cbsa <- core_based_statistical_areas(cb = TRUE) %>% 
   mutate(CBSAFP = as.numeric(CBSAFP)) %>% 
-  inner_join(pop_growth_pos, by = c("CBSAFP" = "cbsa"))
+  inner_join(potential_MSAs, by = c("CBSAFP" = "cbsa")) # merge msa data with shape files
 
@@
                    group = group)) + 
-  geom_sf (data = cbsa, aes(fill = pop_pos)) +
-  labs(title = "MSAs with the highest population growth in 2020")
+  geom_sf (data = cbsa, fill = "lightgreen") +
+  labs(title = "Metro Areas included in this analysis", x = "", y = "")
 ```
@@
 Figure 1.
-Population growth in Metropolitan Statistical Areas (n=44) in 2020.
+Population growth in Metropolitan Statistical Areas (n=29) in 2020-2021.
 Highlighted areas have a population > 300,000 individuals and a population growth rate greater than 0.5%.
@@
 ```{r population change visual, fig.height = 7, fig.align='center', eval = FALSE}
+# create plot of MSAs sorted grom highest population growth to lowest
 ggplot(data = positive_MSA_pop_change, aes(x = pop_percent_change20_21, y = reorder(MSA, pop_percent_change20_21))) + 
@@
   clean_names() %>% 
-  mutate(median_income = as.numeric(estimate_households_median_income_dollars)) %>% 
-  mutate(income_moe_USD = as.numeric(margin_of_error_households_median_income_dollars)) %>% 
+  mutate(median_income = as.numeric(estimate_households_median_income_dollars),
+  income_moe_USD = as.numeric(margin_of_error_households_median_income_dollars)) %>% 
   separate(col = "geography",
@@
            sep = 9) %>% 
-  mutate(cbsa = as.numeric(cbsa)) # change column class from chr to dbl
+  mutate(cbsa = as.numeric(cbsa)) %>% # change column class from chr to dbl
+  inner_join(potential_MSAs)
 
-# join population growth data set to household median income
-income_potential_MSAs <- MSA_income_data %>% # rename row names
-  select(cbsa, median_income) %>% 
-  inner_join(potential_MSAs) %>% 
-  select(cbsa, MSA, state, median_income, pop_percent_change20_21, popestimate2021, normalized_pop2021) #reorder columns
+# join household median income to existing msa dataframe
+potential_MSAs <- potential_MSAs %>% add_column(med_income = MSA_income_data$median_income)
 
@@
 MSA_employment <- unemployment_data %>% 
-  filter(year >= 2019) %>% 
+  filter(year >= 2018) %>% 
   mutate(cbsa = ifelse(cbsa == '76750', "38860", cbsa)) %>% # change south portland, ME to corresponding cbsa
   mutate(cbsa = as.numeric(cbsa)) %>% 
-  full_join(income_potential_MSAs) %>% 
-  drop_na() %>% 
-  select(cbsa, MSA, year, month, employment_rate)
-
-avg_emp_yoy <- MSA_employment %>% 
+  right_join(potential_MSAs) %>% 
   group_by(MSA, year) %>% 
   summarize(avg_employment_rate = mean(employment_rate)) %>% 
-  mutate(avg_emp_YoY = (avg_employment_rate-lag(avg_employment_rate))/lag(avg_employment_rate)*100) %>% 
-  mutate(avg_emp_YoY = round(avg_emp_YoY, digits = 4)) %>% 
-  drop_na()
+  mutate(avg_emp_rate = (avg_employment_rate-lag(avg_employment_rate))/lag(avg_employment_rate)*100, 
+          avg_emp_rate = round(avg_emp_rate, digits = 4)) %>% 
+  drop_na() %>% 
+  mutate(emp_pos = case_when( 
+    avg_emp_rate > 0 ~ "TRUE",
+    avg_emp_rate < 1 ~ "FALSE"))
 
-MSA_emp_income <- income_potential_MSAs %>% 
-  full_join(avg_emp_yoy) %>% 
-  filter(year == "2021") %>% 
-  select(!c(avg_employment_rate, year))
+MSA_employ_rate <- MSA_employment %>% 
+  filter(year == "2021")
 
-```
-
-```{r}
-# adding portland employment rates from https://data.bls.gov/pdq/SurveyOutputServlet
-portland_employment_data <- unemployment_data %>% 
-  filter(year >= 2019) %>% 
-  filter(cbsa == "76750") %>% 
-   group_by(year) %>% 
-  summarize(avg_employment_rate = mean(employment_rate)) %>% 
-  mutate(avg_emp_YoY = (avg_employment_rate-lag(avg_employment_rate))/lag(avg_employment_rate)*100) %>% 
-  mutate(avg_emp_YoY = round(avg_emp_YoY, digits = 4))
-
-employment_data_jan_2021 <- 100-4.9
-employment_data_jan_2022 <- 100- 3.6
-
-yoy <- (employment_data_jan_2022 - employment_data_jan_2021)/ employment_data_jan_2021 *100
+potential_MSAs$employ_change <- MSA_employ_rate$avg_emp_rate
 
@@
 # visual of MSA employment rates, shows major drop during covid
-ggplot(MSA_employment_yoy, aes(x = year, y = employment_rate, color = MSA)) +
+ggplot(MSA_employment, aes(x = year, y = avg_employment_rate, color = MSA)) +
   geom_line() +
-  xlim(2021, 2022) +
+  xlim(2019, 2022) +
   theme(legend.position = "none") +
-
   labs(x = "Year",
        y = "Employment Rate",
-      title = "Employment Rate Trend since 2016")
+      title = "Employment Rate Trend since COVID-19 Pandemic")
 
 # year over year change in employment rate, split up by year
-ggplot(MSA_employment_yoy, aes(x = employment_YoY, y = reorder(MSA, employment_YoY), fill = emp_pos)) +
-  geom_col(position = "identity", color = "black", size = 0.25) +
+ggplot(MSA_employment, aes(x = avg_emp_rate, y = reorder(MSA, avg_emp_rate))) +
+  geom_col(position = "identity", aes(fill = emp_pos), size = 0.25) +
   scale_fill_manual(values = c("red", "darkolivegreen3"), guide = FALSE) +
   facet_grid(~year) +
-  labs(x = "Year-over_year change in employment rate",
+  labs(x = "Change in employment rate",
        y = " ")
@@
 ```{r median rent}
-# median rents from hud
+# data wrangling 2021 median rent
 MSA_median_rent2021_data <- read_csv(here("data", "raw", "2021MSA_medianrent.csv")) %>% 
@@
            sep = -5) %>% 
-  mutate(cbsa = ifelse(cbsa == 'M6400', "38860", cbsa)) %>% 
-  mutate(cbsa = as.numeric(cbsa)) %>% 
-  mutate(median_rent2021 = as.numeric(rent50_1)) %>% 
+  mutate(cbsa = ifelse(cbsa == 'M6400', "38860", cbsa),
+        cbsa = as.numeric(cbsa) ,
+        median_rent2021 = as.numeric(rent50_1)) %>% # select median rent for 1 bedroom apartments
   select(cbsa, median_rent2021) %>% 
@@
 
+# 2022 median rent
 MSA_median_rent2022_data <- read_csv(here("data", "raw", "2022MSA_median_rent.csv")) %>% 
@@
            sep = -5) %>% 
-  mutate(cbsa = ifelse(unknown == '38860M', "38860", cbsa)) %>% 
-  mutate(median_rent2022 = as.numeric(rent_50_1)) %>% 
+  mutate(cbsa = ifelse(unknown == "38860M" & areaname22 == "Portland, ME HUD Metro FMR Area", "38860", cbsa),
+         cbsa = as.numeric(cbsa),
+         median_rent2022 = as.numeric(rent_50_1)) %>% 
   select(cbsa, median_rent2022) %>% 
-  mutate(cbsa = as.numeric(cbsa)) %>% 
-  right_join(MSA_median_rent2021_data)
+  right_join(potential_MSAs)
 
-potential_MSAs <- MSA_median_rent2022_data %>% 
-  left_join(MSA_emp_income)
+potential_MSAs <- potential_MSAs %>% 
+  add_column(med_rent2021 = MSA_median_rent2021_data$median_rent2021, # add rent data
+             med_rent_2022 = MSA_median_rent2022_data$median_rent2022)
 
@@
 Median housing prices by MSA were obtained from the Association of Realtors for 2021. 
+Copyright ©[2021] NATIONAL ASSOCIATION OF REALTORS®. All rights reserved.  Reprinted with permission.
 
@@
 metro_house_cost_data <- read_csv(here("data","raw", "metro_home_costs", "metro-home-prices-2022-11-10.csv")) %>% 
-  row_to_names(3) %>% 
+  row_to_names(3) %>% # make sure column headers are correct
   clean_names() %>% 
@@
            sep = 5) %>% 
-  mutate(cbsa = as.numeric(cbsa)) %>% 
-  mutate(x2021 = as.numeric(x2021)) %>% 
-  mutate(x2021 = x2021 * 1000) %>% 
-  rename(house_price_2021 = x2021)
+  mutate(cbsa = as.numeric(cbsa),
+        x2021 = as.numeric(x2021), 
+        house_price_2021 = x2021 * 1000) %>% # data is proivded as thousands, so multiply to get correct ratio
+  right_join(potential_MSAs) %>% 
+  mutate(house_price_2021 = ifelse(cbsa == "12260", "284900", house_price_2021), 
+        house_price_2021 = ifelse(cbsa == "32580", "259000", house_price_2021))
 
-metro_house_prices <- metro_house_cost_data %>%
-  select(cbsa, house_price_2021) %>% 
-  right_join(potential_MSAs) %>% 
-  select(cbsa, MSA, state, pop_percent_change20_21, normalized_pop2021, popestimate2021, median_income, avg_emp_YoY, median_rent2021, median_rent2022, house_price_2021) %>% 
-  mutate(house_price_2021 = ifelse(cbsa == "12260", "284900", house_price_2021)) %>% 
-  mutate(house_price_2021 = ifelse(cbsa == "32580", "259000", house_price_2021))
+potential_MSAs$house_price_2021 <- metro_house_cost_data$house_price_2021
 
-MSAs <-metro_house_prices %>% 
-  filter(popestimate2021 >= 500000) %>% 
-  distinct(cbsa, .keep_all = TRUE) # remove duplicate portland row
 ```
@@
 
-```{r, house price visual, fig.height = 7, fig.align='center', eval = FALSE}
-ggplot(metro_house_prices, aes(x = houseprice_change21_22, y = reorder(MSA, houseprice_change21_22))) +
-  geom_col(fill = "dodgerblue4") +
-  geom_text(aes(label = houseprice_change21_22),
-            size = 4,
-            hujust = 1, nudge_x = -1,
-            color = "white") +
-  theme_minimal()  +
-  labs(x = "% change",
-       y = "",
-       title = "Change in House Prices in 2021")
-
-```
-
 ### MSA data table
@@
 # table seperated by MSAs
-datatable(MSAs, extensions = 'Buttons', options = list(
+datatable(potential_MSAs, extensions = 'Buttons', options = list(
     dom = 'Bfrtip',
@@
 
-#gs4_create("updated_MSAs")
-sheet_write(MSAs, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "MSAs")
+#gs4_create("real_estate_inputs")
+#output data into google sheet
+sheet_write(potential_MSAs, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "MSA_demographics")
 ```
@@
 #  split up MSAs into cities to collect utility, rent, and landlord data.
-msa_seperated <- MSAs %>% 
+msa_seperated <- potential_MSAs %>% 
   separate_rows(MSA, sep = "-") %>% 
@@
 
-#gs4_create("updated_MSAs")
+#gs4_create("real_estate_inputs")
 sheet_write(msa_seperated, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "cities")
 
-# table seperated by city
-datatable(msa_seperated, extensions = 'Buttons', options = list(
-    dom = 'Bfrtip',
-    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
-  )
-)
-
 ```

### 2023-02-16 src/anchor_city_real_estate.Rmd
output real estate metrics to google sheet
rewriting object names so they are more clear

@@
 library(tidyverse)
-library(janitor)
+library(janitor) # clean_names()
 library(googlesheets4)
 library(DT)
-library(MCDA)
 
-city_realEstate_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990")
+city_occupancy_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990") %>% 
+  clean_names() %>% 
+  select(!x7) %>% 
+  rename(occ_rate2021 = percent_occupancy_2021)
 
-select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")
+MSA_dempgraphic_data <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing", sheet = "MSA_demographics")
 
@@
 ```{r}
-# aggregate yardi data into MSAs by cbsa
-city_realEstate_data <- city_realEstate_data %>% 
-  clean_names() %>% 
-  select(!x7) %>% 
-  rename(avg_rent2022 = x2022_rent) %>% 
-  rename(occ_rate2021 = percent_occupancy_2021)
 
@@
 
-anchor_city_data <- city_realEstate_data %>% 
+anchor_city_data <- city_occupancy_data %>% 
   filter(city %in% c(anchor_cities_list))
@@
   mutate(med_rent_to_income = (median_rent2021*12)/median_income)
+
+sheet_write(compliedMSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=426650684", sheet = "real_estate_metrics")
 

### 2023-02-16 solar_fam_shiny/app.R
created shiny file

@@ -1,0 +1,72 @@
+
+
+library(shiny)
+library(ggplot2)
+library(tidyverse)
+library(bslib)
+
+
+# Define UI for application that draws a histogram
+ui <- fluidPage(
+  # Application title
+  titlePanel("TITLE HEre "),
+  
+  # Sidebar with a slider input for number of bins 
+  sidebarLayout(
+    sidebarPanel("Put my widgets here",
+                 radioButtons(inputId = "penguin_species",
+                              label = "Select a criteria to map",
+                              choices = c("Real Estate", "Landlord Policy", "Solar IRR", "Climate Risk")), # left value is displayed
+                 ### select input box
+                 "Choice a color",
+                 selectInput(inputId = 'pt_color',
+                             label = "choose your favorite color!",
+                             choices = c("Awesome red" = "red",
+                                         "Pretty purple" = "purple",
+                                         "cornflowerblue",
+                                         "orange"))
+    ), ### end sidebarPanel
+    
+    mainPanel("put my graph here!",
+              plotOutput(outputId = "penguin_plot"),
+              tableOutput(outputId = "penguin_table")
+    ) ### end mainPanel
+    
+  ) ### end sidebarLayout
+  
+) ### end fluidPage
+
+
+
+# Define server logic required to draw a histogram
+server <- function(input, output) {
+  penguin_select <- reactive({ ### whatever is in reactive function()
+    penguins %>%
+      filter(species == input$penguin_species) # penguin species comes from inputId in radio buttons
+  })
+  
+  penguin_table <- reactive({
+    penguins %>%
+      filter(species == input$penguin_species) %>%
+      group_by(sex) %>%
+      summarize(mean_flip = mean(flipper_length_mm),
+                mean_mass = mean(body_mass_g))
+  }) ### end penguin table
+  
+  output$penguin_plot <- renderPlot({
+    ggplot(data = penguin_select(),
+           aes(x = flipper_length_mm, y = body_mass_g )) +
+      geom_point(color = input$pt_color) + ### change color based on selection
+      theme_minimal()
+  })
+  
+  output$penguin_table <- renderTable({
+    penguin_table()
+  })
+}
+
+# Run the application 
+shinyApp(ui = ui, server = server)
+
+
+

### 2023-02-21 solar_fam_shiny/app.R
trying to read in census data

@@ -4,64 +4,90 @@ library(shiny)
 library(ggplot2)
 library(tidyverse)
 library(bslib)
+library(tidycensus)
+library(RJSONIO)
+library(googlesheets4)
+library(tigris) # for shape files
+options(tigris_use_cache = TRUE)
+
+census_api_key("ebab921b3002df9b71881ad6c426f34281ce0e11", overwrite = "TRUE")
+resURL <- "https://api.census.gov/data/2021/acs/acs1?get=NAME,B01001_001E&for=metropolitan%20statistical%20area/micropolitan%20statistical%20area:35620&key=ebab921b3002df9b71881ad6c426f34281ce0e11" 
+ljson <- fromJSON(resURL)
+
+# reads in metro areas
+census_data <- get_acs(geography = "cbsa",  year = 2021, 
+              variables = c("B19013_001E", # median household income
+              "B23025_003E", # total civilian labor force
+              "B23025_004E"),   # employed labor force
+              output = "wide",
+              keep_geo_vars = TRUE) %>% 
+  rename("median_income" = "B19013_001E")
+
+# census metro area shape files
+cbsa_shp <- core_based_statistical_areas() 
+
+real_estate_inputs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=519968233")
 
 
-# Define UI for application that draws a histogram
 ui <- fluidPage(
   # Application title
-  titlePanel("TITLE HEre "),
-  
-  # Sidebar with a slider input for number of bins 
-  sidebarLayout(
-    sidebarPanel("Put my widgets here",
-                 radioButtons(inputId = "penguin_species",
-                              label = "Select a criteria to map",
-                              choices = c("Real Estate", "Landlord Policy", "Solar IRR", "Climate Risk")), # left value is displayed
-                 ### select input box
-                 "Choice a color",
-                 selectInput(inputId = 'pt_color',
-                             label = "choose your favorite color!",
-                             choices = c("Awesome red" = "red",
-                                         "Pretty purple" = "purple",
-                                         "cornflowerblue",
-                                         "orange"))
-    ), ### end sidebarPanel
-    
-    mainPanel("put my graph here!",
-              plotOutput(outputId = "penguin_plot"),
-              tableOutput(outputId = "penguin_table")
-    ) ### end mainPanel
-    
-  ) ### end sidebarLayout
-  
[…]

### 2023-02-22 solar_fam_shiny/app.R
trying to read in census data

@@ -11,17 +11,24 @@ library(tigris) # for shape files
 options(tigris_use_cache = TRUE)
 
 census_api_key("ebab921b3002df9b71881ad6c426f34281ce0e11", overwrite = "TRUE")
-resURL <- "https://api.census.gov/data/2021/acs/acs1?get=NAME,B01001_001E&for=metropolitan%20statistical%20area/micropolitan%20statistical%20area:35620&key=ebab921b3002df9b71881ad6c426f34281ce0e11" 
-ljson <- fromJSON(resURL)
+popURL <-"http://api.census.gov/data/2021/acs/acs1?key=ebab921b3002df9b71881ad6c426f34281ce0e11"
+  #"https://api.census.gov/data/2021/acs/acs1?get=NAME,B01001_001E&for=metropolitan%20statistical%20area/micropolitan%20statistical%20area:35620&key=ebab921b3002df9b71881ad6c426f34281ce0e11" 
+ljson <- fromJSON(popURL)
+
+pop_est <- get_estimates(geography = "metropolitan statistical area/micropolitan statistical area",
+                         product = "housing")
+
+msa_flows <- get_flows(geography = "metropolitan statistical area/micropolitan statistical area",
+                      year = 2021)
+
 
 # reads in metro areas
 census_data <- get_acs(geography = "cbsa",  year = 2021, 
-              variables = c("B19013_001E", # median household income
-              "B23025_003E", # total civilian labor force
-              "B23025_004E"),   # employed labor force
+              variables = c(median_income = "B19013_001E", # median household income
+              total_labor_force = "B23025_003E", # total civilian labor force
+              employed_labor_force = "B23025_004E"),   # employed labor force
               output = "wide",
-              keep_geo_vars = TRUE) %>% 
-  rename("median_income" = "B19013_001E")
+              keep_geo_vars = TRUE) 
 
 # census metro area shape files
 cbsa_shp <- core_based_statistical_areas() 

### 2023-02-22 src/MSAtotal_scores.Rmd
trying to create a function to calculate weighted scores

@@
 
-criteria_all <- criteria_all[, col_order]
+criteria_all <- criteria_all[, col_order] %>% 
+  select(!inverted_n_eal_score)
 
@@
 
-```{r original weights, eval = FALSE}
-# weight criteria for Owen
-criteria_weights_owen <- c(0.37, 0.12, 0.22, 0.07, 0.0, 0.07, 0.15)
+```{r}
+# weighting criteria
+weighted_criteria <- criteria_all %>% 
+  select(4:10)
 
-#make landlord_score column
-criteria_all <- criteria_all %>% 
-  mutate(total_score_owen = cbsa, #this is just a place holder
-         real_estate_score = as.numeric(real_estate_score),
-         landlord_score = as.numeric(landlord_score)
-  )
-
-size <- dim(criteria_all)
-for (i in 1:size[1]) {
-  cbsa_data <- c(criteria_all$real_estate_score[i],
-                criteria_all$landlord_score[i],
-                 criteria_all$normalized_irr[i],
-                 criteria_all$inverted_n_total_co2[i],
-                criteria_all$inverted_n_health_emissions_cost[i],
-                criteria_all$normalized_annual_renewable[i],
-                criteria_all$normalized_inverted_climate_risk[i])
-  criteria_all$total_score_owen[i] = weighted.mean(cbsa_data, criteria_weights_owen)
-}
-
+weighted <- t(t(weighted_criteria)*criteria_weights_client)
 ```
@@
   mutate(final_tot_students = cbsa,
-         stud_EAL_total = cbsa,
          real_estate_score = as.numeric(real_estate_score),
@@
 
-```{r student weights with EAL, eval = FALSE}
-size <- dim(criteria_all)
-for (i in 1:size[1]) {
-  cbsa_data <- c(criteria_all$real_estate_score[i],
-                criteria_all$landlord_score[i],
-                 criteria_all$normalized_irr[i],
-                 criteria_all$inverted_n_total_co2[i],
-                criteria_all$inverted_n_health_emissions_cost[i],
-                criteria_all$normalized_annual_renewable[i],
-                criteria_all$inverted_n_eal_score[i]
-                  )
-  criteria_all$stud_EAL_total[i] = weighted.mean(cbsa_data, criteria_weights_students)
-}
-```
+```{r Client FINAL WEIGHTS}
 
-```{r OWEN FINAL WEIGHTS climate risk 50 and proportional weights for others}
-
-weights_climate50_prop <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)
+client_weights <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)
 
@@
 criteria_all <- criteria_all %>% 
-  mutate(owen_FINAL_score = cbsa, #this is just a place holder
+  mutate(client_FINAL_score = cbsa, #this is just a place holder
          score_climate50_prop_EAL = cbsa)
@@
                   )
-  criteria_all$owen_FINAL_score[i] = weighted.mean(cbsa_data, weights_climate50_prop) 
+  criteria_all$client_FINAL_score[i] = weighted.mean(cbsa_data, client_weights) 
 }
 
-```
-
-```{r Climate with 50 & proportional others with EAL, eval=FALSE}
-### Using climate weight 50 and proportional others with EAL
-for (i in 1:size[1]) {
-  cbsa_data <- c(criteria_all$real_estate_score[i],
-                criteria_all$landlord_score[i],
-                 criteria_all$normalized_irr[i],
-                 criteria_all$inverted_n_total_co2[i],
-                criteria_all$inverted_n_health_emissions_cost[i],
-                criteria_all$normalized_annual_renewable[i],
-                criteria_all$inverted_n_eal_score[i]
-                  )
-  criteria_all$score_climate50_prop_EAL[i] = weighted.mean(cbsa_data, weights_climate50_prop) 
-}
-
-```
-
-```{r climate risk 35 and Jan 31 weights from owen, eval = FALSE}
-
-weights_climate35_owen <- c(0.25, 0.1, 0.15, 0.075, 0.0, 0.075, 0.35)
-
-#make landlord_score column
-criteria_all <- criteria_all %>% 
-  mutate(score_climate35_owen = cbsa)#this is just a place holder
-
-size <- dim(criteria_all)
-for (i in 1:size[1]) {
-  cbsa_data <- c(criteria_all$real_estate_score[i],
-                criteria_all$landlord_score[i],
-                 criteria_all$normalized_irr[i],
-                 criteria_all$inverted_n_total_co2[i],
-                criteria_all$inverted_n_health_emissions_cost[i],
-                criteria_all$normalized_annual_renewable[i],
-                criteria_all$normalized_inverted_climate_risk[i]
-                  )
-  criteria_all$score_climate35_owen[i] = weighted.mean(cbsa_data, weights_climate35_owen) 
-}
 ```
@@
          electricity = normalized_annual_renewable,
-         climate_risk = normalized_inverted_climate_risk,
-         climate_EAL_risk = inverted_n_eal_score) %>% 
+         climate_risk = normalized_inverted_climate_risk) %>% 
   mutate_if(is.numeric, ~round(., 5))
@@
 ```
+
+### Data Visualization- barcharts
 
@@
 
-```{r}
+```{r student data viz}
 ## read in weighted criteria sheet

### 2023-02-23 solar_fam_shiny/app.R
creating new branch

@@ -12,23 +12,22 @@ options(tigris_use_cache = TRUE)
 
 census_api_key("ebab921b3002df9b71881ad6c426f34281ce0e11", overwrite = "TRUE")
 popURL <-"http://api.census.gov/data/2021/acs/acs1?key=ebab921b3002df9b71881ad6c426f34281ce0e11"
-  #"https://api.census.gov/data/2021/acs/acs1?get=NAME,B01001_001E&for=metropolitan%20statistical%20area/micropolitan%20statistical%20area:35620&key=ebab921b3002df9b71881ad6c426f34281ce0e11" 
+#"https://api.census.gov/data/2021/acs/acs1?get=NAME,B01001_001E&for=metropolitan%20statistical%20area/micropolitan%20statistical%20area:35620&key=ebab921b3002df9b71881ad6c426f34281ce0e11" 
 ljson <- fromJSON(popURL)
 
 pop_est <- get_estimates(geography = "metropolitan statistical area/micropolitan statistical area",
                          product = "housing")
 
 msa_flows <- get_flows(geography = "metropolitan statistical area/micropolitan statistical area",
-                      year = 2021)
-
+                       year = 2021)
 
 # reads in metro areas
 census_data <- get_acs(geography = "cbsa",  year = 2021, 
-              variables = c(median_income = "B19013_001E", # median household income
-              total_labor_force = "B23025_003E", # total civilian labor force
-              employed_labor_force = "B23025_004E"),   # employed labor force
-              output = "wide",
-              keep_geo_vars = TRUE) 
+                       variables = c(median_income = "B19013_001E", # median household income
+                                     total_labor_force = "B23025_003E", # total civilian labor force
+                                     employed_labor_force = "B23025_004E"),   # employed labor force
+                       output = "wide",
+                       keep_geo_vars = TRUE) 
 
 # census metro area shape files
 cbsa_shp <- core_based_statistical_areas() 

### 2023-02-23 solar_fam_shiny/app.R
removing census data

@@ -11,24 +11,6 @@ library(tigris) # for shape files
 options(tigris_use_cache = TRUE)
 
 census_api_key("ebab921b3002df9b71881ad6c426f34281ce0e11", overwrite = "TRUE")
-popURL <-"http://api.census.gov/data/2021/acs/acs1?key=ebab921b3002df9b71881ad6c426f34281ce0e11"
-  #"https://api.census.gov/data/2021/acs/acs1?get=NAME,B01001_001E&for=metropolitan%20statistical%20area/micropolitan%20statistical%20area:35620&key=ebab921b3002df9b71881ad6c426f34281ce0e11" 
-ljson <- fromJSON(popURL)
-
-pop_est <- get_estimates(geography = "metropolitan statistical area/micropolitan statistical area",
-                         product = "housing")
-
-msa_flows <- get_flows(geography = "metropolitan statistical area/micropolitan statistical area",
-                      year = 2021)
-
-
-# reads in metro areas
-census_data <- get_acs(geography = "cbsa",  year = 2021, 
-              variables = c(median_income = "B19013_001E", # median household income
-              total_labor_force = "B23025_003E", # total civilian labor force
-              employed_labor_force = "B23025_004E"),   # employed labor force
-              output = "wide",
-              keep_geo_vars = TRUE) 
 
 # census metro area shape files
 cbsa_shp <- core_based_statistical_areas() 

### 2023-02-23 solar_fam_shiny/app.R
removing google sheet data

@@ -32,8 +32,6 @@ census_data <- get_acs(geography = "cbsa",  year = 2021,
 # census metro area shape files
 cbsa_shp <- core_based_statistical_areas() 
 
-real_estate_inputs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=519968233")
-
 
 ui <- fluidPage(
   # Application title

### 2023-02-24 solar_fam_shiny/app.R
created input for population size, but output doesnt work

@@ -15,53 +15,78 @@ census_api_key("ebab921b3002df9b71881ad6c426f34281ce0e11", overwrite = "TRUE")
 # census metro area shape files
 cbsa_shp <- core_based_statistical_areas() 
 
-real_estate_inputs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=519968233")
+real_estate_metrics2021 <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=519968233")
 
 
 ui <- fluidPage(
   # Application title
   titlePanel("Identifying the best areas for rooftop solar on multifamily housing"),
   navbarPage("Solar Family", #navbarPage allows us to create our tabs
-             tabPanel("Landing Page", #this is how we add tabs.
+             
+             ### tab 1 - background
+             tabPanel("Landing Page", 
+                      #sidebarLayout(
+                        #sidebarPanel(" ",
+                          mainPanel(
+                            h2("Background"),
+                            h5("Sixth level title"),
+                            plotOutput("map"),
+                            h4("model flow chart") ### Add model image here
+                            ) ### end mainPanel
+                       # ) #end sidebar panel
+                      # ) #end sidebar layour
+             ), # ## END of tab 1
+             
+             
+             ### tab2
+             tabPanel("Coefficient Customization",
                       sidebarLayout(
-                        sidebarPanel("WIDGETS",
-                                     helpText("Create demographic maps with information from the 2010 US Census."),
-                                     selectInput("var", 
-                                                 label = "Choose a variable to display",
-                                                 choices = c("Population Growth", "Employment Growth", "Rent Change"),
-                                                 selected = "Population Growth"),
-               
-                        ), #End sidebarPanel widgets
-               
-                        mainPanel(h2("Background"),
-                                  h5("Sixth level title"),
-                                  "put my graph here!",
-                                  plotOutput("map"),
-                                  tableOutput(outputId = "census_table"),
-                                  
-                        ), ### end mainPanel
-                        
-                       ) #end sidebar (tab1) layout
-             ), # end tabpanel thing 1
-             tabPanel("Metric Data Exploration",
-             sidebarPanel(" ",
-                          radioButtons(inputId = "criteria_select",
[…]

### 2023-02-24 solar_fam_shiny/app.R
added input for minimum population size

@@ -3,23 +3,26 @@
 library(shiny)
 library(ggplot2)
 library(tidyverse)
-library(bslib)
+#library(bslib)
 library(tidycensus)
-library(RJSONIO)
 library(googlesheets4)
 library(tigris) # for shape files
 options(tigris_use_cache = TRUE)
 
-census_api_key("ebab921b3002df9b71881ad6c426f34281ce0e11", overwrite = "TRUE")
+#census_api_key("ebab921b3002df9b71881ad6c426f34281ce0e11", overwrite = "TRUE")
 
 # census metro area shape files
 cbsa_shp <- core_based_statistical_areas() 
 
 real_estate_metrics2021 <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=519968233")
+criteria_unweighted <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202", sheet = "FINAL_total_score")
 
-
+model_img <- img(src = "model_no_weights.jpg",
+                 alt = "Model Visualization",
+                 width="500%")
+  
 ui <- fluidPage(
-  # Application title
+  # Application title43
   titlePanel("Identifying the best areas for rooftop solar on multifamily housing"),
   navbarPage("Solar Family", #navbarPage allows us to create our tabs
              
@@ -31,7 +34,13 @@ ui <- fluidPage(
                             h2("Background"),
                             h5("Sixth level title"),
                             plotOutput("map"),
-                            h4("model flow chart") ### Add model image here
+                           
+                             h4("model flow chart"), 
+                            ### Add model image here
+                            div(img(src = "model_no_weights.jpg", 
+                                    width = "800",
+                                    height = "400"), 
+                                style="text-align: center;")
                             ) ### end mainPanel
                        # ) #end sidebar panel
                       # ) #end sidebar layour
@@ -45,16 +54,16 @@ ui <- fluidPage(
                           # input for population size
                           textInput(inputId = "PopulationSizeMin",
                                     label = "Minimum Population Size",
-                                    value = "300,000",
+                                    value = "300000",
                                     width = "100px"), # end population input
                          
                            # help text under input box
-                          helpText("Format example: 100,000"),
+                          helpText("Format example: 100000"),
[…]

### 2023-02-25 solar_fam_shiny/app.R
started to make tmap

@@ -8,19 +8,30 @@ library(tidycensus)
 library(googlesheets4)
 library(tigris) # for shape files
 options(tigris_use_cache = TRUE)
+library(tmap) # interactive map
+# map states and US
+library(sf)
+
+
+
+real_estate_metrics2021 <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=519968233")
+criteria_unweighted <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202", sheet = "FINAL_total_score") %>% 
+  mutate(cbsa = as.character(cbsa))
 
 #census_api_key("ebab921b3002df9b71881ad6c426f34281ce0e11", overwrite = "TRUE")
 
 # census metro area shape files
-cbsa_shp <- core_based_statistical_areas() 
+msa_shp <- core_based_statistical_areas() %>% 
+  janitor::clean_names() %>% 
+  select(!c(csafp, lsad, cbsafp))
 
-real_estate_metrics2021 <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=519968233")
-criteria_unweighted <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202", sheet = "FINAL_total_score")
+state_shp <- data(state_laea) %>% 
+  st_transform(state_laes, st_crs(msa_shp))
 
-model_img <- img(src = "model_no_weights.jpg",
-                 alt = "Model Visualization",
-                 width="500%")
-  
+# merge criteria table with shp data
+unwt_criteria_shp <- msa_shp %>% 
+  merge(msa_shp, by.x = "cbsa", by.y = "geoid")
+
 ui <- fluidPage(
   # Application title43
   titlePanel("Identifying the best areas for rooftop solar on multifamily housing"),
@@ -29,20 +40,19 @@ ui <- fluidPage(
              ### tab 1 - background
              tabPanel("Landing Page", 
                       #sidebarLayout(
-                        #sidebarPanel(" ",
-                          mainPanel(
[…]

### 2023-02-25 solar_fam_shiny/app.R
map still wont output

@@ -25,11 +25,11 @@ msa_shp <- core_based_statistical_areas() %>%
   janitor::clean_names() %>% 
   select(!c(csafp, lsad, cbsafp))
 
-state_shp <- data(state_laea) %>% 
-  st_transform(state_laes, st_crs(msa_shp))
+state_shp <- states(class = "sf", progress_bar = TRUE) %>% 
+  st_transform(st_crs(msa_shp))
 
 # merge criteria table with shp data
-unwt_criteria_shp <- msa_shp %>% 
+unwt_criteria_shp <- criteria_unweighted %>% 
   merge(msa_shp, by.x = "cbsa", by.y = "geoid")
 
 ui <- fluidPage(

### 2023-03-14 src/MSAtotal_scores.Rmd
renamed student scores to equity scenario

@@
 library(googlesheets4)
-
+library(dplyr)
 ```
@@
 climate_risk_data <- read_sheet("https://docs.google.com/spreadsheets/d/1HZCeNGf61RJ9GJs3cLJcl9aNpvmPMBF1Yoss2fytltc/edit#gid=2045475301", sheet = "climate_risk+EAL") %>% 
-  select(cbsa, normalized_inverted_climate_risk, inverted_n_eal_score)
+  select(cbsa, normalized_inverted_climate_risk)
 
@@
 # reorder columns
-col_order <-c("cbsa", "city_msa", "state", "real_estate_score", "landlord_score", "normalized_annual_renewable", "normalized_irr", "inverted_n_total_co2", "normalized_inverted_climate_risk", "inverted_n_eal_score", "inverted_n_health_emissions_cost")
+col_order <-c("cbsa", "city_msa", "state", "real_estate_score", "landlord_score", "normalized_annual_renewable", "normalized_irr", "inverted_n_total_co2",  "inverted_n_health_emissions_cost", "normalized_inverted_climate_risk")
 
 criteria_all <- criteria_all[, col_order] %>% 
-  select(!inverted_n_eal_score)
+  rename(financial_score = normalized_irr,
+         co2_e_score = inverted_n_total_co2,
+         health_impact_score = inverted_n_health_emissions_cost,
+         electricity_score = normalized_annual_renewable,
+         climate_risk_score = normalized_inverted_climate_risk) %>% 
+  mutate(real_estate_score = as.numeric(real_estate_score),
+         landlord_score = as.numeric(landlord_score)) %>% 
+  mutate_if(is.numeric, ~round(., 5))
 
-```
-
-```{r}
-# weighting criteria
-weighted_criteria <- criteria_all %>% 
-  select(4:10)
-
-weighted <- t(t(weighted_criteria)*criteria_weights_client)
 ```
@@
 criteria_all <- criteria_all %>% 
-  mutate(final_tot_students = cbsa,
-         real_estate_score = as.numeric(real_estate_score),
-         landlord_score = as.numeric(landlord_score)) #this is just a place holder
+  mutate(equity_scenario_score = cbsa,
+         client_FINAL_score = cbsa) #this is just a place holder
 
-size <- dim(criteria_all)
-for (i in 1:size[1]) {
+for (i in 1:nrow(criteria_all)) {
   cbsa_data <- c(criteria_all$real_estate_score[i],
                 criteria_all$landlord_score[i],
-                 criteria_all$normalized_irr[i],
-                 criteria_all$inverted_n_total_co2[i],
-                criteria_all$inverted_n_health_emissions_cost[i],
-                criteria_all$normalized_annual_renewable[i],
-                criteria_all$normalized_inverted_climate_risk[i]
+                 criteria_all$financial_score[i],
+                 criteria_all$co2_e_score[i],
+                criteria_all$health_impact_score[i],
+                criteria_all$electricity_score[i],
+                criteria_all$climate_risk_score[i]
                   )
-  criteria_all$final_tot_students[i] = weighted.mean(cbsa_data, criteria_weights_students)
+  criteria_all$equity_scenario_score[i] = weighted.mean(cbsa_data, criteria_weights_students)
 }
@@
 
-#make landlord_score column
-criteria_all <- criteria_all %>% 
-  mutate(client_FINAL_score = cbsa, #this is just a place holder
-         score_climate50_prop_EAL = cbsa)
+# multiply column by respective weight and sum for total
 
-size <- dim(criteria_all)
-for (i in 1:size[1]) {
+for (i in 1:nrow(criteria_all)) {
   cbsa_data <- c(criteria_all$real_estate_score[i],
                 criteria_all$landlord_score[i],
-                 criteria_all$normalized_irr[i],
-                 criteria_all$inverted_n_total_co2[i],
-                criteria_all$inverted_n_health_emissions_cost[i],
-                criteria_all$normalized_annual_renewable[i],
-                criteria_all$normalized_inverted_climate_risk[i]
+                 criteria_all$financial_score[i],
+                 criteria_all$co2_e_score[i],
+                criteria_all$health_impact_score[i],
+                criteria_all$electricity_score[i],
+                criteria_all$climate_risk_score[i]
                   )
@@
 
-```{r rename columns and output to google sheet}
+```{output to google sheet}
 
 criteria_all <- criteria_all %>% 
-  rename(financial_score = normalized_irr,
-         co2_e = inverted_n_total_co2,
-         health_impact = inverted_n_health_emissions_cost,
-         electricity = normalized_annual_renewable,
-         climate_risk = normalized_inverted_climate_risk) %>% 
   mutate_if(is.numeric, ~round(., 5))
+
+weighted <- t(t(weighted_criteria)*criteria_weights_client)
 
@@
   unite("msa_state", city_msa:state, sep = ", ") 
+ # mutate(test2 = rowSums(client_weighted_criteria[,c(3:9)]))
 
@@
                values_to = "criteria_score") %>% 
- filter(msa_state != "Portland, ME") %>%  # removing portland (negative number)
-  group_by(cbsa)
+ filter(msa_state != "Portland, ME")  # removing portland (negative number)
 
@@
                names_to = "criteria",
-               values_to = "criteria_score") %>% 
-# filter(msa_state != "Portland, ME") %>%  # removing portland (negative number)
-  group_by(cbsa)
+               values_to = "criteria_score")
 

### 2023-03-14 solar_fam_shiny/app.R
added text about project and additional tabs for data

@@ -44,9 +44,11 @@ ui <- fluidPage(
              theme = bs_theme(bootswatch = "cerulean", primary = "#8aab57"),
              ### tab 1 - background
              tabPanel("Landing Page", 
-                      mainPanel("Background Info",
-                        h5("Sixth level title"),
-                        h4("model flow chart"),  ### Add model image here
+                      mainPanel(
+                        h5("Importance & Objectives"),
+                        h5("Background"),
+                        h5("Model Flow Chart"),  ### Add model image here
+                        p("Visualization of methodology to calculate investment favorability score for each metropolitan area. Seven criteria comprise the total score, with each weighted by a value given preferences."),
                         div(img(src = "model_no_weights.jpg",  width = "800", height = "400"), style="text-align: center;")
                       ) ### end mainPanel
              ), # ## END of tab 1
@@ -76,16 +78,16 @@ ui <- fluidPage(
                                                              width = "100px",min = 0.0,max = 1.0, step = 0.05), # end health impact weight input
                                      numericInput(inputId = "SolarIrrWtInput", label = "Solar IRR",
                                                           value = 0, width = "100px", min = 0.0, max = 1.0, step = 0.05), # end irr input
-                                     actionButton(inputId = "EnterWeights", label = "Calculate")),
-                                 
-                        mainPanel(
-                          verbatimTextOutput(outputId = "wt_sum"), 
-                          plotOutput(outputId = "wt_criteria_chart"))
-                        ) # tab 1end main panel
-                        ), # end 2 tabpanel
-                        
+                                     actionButton(inputId = "EnterWeights", label = "Calculate")), # end sidebar panel
+                                   
+                                   mainPanel(
+                                     verbatimTextOutput(outputId = "wt_sum"), 
+                                     h5("Ranking the top metropolitan areas based on solar PV investment favorability"),
+                                     plotOutput(outputId = "wt_criteria_chart"))
+                                   ) # end sidebar layout
+                                 ), # end tab panel
                         tabPanel("Data Table", dataTableOutput("wt_table_test"))
-                                 
+                             
                       ) # end tabset panel
              ), # end tab panel
              
[…]

### 2023-03-16 solar_fam_shiny/app.R
expanding input boxes so text is in one line

@@ -49,7 +49,7 @@ ui <- fluidPage(
                         h5("Background"),
                         h5("Model Flow Chart"),  ### Add model image here
                         p("Visualization of methodology to calculate investment favorability score for each metropolitan area. Seven criteria comprise the total score, with each weighted by a value given preferences."),
-                        div(img(src = "model_no_weights.jpg",  width = "800", height = "400"), style="text-align: center;")
+                        div(img(src = "model_no_weights.jpg",  width = "1000", height = "500"), style="text-align: center;")
                       ) ### end mainPanel
              ), # ## END of tab 1
              ### tab2
@@ -61,23 +61,23 @@ ui <- fluidPage(
                                    sidebarPanel(
                                      ### inputs for weights
                                      numericInput(inputId = "REWtInput", label = "Real Estate",
-                                                  value = 0,width = "100px", min = 0.0,max = 1.0,step = 0.05), # end RE weight input
+                                                  value = 0,width = "150px", min = 0.0,max = 1.0,step = 0.05), # end RE weight input
               
                                      numericInput(inputId = "LandlordWtInput", label = "Landlord Policy",
-                                                       value = 0, width = "100px", min = 0.0,max = 1.0,step = 0.05),
+                                                       value = 0, width = "150px", min = 0.0,max = 1.0,step = 0.05),
                                      
                                      numericInput(inputId = "ElectrictyWtInput", label = "Electricity",
-                                                          value = 0,width = "100px", min = 0.0, max = 1.0, step = 0.05),
+                                                          value = 0,width = "150px", min = 0.0, max = 1.0, step = 0.05),
                                      
                                      numericInput(inputId = "CO2WtInput", label = "CO2 Emissions",
-                                                  value = 0,width = "100px", min = 0.0, max = 1.0,step = 0.05),# end Co2 weight input
+                                                  value = 0,width = "150px", min = 0.0, max = 1.0,step = 0.05),# end Co2 weight input
                                      
                                      numericInput(inputId = "ClimateRiskWtInput", label = "Climate Risk",
-                                                  value = 0,width = "100px", min = 0.0, max = 1.0, step = 0.05), # end climate risk weight input
+                                                  value = 0,width = "150px", min = 0.0, max = 1.0, step = 0.05), # end climate risk weight input
[…]

### 2023-03-17 src/MSAtotal_scores.Rmd
patchwork bar charts

@@
 date: '2022-11-14'
-output: html_document
+output: 
+  html_document: 
+    fig_caption: yes
+editor_options: 
+  chunk_output_type: console
 ---
@@
   rename(financial_score = normalized_irr,
-         co2_e_score = inverted_n_total_co2,
+         co2_score = inverted_n_total_co2,
          health_impact_score = inverted_n_health_emissions_cost,
@@
   mutate(real_estate_score = as.numeric(real_estate_score),
-         landlord_score = as.numeric(landlord_score)) %>% 
-  mutate_if(is.numeric, ~round(., 5))
+         landlord_score = as.numeric(landlord_score))
 
@@
                  criteria_all$financial_score[i],
-                 criteria_all$co2_e_score[i],
+                 criteria_all$co2_score[i],
                 criteria_all$health_impact_score[i],
@@
 
-```{output to google sheet}
+```{r}
 
 criteria_all <- criteria_all %>% 
-  mutate_if(is.numeric, ~round(., 5))
+  mutate_if(is.numeric,
+            round,
+            digits = 5)
 
-weighted <- t(t(weighted_criteria)*criteria_weights_client)
+weighted <- criteria_all %>% 
+  t(t(criteria_all[4:10])*criteria_weights_students) %>% 
+  as.data.frame(row.names=criteria_all$cbsa)
+
+ calculate_wt_criteria <- function(data, weights){
+    wt_criteria <- data %>%
+      mutate(wt_real_estate = real_estate_score * weights[1],
+             wt_landlord_policy = landlord_score * weights[2], 
+             wt_solar_irr = financial_score * weights[3],
+            wt_co2 = co2_score * weights[4], 
+             wt_health_impacts = health_impact_score * weights[5],
+             wt_electricity = electricity_score * weights[6],
+             wt_climate_risk = climate_risk_score * weights[7]) %>% 
+      select(!c(ends_with("_score"))) %>%
+      rename("Real Estate"="wt_real_estate",
+             "Landlord Policy"="wt_landlord_policy",
+             "Electricity"="wt_electricity", 
+             "CO2 Emissions"="wt_co2",
+             "Climate Risk"="wt_climate_risk",
+             "Health Impacts"="wt_health_impacts", 
+             "Solar IRR"="wt_solar_irr") 
+ }
+
+ wt_client <- calculate_wt_criteria(criteria_all, client_weights)  %>%  
+  unite("msa_state", city_msa:state, sep = ", ")
 
@@
 ```{r client data viz}
-## read in weighted criteria sheet
-client_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_Owen") %>% 
-  unite("msa_state", city_msa:state, sep = ", ") 
- # mutate(test2 = rowSums(client_weighted_criteria[,c(3:9)]))
 
 # read in cbsa data with full names
-score_tidy <- client_weighted_criteria %>% 
-  pivot_longer(cols = c("WT_real_estate":"WT_climRisk"), # selecting criteria cols
+client_score_tidy <- wt_client %>% 
+  pivot_longer(cols = c(3:9), # selecting criteria cols
                names_to = "criteria",
-               values_to = "criteria_score") %>% 
+               values_to = "wt_criteria_score") %>% 
  filter(msa_state != "Portland, ME")  # removing portland (negative number)
@@
 tiff(filename = "wt_criteria_client_chart.tiff",width = 8, height = 6, units = "in", res = 300, pointsize = 12)
-ggplot(data = score_tidy, aes(y = reorder(msa_state, FINAL_owen_SCORE), x = criteria_score)) +
+p1 <- ggplot(data = client_score_tidy, aes(y = reorder(msa_state,  wt_criteria_score, FUN=sum), x = wt_criteria_score)) +
     geom_col(aes(fill = criteria)) +
@@
        y = " ",
-       title = "Client") +
+       title = "") +
   theme_bw() +
   guides(fill=guide_legend(title="Criteria (weighted)")) +
-   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD"),
-                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Solar IRR", "Real Estate", "Landlord Policy")) +
-  scale_x_continuous(limits = c(0, 0.7), expand = c(0,0), breaks=seq(0, 0.7, 0.1))
+   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
+      breaks = c("Climate Risk", "CO2 Emissions", "Electricity", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
+        scale_x_continuous(limits = c(0, 1.0), expand = c(0,0), breaks=seq(0, 01.0, 0.1)) +
+  theme(legend.position = "none")
 dev.off()
@@
 ## read in weighted criteria sheet
-student_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_stud") %>% 
+student_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_stud")
+
+equity_wt <- calculate_wt_criteria(criteria_all, criteria_weights_students) %>% 
   unite("msa_state", city_msa:state, sep = ", ")
@@
 # read in cbsa data with full names
-score_tidy2 <- student_weighted_criteria %>% 
-  pivot_longer(cols = c("WT_real_estate":"WT_climRisk"), # selecting criteria cols
+score_tidy2 <- equity_wt %>% 
+  pivot_longer(cols = c(3:9), # selecting criteria cols
                names_to = "criteria",
-               values_to = "criteria_score")
+               values_to = "wt_criteria_score")
 
 tiff(filename = "wt_criteria_student_chart.tiff", width = 8, height = 6, units = "in", res = 300)
-ggplot(data = score_tidy2, aes(y = reorder(msa_state, FINAL_st_score), x = criteria_score)) + 
+
+p2 <- ggplot(data = score_tidy2, aes(y = reorder(msa_state, wt_criteria_score, FUN=sum), x = wt_criteria_score)) + 
     geom_col(aes(fill = criteria)) +
@@
        y = " ",
-       title = "Student") +
+       title = "") +
   theme_bw() +
   guides(fill=guide_legend(title="Criteria (weighted)")) +
-   scale_fill_manual(values = c("palevioletred1", "orange",  "gold", "#44AA99", "dodgerblue4", "#6495ED"),
-                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Solar IRR", "Real Estate", "Health Impacts")) +
-  scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1))
+   scale_fill_manual(values = c("palevioletred1", "orange",  "gold", "#44AA99", "dodgerblue4", "#6495ED", "#BC80BD"),
+                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Solar IRR", "Real Estate", "Health Impacts", "Landlord Policy")) +
+  scale_x_continuous(limits = c(0, 1.0), expand = c(0,0), breaks=seq(0, 1.0, 0.1))
+
 dev.off()
 ```
+
+```{r}
+library(patchwork)
+
+p1 <- ggplot(data = client_score_tidy, aes(y = reorder(msa_state,  wt_criteria_score, FUN=sum), x = wt_criteria_score)) +
+    geom_col(aes(fill = criteria)) +
+  labs(x = "Total Score",
+       y = " ",
+       title = "") +
+  theme_bw() +
+  guides(fill=guide_legend(title="Criteria (weighted)")) +
+   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
+      breaks = c("Climate Risk", "CO2 Emissions", "Electricity", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
+        scale_x_continuous(limits = c(0, 1.0), expand = c(0,0), breaks=seq(0, 01.0, 0.1)) +
+  theme(legend.position = "none")
+
+p2 <- ggplot(data = score_tidy2, aes(y = reorder(msa_state, wt_criteria_score, FUN=sum), x = wt_criteria_score)) + 
+    geom_col(aes(fill = criteria)) +
+  labs(x = "Total Score",
+       y = " ",
+       title = "") +
+  theme_bw() +
+  guides(fill=guide_legend(title="Criteria (weighted)")) +
+   scale_fill_manual(values = c("palevioletred1", "orange",  "gold", "#44AA99", "dodgerblue4", "#6495ED", "#BC80BD"),
+                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Solar IRR", "Real Estate", "Health Impacts", "Landlord Policy")) +
+  scale_x_continuous(limits = c(0, 1.0), expand = c(0,0), breaks=seq(0, 1.0, 0.1))
+
+tiff(filename = "data_vz.tiff", width=12, units = "in", res=100)
+p3<- p1+p2 + plot_layout(widths = c(6, 6)) +
+  plot_annotation(tag_levels = 'A',
+                  caption = "Figure 1.
+Ranked metropolitan areas (n=29) based on investment favorability for rooftop solar PV on mutlifamily houses.
+Areas were scored using six criteria, each given a respective weight given stakeholder priorities.
+\n Each criteria is comprised of indicator metrics using data from the U.S.
+Census Bureau and NREL's reOpt tool.
+Graph A represents a real estate investor stakeholder who prioritizes climate risk due to long-term investments.
+\n B represents a equity-focused stakeholder who prioritizes markets that reduce CO2 emissions and associated health impacts.")
+dev.off()
+
+```
+
+"Figure 1.
+Ranked metropolitan areas (n=29) based on investment favorability for rooftop solar PV on mutlifamily houses.
+Areas were scored using six criteria, each given a respective weight given stakeholder priorities.
+Each criteria is comprised of indicator metrics using data from the U.S.
+Census Bureau and NREL's reOpt tool.
+Graph A represents a real estate investor stakeholder who prioritizes climate risk due to long-term investments.
+B represents a equity-focused stakeholder who values impact one

### 2023-04-19 src/MSAtotal_scores.Rmd
updated weighting function to calculate sum of the 7 criteria

@@
 date: '2022-11-14'
-output: 
-  html_document: 
-    fig_caption: yes
+output: pdf_document
 editor_options: 
@@
 
+### Cleaning up the data order and column names
 # reorder columns
@@
 
+# rename columns
 criteria_all <- criteria_all[, col_order] %>% 
@@
 
-weighted <- criteria_all %>% 
-  t(t(criteria_all[4:10])*criteria_weights_students) %>% 
-  as.data.frame(row.names=criteria_all$cbsa)
-
+# creating a function with criteria data and weight inputs
 calculate_wt_criteria <- function(data, weights){
     wt_criteria <- data %>%
+      # multiply each criteria by respective weight
       mutate(wt_real_estate = real_estate_score * weights[1],
@@
              wt_climate_risk = climate_risk_score * weights[7]) %>% 
+      # select weighted columns and rename them
       select(!c(ends_with("_score"))) %>%
+      # create column for investment score
+      mutate(investment_score = 0) %>%
+      # for loop to sum weighted columns for each row
+      mutate(investment_score = apply(.[, 4:10], 1, sum)) %>% 
       rename("Real Estate"="wt_real_estate",
              "Landlord Policy"="wt_landlord_policy",
-             "Electricity"="wt_electricity", 
-             "CO2 Emissions"="wt_co2",
-             "Climate Risk"="wt_climate_risk",
+             "Electricity Production"="wt_electricity", 
+             "CO2 Emission Reduction"="wt_co2",
+             "Climate Risk Avoided"="wt_climate_risk",
              "Health Impacts"="wt_health_impacts", 
@@
  }
-
- wt_client <- calculate_wt_criteria(criteria_all, client_weights)  %>%  
-  unite("msa_state", city_msa:state, sep = ", ")
 
@@
 
-```{r client data viz}
+```{r tidy data}
+# client weights
+wt_criteria_client <- calculate_wt_criteria(criteria_all, client_weights)  %>%  
+  unite("msa_state", city_msa:state, sep = ", ")
 
-# read in cbsa data with full names
-client_score_tidy <- wt_client %>% 
+client_score_tidy <- wt_criteria_client  %>% 
   pivot_longer(cols = c(3:9), # selecting criteria cols
@@
 
-tiff(filename = "wt_criteria_client_chart.tiff",width = 8, height = 6, units = "in", res = 300, pointsize = 12)
+# equity weights
+wt_criteria_equity <- calculate_wt_criteria(criteria_all, criteria_weights_students) %>% 
+  unite("msa_state", city_msa:state, sep = ", ")
+
+# tidy data for bar chart
+equity_tidy <- wt_criteria_equity %>% 
+  pivot_longer(cols = c(3:9), # selecting criteria cols
+               names_to = "criteria",
+               values_to = "wt_criteria_score") 
+```
+
+```{r client data viz}
+tiff(filename = "client_criteria_barchart.tiff",width = 8, height = 6, units = "in", res = 300, pointsize = 12)
+ggplot(data = client_score_tidy, aes(y = reorder(msa_state,  wt_criteria_score, FUN=sum), x = wt_criteria_score)) +
+    geom_col(aes(fill = factor(criteria, levels = c("Climate Risk Avoided", "CO2 Emission Reduction", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")))) +
+  theme_bw() +
+  labs(x = "Total Score",
+       y = " ",
+       title = "") +
+  guides(fill=guide_legend(title="Criteria (weighted)")) +
+   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
+      breaks = c("Climate Risk Avoided", "CO2 Emission Reduction", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
+        scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1)) 
+#+ theme(legend.position = "none")
+dev.off()
+```
+
+```{r student data viz}
+# ## read in weighted criteria sheet
+# student_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_stud")
+
+#tiff(filename = "wt_criteria_student_chart.tiff", width = 8, height = 6, units = "in", res = 300)
+ ggplot(data = equity_tidy, aes(y = reorder(msa_state, wt_criteria_score, FUN=sum), x = wt_criteria_score)) + 
+    geom_col(aes(fill = criteria)) +
+  labs(x = "Total Score",
+       y = " ",
+       title = "") +
+  theme_bw() +
+  guides(fill=guide_legend(title="Criteria (weighted)")) +
+   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
+      breaks = c("Climate Risk Avoidance", "CO2 Abatement", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
+  scale_x_continuous(limits = c(0, 1.0), expand = c(0,0), breaks=seq(0, 1.0, 0.1))
+
+#dev.off()
+```
+
+```{r}
+library(patchwork)
+
+# combined visual of client and equity score results
 p1 <- ggplot(data = client_score_tidy, aes(y = reorder(msa_state,  wt_criteria_score, FUN=sum), x = wt_criteria_score)) +
@@
    scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
-      breaks = c("Climate Risk", "CO2 Emissions", "Electricity", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
-        scale_x_continuous(limits = c(0, 1.0), expand = c(0,0), breaks=seq(0, 01.0, 0.1)) +
-  theme(legend.position = "none")
-dev.off()
-```
+      breaks = c("Climate Risk Avoidance", "CO2 Abatement", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
+        scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1)) +
+  theme(legend.position = "none") +
+  theme(axis.title.y = element_text(size = 20))
 
-```{r student data viz}
-## read in weighted criteria sheet
-student_weighted_criteria <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=0", sheet = "wtd_criteria_stud")
-
-equity_wt <- calculate_wt_criteria(criteria_all, criteria_weights_students) %>% 
-  unite("msa_state", city_msa:state, sep = ", ")
-
-# read in cbsa data with full names
-score_tidy2 <- equity_wt %>% 
-  pivot_longer(cols = c(3:9), # selecting criteria cols
-               names_to = "criteria",
-               values_to = "wt_criteria_score")
-
-tiff(filename = "wt_criteria_student_chart.tiff", width = 8, height = 6, units = "in", res = 300)
-
-p2 <- ggplot(data = score_tidy2, aes(y = reorder(msa_state, wt_criteria_score, FUN=sum), x = wt_criteria_score)) + 
+p2 <- ggplot(data = equity_tidy, aes(y = reorder(msa_state, wt_criteria_score, FUN=sum), x = wt_criteria_score)) + 
     geom_col(aes(fill = criteria)) +
@@
   theme_bw() +
-  guides(fill=guide_legend(title="Criteria (weighted)")) +
-   scale_fill_manual(values = c("palevioletred1", "orange",  "gold", "#44AA99", "dodgerblue4", "#6495ED", "#BC80BD"),
-                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Solar IRR", "Real Estate", "Health Impacts", "Landlord Policy")) +
-  scale_x_continuous(limits = c(0, 1.0), expand = c(0,0), breaks=seq(0, 1.0, 0.1))
-
-dev.off()
-```
-
-```{r}
-library(patchwork)
-
-p1 <- ggplot(data = client_score_tidy, aes(y = reorder(msa_state,  wt_criteria_score, FUN=sum), x = wt_criteria_score)) +
-    geom_col(aes(fill = criteria)) +
-  labs(x = "Total Score",
-       y = " ",
-       title = "") +
-  theme_bw() +
+  guides(shape = guide_legend(override.aes = list(size = 0.5))) +
   guides(fill=guide_legend(title="Criteria (weighted)")) +
    scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
-      breaks = c("Climate Risk", "CO2 Emissions", "Electricity", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
-        scale_x_continuous(limits = c(0, 1.0), expand = c(0,0), breaks=seq(0, 01.0, 0.1)) +
-  theme(legend.position = "none")
+      breaks = c("Climate Risk Avoidance", "CO2 Abatement", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
+        scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1)) +
+    theme(axis.title.y = element_text(size = 11))
 
-p2 <- ggplot(data = score_tidy2, aes(y = reorder(msa_state, wt_criteria_score, FUN=sum), x = wt_criteria_score)) + 
-    geom_col(aes(fill = criteria)) +
-  labs(x = "Total Score",
-       y = " ",
-       title = "") +
-  theme_bw() +
-  guides(fill=guide_legend(title="Criteria (weighted)")) +
-   scale_fill_manual(values = c("palevioletred1", "orange",  "gold", "#44AA99", "dodgerblue4", "#6495ED", "#BC80BD"),
-                     labels=c('Climate Risk', 'CO2 Emissions', 'Electricity', "Solar IRR", "Real Estate", "Health Impacts", "Landlord Policy")) +
-  scale_x_continuous(limits = c(0, 1.0), expand = c(0,0), breaks=seq(0, 1.0, 0.1))
+# tiff(filename = "data_vz.tiff", width=12, units = "in", res=100)
+ p1/p2 + 
+  plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 16)) 
+ ggsave("test_dim_2plots.jpeg", plot = plots, width = 6, height = 4, units = "in", dpi = 300)
 
-tiff(filename = "data_vz.tiff", width=12, units = "in", res=100)
-p3<- p1+p2 + plot_layout(widths = c(6, 6)) +
-  plot_annotation(tag_levels = 'A',
-                  caption = "Figure 1.
-Ranked metropolitan areas (n=29) based on investment favorability for rooftop solar PV on mutlifamily houses.
-Areas were scored using six criteria, each given a respective weight given stakeholder priorities.
-\n Each criteria is comprised of indicator metrics using data from the U.S.
-Census Bureau and NREL's reOpt tool.
-Graph A represents a real estate investor stakeholder who prioritizes climate risk due to long-term investments.
-\n B represents a equity-focused stakeholder who prioritizes markets that reduce CO2 emissions and associated health impacts.")
-dev.off()
+ #rstudioapi::savePlotAsImage(file = "myplot.png",width = dev.size(units="px")[[1]],height=dev.size(units="px")[[2]])
 
@@
 
-"Figure 1.
+Figure 1.
 Ranked metropolitan areas (n=29) based on investment favorability for rooftop solar PV on mutlifamily houses.
@@
 Graph A represents a real estate investor stakeholder who prioritizes climate risk due to long-term investments.
-B represents a equity-focused stakeholder who values impact one
+B represents a equity-focused stakeholder who values impacts on the community, emissions, and

### 2023-04-20 src/MSAtotal_scores.Rmd
github check
cleaning up code

@@
 date: '2022-11-14'
-output: pdf_document
-editor_options: 
-  chunk_output_type: console
+output: 
+  html_document:
+    code_folding: hide
 ---
 
-```{r setup, include=FALSE, warning=FALSE, message=FALSE}
+```{r setup, include=TRUE, warning=FALSE, message=FALSE}
 knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
@@
 ```{r}
-# read in the data
+# read in the normalized data
 real_estate_score_data <- read_sheet("https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit#gid=1038418521") %>% 
@@
 
-```{r read in data}
+```{r data wrangling}
 criteria_all <- real_estate_score_data %>% 
@@
   mutate(real_estate_score = as.numeric(real_estate_score),
-         landlord_score = as.numeric(landlord_score))
+         landlord_score = as.numeric(landlord_score)) %>%
+  # round criteria scores
+  mutate_if(is.numeric,round, digits = 5)
 
@@
 
-```{r FINAL student weights}
-# weight criteria for students
-criteria_weights_students <- c(0.18, 0.0, 0.10, 0.19, 0.17, 0.14, 0.22)
+```{r model weights}
+# weights for each criteria must be in this order
+equity_weights <- c(0.18, # real estate weight
+                    0.0, # landlord weight
+                    0.10, # financial weight
+                    0.19, # co2 emissions avoided weight
+                    0.17, # health impacts weight
+                    0.14, # electricity weight
+                    0.22) # climate risk weight
 
-#make landlord_score column
-criteria_all <- criteria_all %>% 
-  mutate(equity_scenario_score = cbsa,
-         client_FINAL_score = cbsa) #this is just a place holder
-
-for (i in 1:nrow(criteria_all)) {
-  cbsa_data <- c(criteria_all$real_estate_score[i],
-                criteria_all$landlord_score[i],
-                 criteria_all$financial_score[i],
-                 criteria_all$co2_score[i],
-                criteria_all$health_impact_score[i],
-                criteria_all$electricity_score[i],
-                criteria_all$climate_risk_score[i]
-                  )
-  criteria_all$equity_scenario_score[i] = weighted.mean(cbsa_data, criteria_weights_students)
-}
-
+client_weights <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)
 ```
 
-```{r Client FINAL WEIGHTS}
-
-client_weights <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)
-
-# multiply column by respective weight and sum for total
-
-for (i in 1:nrow(criteria_all)) {
-  cbsa_data <- c(criteria_all$real_estate_score[i],
-                criteria_all$landlord_score[i],
-                 criteria_all$financial_score[i],
-                 criteria_all$co2_score[i],
-                criteria_all$health_impact_score[i],
-                criteria_all$electricity_score[i],
-                criteria_all$climate_risk_score[i]
-                  )
-  criteria_all$client_FINAL_score[i] = weighted.mean(cbsa_data, client_weights) 
-}
-
-```
+# error check so weights sum to 1
+# incorporate shiny??
 
 ```{r}
-
-criteria_all <- criteria_all %>% 
-  mutate_if(is.numeric,
-            round,
-            digits = 5)
-
 # creating a function with criteria data and weight inputs
@@
              "Health Impacts"="wt_health_impacts", 
-             "Solar IRR"="wt_solar_irr") 
+             "Solar IRR"="wt_solar_irr") %>%  
+      unite("msa_state", city_msa:state, sep = ", ")
  }
@@
 # equity weights
-wt_criteria_equity <- calculate_wt_criteria(criteria_all, criteria_weights_students) %>% 
+wt_criteria_equity <- calculate_wt_criteria(criteria_all, equity_weights) %>% 
   unite("msa_state", city_msa:state, sep = ", ")

### 2023-04-20 src/demo_data_wrangling.Rmd
cleaning up code

@@
 ### Population Growth
+
 All states were considered with the exception of CA, HI, and AK.
 Initial analysis of Metropolitan Statistical Areas (MSAs) relied on Census data of annual population estimates, in which the population had to be greater than 500,000 individuals for the purpose of attracting investors.
-Population change from 2020 to 2021 was calculated by $$\frac{2021 estimate- 2020 estimate}{2020 estimate} * 100$$
-MSAs with a population growth > 0.5%, were selected to gather further data on. The 0.5% growth was determined.
+Population change from 2020 to 2021 was calculated by $$\frac{2021 estimate- 2020 estimate}{2020 estimate} * 100$$ MSAs with a population growth \> 0.5%, were selected to gather further data on.
+The 0.5% growth was determined.
 
@@
 Population growth in Metropolitan Statistical Areas (n=29) in 2020-2021.
-Highlighted areas have a population > 300,000 individuals and a population growth rate greater than 0.5%.
+Highlighted areas have a population \> 300,000 individuals and a population growth rate greater than 0.5%.
 The colors of the MSAs represent incremental growth rates, in which low(green) is 0.5-1% increase, medium(blue) is 1-2% increase, and high(pink) is 2-3.5% increase.
@@
 ### Household Income
+
 Household Median Income data was also obtained from Census data at the MSA level and amended to existing MSAs with a positive population increase.
@@
 ### Employment Rates
+
 MSA monthly unemployment rates (seasonally-adjusted) were provided by The Bureau of Labor statistics in the civil labor force data at the MSA level.
@@
 The employment data for the month of January in years 2018-2021 was selected to calculate the change in employment rates over time.
-Year-over-Year employment growth was calculated by dividing the employment rate of the year by the previous year and subtracting one; for example, 2018 employment YoY growth = (2019 employment rate / 2018 employment rate) - 1.
+Year-over-Year employment growth was calculated by dividing the employment rate of the year by the previous year and subtracting one; for example, 2018 employment growth = (2019 employment rate / 2018 employment rate) - 1.
 
@@
 
-<https://www.bls.gov/web/metro.supp.toc.htm> <https://www.bls.gov/web/metro/metro_oty_change.htm> - significant change
-
-2021 Median Rent from https://www.huduser.gov/portal/datasets/50per.html#2021
+2021 Median Rent from <https://www.huduser.gov/portal/datasets/50per.html#2021>
 
@@
 
-Median housing prices by MSA were obtained from the Association of Realtors for 2021. 
-Copyright ©[2021] NATIONAL ASSOCIATION OF REALTORS®. All rights reserved.  Reprinted with permission.
+Median housing prices by MSA were obtained from the Association of Realtors for 2021. Copyright ©[2021] NATIONAL ASSOCIATION OF REALTORS®. All rights reserved. Reprinted with permission.
 
@@
 ### City Data Table
+
 MSAs were split up by cities to collect utility, rent, and landlord data.

### 2023-04-26 src/reopt_api.Rmd
testing out reopt API

---
title: "Reopt"
author: "Grace Bianchi"
date: "2023-04-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(RJSONIO)
library(jsonlite)
```

```{r}
# Set your API key
api_key <- "GZtZAUcX97VchjrLAeeN3amg2pslaO3hbtCQgXBl"

# Set the URL for the job submission endpoint
url <- "https://developer.nrel.gov/api/reopt/stable/job"

job_url <- paste0(url, "?api_key=", api_key)
```

```{r}
# create list of default inputs for each city
inputs <- list(
  ModuleType = "Premium",
  DCACratio = 1.35,
  System_losses = 5,
  FederalIncentive = 30
)
```

```{r}

# Set the payload for the job submission request
payload <- list(
  "Scenario" = list(
    "Site" = list(
      "longitude" = -118.1164613,
      "latitude" = 34.5794343,
      "roof_squarefeet" = 5000.0,
      "land_acres" = 1.0,
      "PV" = list(
        "installed_cost_us_dollars_per_kw" = 2000.0,
        "tilt" = 34.579,
        "degradation_pct" = 0.005,
        "macrs_option_years" = 5,
        "federal_itc_pct" = 0.3,
        "module_type" = 0,
        "array_type" = 1,
        "om_cost_us_dollars_per_kw" = 16.0,
        "macrs_itc_reduction" = 0.5,
        "azimuth" = 180.0,
        "federal_rebate_us_dollars_per_kw" = 350.0
      ),
      "LoadProfile" = list(
        "doe_reference_name" = "MidriseApartment",
        "annual_kwh" = 10000000.0,
        "city" = "LosAngeles"
      ),
      "Storage" = list(
        "total_rebate_per_kw" = 100.0,
        "macrs_option_years" = 5,
        "can_grid_charge" = TRUE,
        "macrs_bonus_pct" = 0.4,
        "macrs_itc_reduction" = 0.5,
        "total_itc_pct" = 0,
        "installed_cost_us_dollars_per_kw" = 1000.0,
        "installed_cost_us_dollars_per_kwh" = 500.0,
        "replace_cost_us_dollars_per_kw" = 460.0,
        "replace_cost_us_dollars_per_kwh" = 230.0
      ),
      "ElectricTariff" = list(
        "urdb_label" = "5ed6c1a15457a3367add15ae"
      ),
      "Financial" = list(
        "escalation_pct" = 0.026,
        "offtaker_discount_pct" = 0.081,
        "owner_discount_pct" = 0.081,
        "analysis_years" = 20,
        "offtaker_tax_pct" = 0.4,
        "owner_tax_pct" = 0.4,
        "om_cost_escalation_pct" = 0.025
      )
    )
  )
)

```

```{r}
# Make the request to submit the job
response <- POST(url, query = list(api_key = api_key), body = payload, encode = "json", timeout(30))

# Check the response status code
status_code <- response$status_code
if (status_code == 200) {
  message("Inputs updated successfully.")
} else {
  message("Error updating inputs.")
}

# get run_uuid to obtain outputs
output <- httr::content(response)
run_uuid <- output$run_uuid
```

```{r}
results_url <- paste0("https://developer.nrel.gov/api/reopt/stable/job/", run_uuid, "/results")

# Send the GET request to retrieve the results
response <- GET(url = results_url, query = list(api_key = api_key), timeout(30))

# Extract the response content as JSON and convert it to a list
response_content <- content(response, as="text")
response_json <- fromJSON(response_content)

# pull out IMPORTANT data
output_df <- data.frame(
  IRR = response_json$outputs$Scenario$Site$Financial$irr_pct,
  Health_Emissions_Cost = response_json$outputs$Scenario$Site$lifecycle_emissions_cost_Health_bau,
  Annual_Generation_Renewables = response_json$outputs$Scenario$Site$annual_renewable_electricity_pct)

```

```{r}
host <- "https://developer.nrel.gov/api/reopt"
api_key <- "?API_KEY=GZtZAUcX97VchjrLAeeN3amg2pslaO3hbtCQgXBl"
version_url <- file.path(host, help_end, api)
api_version <- jsonlite::fromJSON(version_url) # retrieved

```

```{r}

## energy goals
#cost_savings and clean_energy
## inputs not default 
host_eff <- 0
scc = 185

#PV & wind space available:  ❌ Land only ✅ Roofspace only ❌ Land & roofspace

#PV System Characteristics
ModuleType= Premium
DCACratio = 1.35
System_losses = 5
FederalIncentive= 30

```

```{r}
# test scenario post
test <- "https://reopt.nrel.gov/tool/results/f4e6c9db-3c22-43a4-b2bf-15146aec9896"

```

### 2023-04-26 src/reopt_api.Rmd
trying to read in utility data

@@
 inputs <- list(
-  ModuleType = "Premium",
-  DCACratio = 1.35,
-  System_losses = 5,
-  FederalIncentive = 30
+  offtaker_tax_pct = 0, # host effective tax rate
+  co2_cost_us_dollars_per_tonne = 185,
+  module_type = 1,
+  dc_ac_ratio = 1.35,
+  losses = 0.05,
+  federal_itc_pct = 0.3,
+  location = "roof", # restrict projects to roof
+  can_net_meter = TRUE
+
 )
@@
 ```{r}
-
 # Set the payload for the job submission request
@@
         "installed_cost_us_dollars_per_kw" = 2000.0,
-        "tilt" = 34.579,
-        "degradation_pct" = 0.005,
-        "macrs_option_years" = 5,
+        "federal_rebate_us_dollars_per_kw" = 350.0,
+        # changes from default
+        "module_type" = 1,
+        "dc_ac_ratio" = 1.35,
+        "losses" = 0.05,
         "federal_itc_pct" = 0.3,
-        "module_type" = 0,
-        "array_type" = 1,
-        "om_cost_us_dollars_per_kw" = 16.0,
-        "macrs_itc_reduction" = 0.5,
-        "azimuth" = 180.0,
-        "federal_rebate_us_dollars_per_kw" = 350.0
+        "location" = "roof", # restrict projects to roof
+        "can_net_meter" = TRUE
       ),
@@
       ),
-      "Storage" = list(
-        "total_rebate_per_kw" = 100.0,
-        "macrs_option_years" = 5,
-        "can_grid_charge" = TRUE,
-        "macrs_bonus_pct" = 0.4,
-        "macrs_itc_reduction" = 0.5,
-        "total_itc_pct" = 0,
-        "installed_cost_us_dollars_per_kw" = 1000.0,
-        "installed_cost_us_dollars_per_kwh" = 500.0,
-        "replace_cost_us_dollars_per_kw" = 460.0,
-        "replace_cost_us_dollars_per_kwh" = 230.0
-      ),
       "ElectricTariff" = list(
-        "urdb_label" = "5ed6c1a15457a3367add15ae"
+        "urdb_label" = "5ed6c1a15457a3367add15ae",
+        "add_tou_energy_rates_to_urdb_rate" = TRUE
       ),
@@
 )
+
+```
+
+```{r}
+## open EI api
+openEI_api_key <- "rU6MEqhK9OJ7obAnmcbzYhpFOenfBM5Q8GF0yJFA"
+
+openEI_start <- "https://api.openei.org/utility_rates?version=latest&format=json&sector=Residential&detail=full&approved=true&address=26900&"
+# Set the URL for the job submission endpoint
+
+openEI_url <- paste0(openEI_start, "api_key=", openEI_api_key)
+
+EIresponse <- GET(openEI_url)
+content <- content(EIresponse, "text")
+json <- fromJSON(content)
+
+EIoutput_df <- data.frame(
+  label = json$items$label,
+  utility = json$items$utility,
+  approved = json$items$approved,
+  rate_name = json$items$name)
+
+  # extract utility rate information
+rates <- json$items$rate
+  # print utility rate information for the city
+  cat(city, "utility rates:\n")
+  for (rate in rates) {
+    cat("  - Name:", rate$name, "\n")
+    cat("    Description:", rate$description, "\n")
+    cat("    Value:", rate$value, "\n")
+
+if (http_type(response)$status_code == 200) {
+  data <- fromJSON(content(EIresponse, "text"), simplifyDataFrame = TRUE)
+
+  # Extract relevant information from the data
+  utility_name <- data$items[[1]]$utility_name
+  utility_type <- data$items[[1]]$sector
+  rate_name <- data$items[[1]]$name
+  start_date <- data$items[[1]]$start_date
+  end_date <- data$items[[1]]$end_date
+  rate_structure <- data$items[[1]]$rate_structure
+  energy_rate <- data$items[[1]]$rate
+  demand_rate <- data$items[[1]]$demand_rate
+
+  # Print the results
+  cat(sprintf("Utility name: %s\n", utility_name))
+  cat(sprintf("Utility type: %s\n", utility_type))
+  cat(sprintf("Rate name: %s\n", rate_name))
+  cat(sprintf("Start date: %s\n", start_date))
+  cat(sprintf("End date: %s\n", end_date))
+  cat(sprintf("Rate structure: %s\n", rate_structure))
+  cat(sprintf("Energy rate: %s\n", energy_rate))
+  cat(sprintf("Demand rate: %s\n", demand_rate))
+} else {
+  cat(sprintf("Failed to retrieve data (status code %s)", http_type(response)$status_code))
+}
+
+```
+
+```{r}
+
+# list of cities
+cities <- c("San Francisco", "Los Angeles", "New York")
+
+# loop through each city and get utility rate information
+for (city in cities) {
+  # construct URL with city name as parameter
+  url <- paste0("https://api.openei.org/utility_rates?version=3&format=json&address=", gsub(" ", "+", city))
+  # make API call
+  response <- GET(url)
+  # extract content and parse as JSON
+  content <- content(response, "text")
+  json <- fromJSON(content)
+  # extract utility rate information
+  rates <- json$items$rate
+  # print utility rate information for the city
+  cat(city, "utility rates:\n")
+  for (rate in rates) {
+    cat("  - Name:", rate$name, "\n")
+    cat("    Description:", rate$description, "\n")
+    cat("    Value:", rate$value, "\n")
+  }
+}
+```
+
+```{r STUFF TO WORK ON}
+
+net_metering_limit_kw # by state
 
@@
 ```{r}
+## GET RESULTS FROM REOPT
 results_url <- paste0("https://developer.nrel.gov/api/reopt/stable/job/", run_uuid, "/results")
@@
 
-## energy goals
-#cost_savings and clean_energy
-## inputs not default 
-host_eff <- 0
-scc = 185
-
 #PV & wind space available:  ❌ Land only ✅ Roofspace only ❌ Land & roofspace

### 2023-04-26 src/reopt_api.Rmd
created function to use list of zipcodes and collect utility data from OpenEI website api

@@
 ```{r}
+```
+
+```{r}
+
+## open EI api
 ## open EI api
@@
 
-openEI_start <- "https://api.openei.org/utility_rates?version=latest&format=json&sector=Residential&detail=full&approved=true&address=26900&"
+openEI_start <- "https://api.openei.org/utility_rates?version=latest&format=json&sector=Residential&detail=full&approved=true&"
 # Set the URL for the job submission endpoint
 
-openEI_url <- paste0(openEI_start, "api_key=", openEI_api_key)
-
+openEI_url <- paste0(openEI_start, "address=", zipcodes, "&api_key=", openEI_api_key)
 EIresponse <- GET(openEI_url)
@@
 
-EIoutput_df <- data.frame(
+ # extract utility rate information
+EIutility_df <- list(
   label = json$items$label,
   utility = json$items$utility,
-  approved = json$items$approved,
   rate_name = json$items$name)
+```
 
-  # extract utility rate information
-rates <- json$items$rate
-  # print utility rate information for the city
-  cat(city, "utility rates:\n")
-  for (rate in rates) {
-    cat("  - Name:", rate$name, "\n")
-    cat("    Description:", rate$description, "\n")
-    cat("    Value:", rate$value, "\n")
+```{r}
 
-if (http_type(response)$status_code == 200) {
-  data <- fromJSON(content(EIresponse, "text"), simplifyDataFrame = TRUE)
+openEI_start_url <- "https://api.openei.org/utility_rates?version=latest&format=json&sector=Residential&detail=full&approved=true&address="
+openEI_api_end <- "&api_key=rU6MEqhK9OJ7obAnmcbzYhpFOenfBM5Q8GF0yJFA"
 
-  # Extract relevant information from the data
-  utility_name <- data$items[[1]]$utility_name
-  utility_type <- data$items[[1]]$sector
-  rate_name <- data$items[[1]]$name
-  start_date <- data$items[[1]]$start_date
-  end_date <- data$items[[1]]$end_date
-  rate_structure <- data$items[[1]]$rate_structure
-  energy_rate <- data$items[[1]]$rate
-  demand_rate <- data$items[[1]]$demand_rate
+### Function w for loop
+get_utility_data <- function(zipcodes) {
+  # Create an empty list to store the resulting data frames
+  utility_info <- list()
+  # get data for each zipcode in vector
+  for (zipcode in zipcodes) {
+    openEI_start_url <- "https://api.openei.org/utility_rates?version=latest&format=json&sector=Residential&detail=full&approved=true&address="
+    openEI_api_end <- "&api_key=rU6MEqhK9OJ7obAnmcbzYhpFOenfBM5Q8GF0yJFA"
 
-  # Print the results
-  cat(sprintf("Utility name: %s\n", utility_name))
-  cat(sprintf("Utility type: %s\n", utility_type))
-  cat(sprintf("Rate name: %s\n", rate_name))
-  cat(sprintf("Start date: %s\n", start_date))
-  cat(sprintf("End date: %s\n", end_date))
-  cat(sprintf("Rate structure: %s\n", rate_structure))
-  cat(sprintf("Energy rate: %s\n", energy_rate))
-  cat(sprintf("Demand rate: %s\n", demand_rate))
-} else {
-  cat(sprintf("Failed to retrieve data (status code %s)", http_type(response)$status_code))
+    openEI_url <- paste0(openEI_start_url, zipcode, openEI_api_end)
+    response <- GET(openEI_url)
+    # extract content and parse as JSON
+    content <- content(response, "text")
+    json <- fromJSON(content)
+
+    # create a data frame with the extracted utility info
+    utility_df <- data.frame(
+      "zipcode" = rep(zipcode, each = nrow(json$items$label)),
+      "label" = json$items$label,
+      "utility" = json$items$utility,
+      "rate_name" = json$items$name
+    )
+
+    # append the resulting data frame to the list
+    utility_info[[length(utility_info) + 1]] <- utility_df
   }
+
+  # combine the resulting data frames into a single data frame
+  combined_utility_info <- do.call(rbind, utility_info)
+
+  return(combined_utility_info)
+}
+
+utility_info_zips <- get_utility_data(zipcodes = zipcodes)
 
@@
 ```{r}
+zipcodes <- c("46033", "93105", "90404")
+## functiion wiith out for loop
+get_zip_utility <- function(zipcode){
+  openEI_start_url <- "https://api.openei.org/utility_rates?version=latest&format=json&sector=Residential&detail=full&approved=true&address="
+openEI_api_end <- "&api_key=rU6MEqhK9OJ7obAnmcbzYhpFOenfBM5Q8GF0yJFA"
+zipcode=zipcode
 
-# list of cities
-cities <- c("San Francisco", "Los Angeles", "New York")
+  openEI_url <- paste0(openEI_start_url, zipcode, openEI_api_end)
+  response <- GET(openEI_url)
+  # extract content and parse as JSON
+  content <- content(EIresponse, "text")
+  json <- fromJSON(content)
 
-# loop through each city and get utility rate information
-for (city in cities) {
-  # construct URL with city name as parameter
-  url <- paste0("https://api.openei.org/utility_rates?version=3&format=json&address=", gsub(" ", "+", city))
-  # make API call
-  response <- GET(url)
-  # extract content and parse as JSON
-  content <- content(response, "text")
-  json <- fromJSON(content)
-  # extract utility rate information
-  rates <- json$items$rate
-  # print utility rate information for the city
-  cat(city, "utility rates:\n")
-  for (rate in rates) {
-    cat("  - Name:", rate$name, "\n")
-    cat("    Description:", rate$description, "\n")
-    cat("    Value:", rate$value, "\n")
+  utility_df <- data.frame(
+  # repeat zipcode in each row for each utility observation
+  "zipcode" = rep(zipcode, each=nrow(json$items$label)),
+  "label" = json$items$label,
+    "utility" = json$items$utility,
+    "rate_name" = json$items$name)
+
 }
-}
+
+test <- get_zip_utility("46033")
+utility_info <- pmap(list(zipcode=zipcodes), get_zip_utility)
+
+utilitu_info = as.data.frame(utility_info)
 ```

### 2023-04-26 solar_fam_shiny/app.R
update google sheets

@@ -13,12 +13,10 @@ library(sf) # map states and US
 library(shinydashboard)
 
 
-
 real_estate_metrics2021 <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=519968233")
-criteria_unweighted <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202", sheet = "FINAL_total_score") %>% 
+criteria_unweighted <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202", sheet = "unwt_criteria_scores") %>% 
   mutate(cbsa = as.character(cbsa)) %>% 
-  filter(city_msa != "Portland") %>% 
-  rename(student_tot_score = final_tot_students)
+  filter(city_msa != "Portland") 
 
 #census_api_key("ebab921b3002df9b71881ad6c426f34281ce0e11", overwrite = "TRUE")
 

### 2023-04-27 solar_fam_shiny/app.R
added population size range input

@@ -13,7 +13,8 @@ library(sf) # map states and US
 library(shinydashboard)
 
 
-real_estate_metrics2021 <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=519968233")
+real_estate_metrics2021 <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=519968233") %>% 
+  rename(pop_estimate = popestimate2021)
 criteria_unweighted <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202", sheet = "unwt_criteria_scores") %>% 
   mutate(cbsa = as.character(cbsa)) %>% 
   filter(city_msa != "Portland") 
@@ -51,9 +52,19 @@ ui <- fluidPage(
                       ) ### end mainPanel
              ), # ## END of tab 1
              ### tab2
-             tabPanel("Input Weights",
+             tabPanel("Test the Model",
                       tabsetPanel(
-                        tabPanel("Criteria Weights", 
+                        tabPanel("Step 1: Demographic Preferences",
+                                 sidebarLayout(
+                                   sidebarPanel(
+                                      sliderInput("slider2", label = h3("Population Size Range"), 
+                                                  min = 200000,  max = 5000000, step = 100000, sep = ",",
+                                                  value = c(300000, 600000))),
+                                         
+                                   mainPanel(dataTableOutput(outputId = "pop_test_table"))
+                                 )), # end step 1 tab panel
+                        tabPanel("Step 2: Landlord Inputs"), # end step 2 tab panel
+                        tabPanel("Step 3. Criteria Weights", 
                                  br(),
                                  sidebarLayout(
                                    sidebarPanel(
@@ -84,23 +95,14 @@ ui <- fluidPage(
                                      plotOutput(outputId = "wt_criteria_chart"))
                                    ) # end sidebar layout
                                  ), # end tab panel
-                        tabPanel("Data Table", dataTableOutput("wt_table_test"))
+                        tabPanel("Results:Data Table", 
+                                 mainPanel(dataTableOutput("wt_table_test"))),
+                        tabPanel("Results: Map")
                              
[…]

### 2023-04-27 src/MSAtotal_scores.Rmd
updated bar chart axis and criteria names

@@
              "Landlord Policy"="wt_landlord_policy",
-             "Electricity Production"="wt_electricity", 
-             "CO2 Emission Reduction"="wt_co2",
+             "Electricity Generation"="wt_electricity", 
+             "Avoided CO2 Emissions"="wt_co2",
              "Climate Risk Avoided"="wt_climate_risk",
-             "Health Impacts"="wt_health_impacts", 
-             "Solar IRR"="wt_solar_irr") %>%  
+             "Avoided Health Costs"="wt_health_impacts", 
+             "Solar Financials"="wt_solar_irr") %>%  
       unite("msa_state", city_msa:state, sep = ", ")
@@
 # client weights
-wt_criteria_client <- calculate_wt_criteria(criteria_all, client_weights)  %>%  
-  unite("msa_state", city_msa:state, sep = ", ")
+wt_criteria_client <- calculate_wt_criteria(criteria_all, client_weights)
 
@@
 # equity weights
-wt_criteria_equity <- calculate_wt_criteria(criteria_all, equity_weights) %>% 
-  unite("msa_state", city_msa:state, sep = ", ")
+wt_criteria_equity <- calculate_wt_criteria(criteria_all, equity_weights)
 
@@
 ggplot(data = client_score_tidy, aes(y = reorder(msa_state,  wt_criteria_score, FUN=sum), x = wt_criteria_score)) +
-    geom_col(aes(fill = factor(criteria, levels = c("Climate Risk Avoided", "CO2 Emission Reduction", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")))) +
+    geom_col(aes(fill = factor(criteria, levels = c("Climate Risk Avoided", "Avoided CO2 Emissions", "Electricity Generation", "Solar Financials", "Real Estate", "Landlord Policy", "Avoided Health Costs")))) +
   theme_bw() +
-  labs(x = "Total Score",
+  labs(x = "Investment Favorabliity Score",
        y = " ",
@@
    scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
-      breaks = c("Climate Risk Avoided", "CO2 Emission Reduction", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
+      breaks = c("Climate Risk Avoided", "Avoided CO2 Emissions", "Electricity Generation", "Solar Financials", "Real Estate", "Landlord Policy", "Avoided Health Costs")) +
         scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1)) 
@@
  ggplot(data = equity_tidy, aes(y = reorder(msa_state, wt_criteria_score, FUN=sum), x = wt_criteria_score)) + 
-    geom_col(aes(fill = criteria)) +
-  labs(x = "Total Score",
+  geom_col(aes(fill = factor(criteria, levels = c("Climate Risk Avoided", "Avoided CO2 Emissions", "Electricity Generation", "Solar Financials", "Real Estate", "Landlord Policy", "Avoided Health Costs")))) +
+  labs(x =  "Investment Favorabliity Score",
        y = " ",
@@
    scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
-      breaks = c("Climate Risk Avoidance", "CO2 Abatement", "Electricity Production", "Solar IRR", "Real Estate", "Landlord Policy", "Health Impacts")) +
-  scale_x_continuous(limits = c(0, 1.0), expand = c(0,0), breaks=seq(0, 1.0, 0.1))
+      breaks = c("Climate Risk Avoided", "Avoided CO2 Emissions", "Electricity Generation", "Solar Financials", "Real Estate", "Landlord Policy", "Avoided Health Costs")) +
+  scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1))
 
@@
 
-```{r}
+```{r COMBINED BAR CHARTS}
 library(patchwork)
@@
 ```
-
-Figure 1.
-Ranked metropolitan areas (n=29) based on investment favorability for rooftop solar PV on mutlifamily houses.
-Areas were scored using six criteria, each given a respective weight given stakeholder priorities.
-Each criteria is comprised of indicator metrics using data from the U.S.
-Census Bureau and NREL's reOpt tool.
-Graph A represents a real estate investor stakeholder who prioritizes climate risk due to long-term investments.
-B represents a equity-focused stakeholder who values impacts on the community, emissions, and

### 2023-04-27 src/demo_data_wrangling.Rmd
output demographic data to excel file

@@
 options(tigris_use_cache = TRUE)
+library(openxlsx)
 ```
@@
   clean_names() %>% 
-  rename(cbsa = x3) %>% 
-  rename(area = x4) %>% 
-  rename(year = x5) %>% 
-  rename(month = x6) %>% 
-  rename(civilian_labor_force = x7) %>% 
-  rename(employment = x8) %>% 
-  rename(unemployment = x9) %>% 
-  rename(unemployment_rate = x10) %>% 
+  rename(cbsa = x3,
+         area = x4,
+         year = x5,
+         month = x6,
+         civilian_labor_force = x7,
+         employment = x8,
+         unemployment = x9,
+         unemployment_rate = x10)%>% 
   mutate(employment_rate = (employment/civilian_labor_force)*100) # calculate
@@
 
-The following MSAs were not included in the housing dataset:
-
--   Augusta-Richmond County
--   Clarksville, TN-KY
--   Greeley, CO
--   Killeen, TX
--   Lubbock, TX
--   McAllen-Edinburg-Mission, TX
--   Salisbury, MD
-
 ### MSA data table
@@
 ```{r data table}
+# clean up names
+potential_MSAs <- potential_MSAs %>% 
+  rename(pop_growth = pop_percent_change20_21)
+
 # table seperated by MSAs
@@
 
+write.xlsx(potential_MSAs, here("data/intermediate/demographic.xlsx"), sheetName = "Sheet1")
+
 #gs4_create("real_estate_inputs")
 #output data into google sheet
-sheet_write(potential_MSAs, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "MSA_demographics")
+#sheet_write(potential_MSAs, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "MSA_demographics")
 ```
-
-### City Data Table
-
-MSAs were split up by cities to collect utility, rent, and landlord data.
-
-```{r MSAs seperate into cities}
-#  split up MSAs into cities to collect utility, rent, and landlord data.
-msa_seperated <- potential_MSAs %>% 
-  separate_rows(MSA, sep = "-") %>% 
-  rename(city = MSA) %>% # separate MSA cities into their own columns
-  filter(! city == "")
-
-#gs4_create("real_estate_inputs")
-sheet_write(msa_seperated, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "cities")
-
-```

### 2023-05-01 src/reopt_api.Rmd
Create code to get api data for each zip code

--- 2023-04-26 src/reopt_api.Rmd
@@
 
+### Obtain MSA zip code information
 ```{r}
-# Set your API key
-api_key <- "GZtZAUcX97VchjrLAeeN3amg2pslaO3hbtCQgXBl"
+token <- "eyJ0eXAiOiJKV1QiLCJhbGciOiJSUzI1NiIsImp0aSI6ImNmMmU0Yzg1ZWZlMDNiYTkxNTA4YzdlNWJiYjE1NjVkNzIyNjg4MzgwYTE4OWNjNmJkOWRjNjFiZWUyZjIwYzg0ZmJjNmZkOTc0MDg4YzRmIn0.eyJhdWQiOiI2IiwianRpIjoiY2YyZTRjODVlZmUwM2JhOTE1MDhjN2U1YmJiMTU2NWQ3MjI2ODgzODBhMTg5Y2M2YmQ5ZGM2MWJlZTJmMjBjODRmYmM2ZmQ5NzQwODhjNGYiLCJpYXQiOjE2ODI0ODQyOTYsIm5iZiI6MTY4MjQ4NDI5NiwiZXhwIjoxOTk4MTAzNDk2LCJzdWIiOiI0ODUyMyIsInNjb3BlcyI6W119.G7NL2Khpo602p_fzk6EiAZ0-KdXsVW-IO9mUqwHem0YFLaTQZezotmLqXmbcTi7BxRzc8UHNYBw415CCzKo81w"
 
-# Set the URL for the job submission endpoint
-url <- "https://developer.nrel.gov/api/reopt/stable/job"
+# cbsa-ziip data
+cbsa_url <- "https://www.huduser.gov/hudapi/public/usps?type=8&query=All"
 
-job_url <- paste0(url, "?api_key=", api_key)
+# zip-cbsa : type3
+zip_url <- "https://www.huduser.gov/hudapi/public/usps?type=3&query=All"
+
+# set the authorization header
+headers <- add_headers("Authorization" = paste("Bearer", token))
+
+# Send the GET request to retrieve the results
+cbsa_response <- GET(url = cbsa_url,add_headers("Authorization" = paste("Bearer", token)))
+
+# Extract the response content as JSON and convert it to a list
+response_content <- content(cbsa_response, as="text")
+response_cbsa <- fromJSON(response_content)
+
+# make data frame
+cbsa_city_data <- data.frame("geoid"= response_cbsa$data$results$geoid,
+                             "cbsa"= response_cbsa$data$results$cbsa,
+                            "city"= response_cbsa$data$results$city)
+
+zip_response <- GET(url = zip_url ,add_headers("Authorization" = paste("Bearer", token)))
+# Extract the response content as JSON and convert it to a list
+response_zip_content <- content(zip_response, as="text")
+response_list <- fromJSON(response_zip_content)
+
+zip_data <- data.frame("geoid"= response_list$data$results$geoid,
+                            "zipcode"= response_list$data$results$zip,
+                             "state"= response_list$data$results$state)
+# combine data sets
+zip_cbsa_data <- cbsa_city_data %>% 
+  right_join(zip_data)
+
 ```
 
+```{r polulation estimates}
+api_key <- "ebab921b3002df9b71881ad6c426f34281ce0e11"
+# Add key to .Renviron
+Sys.setenv(CENSUS_KEY="ebab921b3002df9b71881ad6c426f34281ce0e11")
+# Reload .Renviron
+readRenviron("~/.Renviron")
+# Check to see that the expected key is output in your R console
+Sys.getenv("CENSUS_KEY")
+
+host <- "https://api.census.gov/data"
+year <- "2021"
+pop_est_stem <- "pep/population"
+
+job_end <- "acs/acs1?get=NAME,B01001_001E&for=metropolitan%20statistical%20area/micropolitan%20statistical%20area:*&key="
+
+version_url <- file.path(host, year, job_end)
+pop_list <- fromJSON(version_url) %>% # retrieved cbsa numbers, msa names, and some value??
+  row_to_names(1) %>% clean_names() %>% 
+  as.data.frame() %>% 
+  rename("msa"= name,
+                             "cbsa" = metropolitan_statistical_area_micropolitan_statistical_area,
+                             "pop_esta"= b01001_001e)
+
+msa_zip_pop_est_data <- pop_list %>% 
+  left_join(cbsa_city_data) %>% 
+  inner_join(zip_data)
+
+```
+
+# reopt inputs
 ```{r}
@@
 
+# Set the payload for the job submission request for reopt
 ```{r}
-
 # Set the payload for the job submission request
@@
       "roof_squarefeet" = 5000.0,
-      "land_acres" = 1.0,
       "PV" = list(
         "installed_cost_us_dollars_per_kw" = 2000.0,
-        "tilt" = 34.579,
-        "degradation_pct" = 0.005,
-        "macrs_option_years" = 5,
-        "federal_itc_pct" = 0.3,
-        "module_type" = 0,
-        "array_type" = 1,
-        "om_cost_us_dollars_per_kw" = 16.0,
-        "macrs_itc_reduction" = 0.5,
-        "azimuth" = 180.0,
-        "federal_rebate_us_dollars_per_kw" = 350.0
+        "federal_rebate_us_dollars_per_kw" = 350.0,
+        # changes from default
+        "module_type" = 1,
+        "dc_ac_ratio" = 1.35, # changed default
+        "losses" = 0.05,   # changed default
+        "federal_itc_pct" = 0.3,  # changed default
+        "location" = "roof", # restrict projects to roof
+        "can_net_meter" = TRUE
       ),
@@
 
+```
+
+## OpenEI 
+```{r}
+## open EI api
+openEI_api_key <- "rU6MEqhK9OJ7obAnmcbzYhpFOenfBM5Q8GF0yJFA"
+
+openEI_start <- "https://api.openei.org/utility_rates?version=latest&format=json&sector=Residential&detail=full&approved=true&sort_select=effectiveDate&direction=desc&"
+# Set the URL for the job submission endpoint
+
+openEI_url <- paste0(openEI_start, "address=", "46033", "&api_key=", openEI_api_key)
+EIresponse <- GET(openEI_url)
+# extract contents
+content <- content(EIresponse, "text")
+utility_data <- fromJSON(content)
+
+ # extract utility rate information
+EIutility_df <- list(
+  start_date = utility_data$items$startdate,
+  #revisions = utility_data$items$revisions,
+  label = utility_data$items$label,
+  utility = utility_data$items$utility,
+  rate_name = utility_data$items$name)
+
+startdates <- EIutility_df$start_date
+start_dates_rev <- lapply(startdates, function(x) as.Date(x/86400, origin="1970-01-01"))
+
+```
+
+```{r}
+## function to select the most recent utiilty rate information for a given area code
+get_zip_utility <- function(zipcode){
+  openEI_start_url <- "https://api.openei.org/utility_rates?version=latest&format=json&sector=Residential&detail=full&approved=true&address="
+openEI_api_end <- "&api_key=rU6MEqhK9OJ7obAnmcbzYhpFOenfBM5Q8GF0yJFA"
+zipcode=zipcode
+
+  openEI_url <- paste0(openEI_start_url, zipcode, openEI_api_end)
+  response <- GET(openEI_url)
+  # extract content and parse as JSON
+  content <- content(EIresponse, "text")
+  json <- fromJSON(content)
+
+  utility_df <- data.frame(
+  # repeat zipcode in each row for each utility observation
+  "zipcode" = rep(zipcode, each=nrow(json$items$label)),
+  "start_date" = utility_data$items$startdate,
+  "label" = json$items$label,
+    "utility" = json$items$utility,
+    "rate_name" = json$items$name)
+
+  # Find index of row with most recent start date value
+  max_index <- which.max(utility_df$start_date)
+
+# Extract row with greatest seconds value
+  max_row <- utility_df[max_index, ]
+
+}
+
+test <- get_zip_utility(46033) %>% 
+     mutate(start_date = as.Date(start_date/86400, origin="1970-01-01"))
+
+```
+
+```{r}
+# Define a vector of zip codes
+zipcodes <- c("90210", "10001", "60601", "90245", "90001", "90250", "90260", "90278", "90501", "90502")
+
+# Initialize an empty list to store the results
+utility_list <- list()
+
+# Loop through each zip code and call the get_zip_utility function
+for (zipcode in zipcodes) {
+  utility_data <- get_zip_utility(zipcode)
+  utility_list[[zipcode]] <- utility_data
+}
+
+# Combine the results into a single data frame
+utility_df <- do.call(rbind, utility_list)
+
+```
+
+```{r STUFF TO WORK ON}
+
+net_metering_limit_kw # by state
+```
+
+### Reopt API
+
+```{r}
+# Set your API key
+api_key <- "GZtZAUcX97VchjrLAeeN3amg2pslaO3hbtCQgXBl"
+
+# Set the URL for the job submission endpoint
+url <- "https://developer.nrel.gov/api/reopt/stable/job"
+
+job_url <- paste0(url, "?api_key=", api_key)
 ```
@@
 ```{r}
-
-## energy goals
-#cost_savings and clean_energy
-## inputs not default 
-host_eff <- 0
-scc = 185
-
 #PV & wind space available:  ❌ Land only ✅ Roofspace only ❌ Land & roofspace

### 2023-05-01 R/openEI_call.R
create folder for functions

@@ -1,0 +1,38 @@
+# @param zipcode 
+
+
+## function to select the most recent utiilty rate information for a given area code
+get_zip_utility <- function(zipcode){
+  # query for residential rates that are approved by experts and sorted by most recent effective date
+  openEI_start_url <- "https://api.openei.org/utility_rates?version=latest&format=json&sector=Residential&detail=full&approved=true&sort_select=effectiveDate&direction=desc&address="
+  openEI_api_end <- "&api_key=rU6MEqhK9OJ7obAnmcbzYhpFOenfBM5Q8GF0yJFA"
+  
+  # put url together
+  openEI_url <- paste0(openEI_start_url, zipcode, openEI_api_end)
+  # get the results for the query
+  response <- GET(openEI_url)
+  # extract content and parse as JSON
+  content <- content(response, "text")
+  json <- fromJSON(content)
+  
+  # error handling
+  if(length(json$items) == 0) {
+    message(paste("No utility data available for zipcode", zipcode))
+    return(NA)
+  }
+  
+  utility_df <- data.frame(
+    # repeat zipcode in each row for each utility observation
+    "zipcode" = rep(zipcode, each=nrow(json$items$label)),
+    "start_date" = json$items$startdate,
+    "label" = json$items$label,
+    "utility" = json$items$utility,
+    "rate_name" = json$items$name) 
+  
+  # Find index of row with most recent start date value
+  max_index <- which.max(utility_df$start_date) 
+  
+  # Extract row with greatest seconds value
+  max_row <- utility_df[max_index, ]
+  return(max_row)
+}

### 2023-05-04 src/landlord.Rmd
create excel with landlord policies for each state

@@
 library(ggplot2)
+library(readxl)
+library(here)
 
@@
 
-# select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")
+# for shiny app
+state_deposit_data<- read_xlsx(here("data/raw/landlord_state_data.xlsx")) %>% clean_names()
+
+eviction_notice <- read_xlsx(here("data/raw/landlord_state_data.xlsx"), sheet= "Sheet2") %>% 
+  clean_names() %>% 
+  rename(eviction_days = sel_min_nonpayment)
+
 ```
 
-```{r}
-#filter by new MSA list
-# new_MSA_list <- unique(select_MSAs$cbsa)
-# 
-# MSA_landlord_data <- MSA_landlord_data %>% 
-#   filter(cbsa %in% c(new_MSA_list))
+```{r wrangling state deposit limits}
+# Define a regular expression pattern to match target phrases
+phrase_pattern <- "(one|two)\\smonth|no\\sstatutory\\slimit"
+
+# Define a function to extract matched words from each sentence
+extract_words <- function(sentence, pattern) {
+  words <- regmatches(sentence, gregexpr(pattern, sentence, ignore.case = TRUE))[[1]]
+  if (length(words) > 0) {
+    return(paste(words, collapse = " "))
+  } else {
+    return(NA)
+  }
+}
+
+state_deposit_data$deposit_limit <- NA
+
+# Create a new column with matched words
+state_deposit_data <- state_deposit_data %>%
+  mutate(matched_words = sapply(limit, extract_words, pattern = phrase_pattern))
+
+# put matched words in new column
+state_deposit_data$deposit_limit <- sapply(state_deposit_data$matched_words, extract_words, pattern = phrase_pattern)
+
+# rename deposit data to apply weights
+landlord_state_data <- state_deposit_data %>%
+  mutate(deposit_limit = case_when(
+    matched_words == "One month"   ~ "1_month_rent",
+    matched_words == "Two month" ~ "2_months_rent",
+    matched_words == "No statutory limit" ~ "N",
+    TRUE ~ NA_character_
+  )) %>% 
+  select(state, limit, deposit_limit) %>% 
+  inner_join(eviction_notice, by = join_by(state == jurisdictions))
+
+#write result to google sheet
+sheet_write(landlord_state_data, ss = "https://docs.google.com/spreadsheets/d/1A6o934oBp9Sdbxrx30GkVrGaX9PyzpS1pZc8Pz6Aoro/edit#gid=0", sheet = "MSA_landlord_data") # manually edited eviction data
+
+# write output 
+write.xlsx(landlord_state_data, here("data/intermediate/demographic.xlsx"), sheetName = "Sheet1")
 ```
@@
 
-# #notice_of_entry_days
-# MSA_landlord_data <- MSA_landlord_data %>% 
-#   mutate(
-#     notice_for_entry_days_num = case_when(
-#       notice_for_entry_days == "N" ~ 1.0, 
-#       notice_for_entry_days == "0" ~ 1.0,
-#       notice_for_entry_days == "0.5" ~ 0.75,
-#       notice_for_entry_days == "1" ~ 0.5,
-#       notice_for_entry_days == "2" ~ 0.0,))
 ```
@@
 
-datatable(MSA_landlord_data, extensions = 'Buttons', options = list(
-    dom = 'Bfrtip',
-    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
-  )
-)
 ```

### 2023-05-09 src/real_estate.Rmd
outputting metric data to excel files

@@
 library(tidyverse)
-library(janitor)
+library(janitor) # clean_names()
 library(googlesheets4)
 library(DT)
+library(openxlsx)
+library(here)
+source(here("R/normalize.R"))
 
-city_realEstate_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990")
+city_occupancy_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990") %>% 
+  clean_names() %>% 
+  rename(occ_rate2021 = percent_occupancy_2021)
 
-select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")
+MSA_dempgraphic_data <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing", sheet = "MSA_demographics") %>% 
+  clean_names() %>%
+
+  inner_join(city_occupancy_data, by = join_by(cbsa== cbsa, anchor_city==city))
 
@@
 ```{r}
-# aggregate yardi data into MSAs by cbsa
-MSA_realEstate_data <- city_realEstate_data %>% 
-  clean_names() %>% 
-  rename(avg_rent2022 = x2022_rent) %>% 
-   # calculate change in rent
-  mutate(rent_change = (avg_rent2022 - average_rent_2021)/average_rent_2021 * 100) %>% 
-  group_by(cbsa) %>% 
-  # calculate average values for MSAs with more than one city
-  summarize(avg_rent_2021 = mean(average_rent_2021, na.rm = TRUE),
-            avg_rent_2022 = mean(avg_rent2022, na.rm = TRUE),
-            avg_occupancy_21 = mean(percent_occupancy_2021, na.rm = TRUE),
-            avg_rent_change = mean(rent_change, na.rm = TRUE))
+# calculate metrics
+real_estate_metrics<- MSA_dempgraphic_data %>% 
+    # calculate rent change
+  mutate(med_rent_change = (med_rent2022 - med_rent2021)/med_rent2021 * 100,
+         house_price_2021 = as.numeric(house_price_2021),
+         income_to_homeprice = med_income/house_price_2021,
+         med_rent_to_income = (med_rent2021*12)/med_income) %>%  # multiply monthly rent x 12 for annual rent!
+  select(!c(state.y, occ_rate2021, popestimate2021, med_rent2021, med_rent2022, house_price_2021, med_income)) %>% 
+  rename(pop_change = pop_percent_change20_21)
+
+write.xlsx(real_estate_metrics, here("data/intermediate/metric_data.xlsx"), sheetName = "real_estate_metrics")
+
+#sheet_write(compliedMSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=426650684", sheet = "real_estate_metrics")
 
@@
 
+# Normalize Metrics
+
 ```{r}
-# add MSA_median rent data
-compliedMSA_realEstate<- select_MSAs %>% 
-  left_join(MSA_realEstate_data) %>% 
-    # calculate rent change
-  mutate(med_rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100) %>% 
-  # calculate income to homeprice
-  mutate(house_price_2021 = as.numeric(house_price_2021)) %>% 
-  mutate(income_to_homeprice = median_income/house_price_2021) %>% 
-  # calculate rent to income ratio
-  # multiply monthly rent x 12 for annual rent!
-  mutate(med_rent_to_income = (median_rent2021*12)/median_income)
+n_real_estate_metrics <- real_estate_metrics %>%
+  mutate(n_pop_change = normalize(pop_change),
+         n_employ_change = normalize(employ_change),
+         n_income_homprice = 1- normalize(income_to_homeprice), # subtracted from 1 so higher values are more favorable
+         n_rent_income = 1 - normalize(med_rent_to_income),
+         n_rent_change = normalize(med_rent_change)
+
+        )
 
@@
   mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
-  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
-  # normalize rent change
-  mutate(n_avg_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))
+  mutate(n_avg_occupancy = (occ_rate2021 - min(occ_rate2021)) / (max(occ_rate2021) - min(occ_rate2021)), na.rm = FALSE)
 
@@
                   )
-  n_MSA_realEstate$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
+  n_MSA_realEstate$real_estate_score[i] = sum(MSA_x_data * real_estate_weights)
 }
@@
 ```{r}
-# create downloadable data table
-datatable(n_MSA_realEstate, extensions = 'Buttons', options = list(
-    dom = 'Bfrtip',
-    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
-  )
-)
+wt_real_estate_metrics <-n_MSA_realEstate %>%
+
+   mutate(_score = apply(.[, 4:10], 1, sum))
+  mutate(weighted_pop_change = 0.35*n_pop_change,
+                  weighted_emp_change = 0.20*n_employ_change,
+                  weighted_occ = 0.15 * n_avg_occupancy,
+                  weighted_rent_change = 0.10*n_med_rent_change,
+                  weighted_rent_income = 0.10 * n_med_rent_to_income_inverted, 
+                  weighted_income_homeprice = 0.10 * n_income_to_homeprice_inverted) %>% 
+  mutate(real_estate_score = weighted_pop_change + weighted_emp_change + weighted_occ + weighted_rent_change + weighted_rent_income + weighted_income_homeprice)
+
 ```
@@
 #write result to google sheet
-# gs4_create("MSA_real_estate_score")
-sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "n_MSA_realEstate")
+#gs4_create("anchor_real_estate_score")
+sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "anchor_realEstate")
 ```

### 2023-05-09 R/normalize.R

@@ -1,0 +1,12 @@
+#' Normalization Equation
+#'
+#' THis function normalizes values using the min and max of a given vector 
+#' @param       col a column of data to normalize
+#'
+#' @return  Normalized column
+
+normalize = function(col) {
+  normalized_col = (col-min(col))/(max(col)-min(col))
+  return(normalized_col)
+}
+                                    

### 2023-05-09 src/real_estate.Rmd
output weighted real estate metrics into excel with real estate score

@@
     # calculate rent change
-  mutate(med_rent_change = (med_rent2022 - med_rent2021)/med_rent2021 * 100,
+  mutate(cbsa = factor(cbsa),
+         rent_change = (med_rent2022 - med_rent2021)/med_rent2021 * 100,
          house_price_2021 = as.numeric(house_price_2021),
          income_to_homeprice = med_income/house_price_2021,
-         med_rent_to_income = (med_rent2021*12)/med_income) %>%  # multiply monthly rent x 12 for annual rent!
-  select(!c(state.y, occ_rate2021, popestimate2021, med_rent2021, med_rent2022, house_price_2021, med_income)) %>% 
+         rent_to_income = (med_rent2021*12)/med_income) %>%  # multiply monthly rent x 12 for annual rent!
+  select(!c(state.y, popestimate2021, med_rent2021, med_rent2022, house_price_2021, med_income)) %>% 
   rename(pop_change = pop_percent_change20_21)
@@
 n_real_estate_metrics <- real_estate_metrics %>%
-  mutate(n_pop_change = normalize(pop_change),
-         n_employ_change = normalize(employ_change),
-         n_income_homprice = 1- normalize(income_to_homeprice), # subtracted from 1 so higher values are more favorable
-         n_rent_income = 1 - normalize(med_rent_to_income),
-         n_rent_change = normalize(med_rent_change)
+  # normalize cols and rename columns to start with n
+  mutate(across(5:10, normalize, .names = "n_{.col}")) %>% 
+  mutate(n_income_to_homeprice = 1- n_income_to_homeprice, # subtracted from 1 so higher values are more favorable
+         n_rent_to_income = 1 - n_rent_to_income) %>% 
+  select(cbsa, msa, anchor_city, state.x, starts_with("n_"))
 
-        )
-
+write.xlsx(n_real_estate_metrics, here("data/intermediate/normalized_data.xlsx"), sheetName = "real_estate_metrics")
 ```
@@
 ```{r}
-# normalize columns into new data table
-n_MSA_realEstate <- compliedMSA_realEstate %>% 
-  # normalize population data
-  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>%
-  # normalize employment growth
-  mutate(n_employ_change = (avg_emp_YoY - min(avg_emp_YoY)) / (max(avg_emp_YoY) - min(avg_emp_YoY))) %>% 
-  # normalize rent to income
-  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
-  # normalize income to home price
-  mutate(n_med_rent_change = (med_rent_change - min(med_rent_change)) / (max(med_rent_change) - min(med_rent_change))) %>% 
-  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
-  mutate(n_avg_occupancy = (occ_rate2021 - min(occ_rate2021)) / (max(occ_rate2021) - min(occ_rate2021)), na.rm = FALSE)
+real_estate_weights <- c(0.35, # pop change weight
+                         0.20, # employment change weight
+                         0.15, # occupancy weight
+                         0.10, # rent change weight
+                         0.10, # rent to income weight
+                         0.10) # income to homeprice weight
 
-```
+wt_real_estate_metrics <- n_real_estate_metrics %>% 
+  mutate(wt_pop_change = n_pop_change* real_estate_weights[1],
+         wt_employ_change = n_employ_change* real_estate_weights[2],
+         wt_occ = n_occ_rate2021* real_estate_weights[3],
+         wt_rent_change = n_rent_change* real_estate_weights[4],
+         wt_rent_income = n_rent_to_income* real_estate_weights[5],
+         wt_income_homeprice = n_income_to_homeprice* real_estate_weights[6]) %>% 
+  select(!starts_with("n_")) %>% 
+  mutate(real_estate_score = apply(.[, 5:10], 1, sum)) %>% 
+  mutate_if(is.numeric,round, digits = 4)
 
-```{r}
-# invert normalized rent to income ratio so lower values (more favorable) have higher scores
-n_MSA_realEstate <- n_MSA_realEstate %>% 
-  mutate(n_med_rent_to_income_inverted = 1 - n_med_rent_to_income) %>% 
-  mutate(n_income_to_homeprice_inverted = 1 - n_income_to_homeprice)
-
-```
-
-```{r}
-# weights from owen
-real_estate_weights <- c(0.35, 0.20, 0.15, 0.10, 0.10, 0.10)
-
-#make real_estate_score column
-n_MSA_realEstate <- n_MSA_realEstate %>% 
-  select(cbsa, MSA, starts_with("n_"), ) %>% 
-  mutate(real_estate_score = MSA) 
-```
-
-```{r}
-size <- dim(n_MSA_realEstate)
-
-for (i in 1:size[1]) {
-  MSA_x_data <- c(n_MSA_realEstate$n_pop_change[i],
-                  n_MSA_realEstate$n_employ_change[i],
-                  n_MSA_realEstate$n_avg_occupancy[i],
-                  n_MSA_realEstate$n_med_rent_change[i],
-                  n_MSA_realEstate$n_med_rent_to_income_inverted[i], 
-                  n_MSA_realEstate$n_income_to_homeprice_inverted[i]
-                  )
-  n_MSA_realEstate$real_estate_score[i] = sum(MSA_x_data * real_estate_weights)
-}
-
-```
-
-```{r}
-wt_real_estate_metrics <-n_MSA_realEstate %>%
-
-   mutate(_score = apply(.[, 4:10], 1, sum))
-  mutate(weighted_pop_change = 0.35*n_pop_change,
-                  weighted_emp_change = 0.20*n_employ_change,
-                  weighted_occ = 0.15 * n_avg_occupancy,
-                  weighted_rent_change = 0.10*n_med_rent_change,
-                  weighted_rent_income = 0.10 * n_med_rent_to_income_inverted, 
-                  weighted_income_homeprice = 0.10 * n_income_to_homeprice_inverted) %>% 
-  mutate(real_estate_score = weighted_pop_change + weighted_emp_change + weighted_occ + weighted_rent_change + weighted_rent_income + weighted_income_homeprice)
-
+write.xlsx(wt_real_estate_metrics, here("data/intermediate/weighted_criteria_data.xlsx"), sheetName = "wt_real_estate_metrics")
 ```
@@
 #gs4_create("anchor_real_estate_score")
-sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "anchor_realEstate")
+#sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "anchor_realEstate")
 ```

### 2023-05-10 src/landlord.Rmd
cleaning up landlord code and output results into excel

--- 2022-12-06 src/landlord.Rmd
@@
 library(DT)
+library(ggplot2)
+library(readxl)
+library(here)
+source(here("R/normalize.R"))
 
@@
 
-# select_MSAs <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit?usp=sharing")
-```
-
-```{r}
-#filter by new MSA list
-# new_MSA_list <- unique(select_MSAs$cbsa)
-# 
-# MSA_landlord_data <- MSA_landlord_data %>% 
-#   filter(cbsa %in% c(new_MSA_list))
 ```
@@
   mutate(
-    security_depopsit_limit_num = case_when(
+    security_deposit_limit_num = case_when(
       security_deposit_limit == "N" ~ 1.0, 
       security_deposit_limit == "1_month_rent" ~ 0.0,
-      security_deposit_limit == "2_months_rent" ~ 0.5))
+      security_deposit_limit == "2_months_rent" ~ 0.5)) %>% 
+    #"invert" normalized score for eviction notice days
+#1 - normalized score (since more eviction notice is worse for landlords)
+  mutate(n_eviction_notice = 1- normalize(eviction_notice_days))
 
-# #notice_of_entry_days
-# MSA_landlord_data <- MSA_landlord_data %>% 
-#   mutate(
-#     notice_for_entry_days_num = case_when(
-#       notice_for_entry_days == "N" ~ 1.0, 
-#       notice_for_entry_days == "0" ~ 1.0,
-#       notice_for_entry_days == "0.5" ~ 0.75,
-#       notice_for_entry_days == "1" ~ 0.5,
-#       notice_for_entry_days == "2" ~ 0.0,))
+# assign numeric values to security deposit limit for all states
+state_landlord_policy <- state_landlord_policy %>% 
+  mutate(security_deposit_limit = case_when(
+      deposit_limit == "N" ~ 1.0, 
+      deposit_limit == "1_months_rent" ~ 0.0,
+      deposit_limit == "1_month_rent" ~ 0.0,
+      deposit_limit == "1.5_months_rent" ~ 0.25,
+      deposit_limit == "2_months_rent" ~ 0.5,
+      deposit_limit == "2.5_months_rent" ~ 0.75,
+      deposit_limit == "3_months_rent" ~ 0.75)) %>% 
+   #"invert" normalized score for eviction notice days
+#1 - normalized score (since more eviction notice is worse for landlords)
+  mutate(n_eviction_notice = 1- normalize(eviction_days))
+
 ```
 
-```{r}
-#normalize eviction notice (days)
-MSA_landlord_data <- MSA_landlord_data %>% 
-  mutate(
-    eviction_notice_days_normalized = (eviction_notice_days-min(eviction_notice_days))/(max(eviction_notice_days)-min(eviction_notice_days)) 
-  )
-
-#"invert" normalized score for eviction notice days
-#1 - normalized score (since more eviction notice is worse for landlords)
-MSA_landlord_data <- MSA_landlord_data %>% 
-  mutate(
-    eviction_notice_days_normalized_inverted = 1 - eviction_notice_days_normalized
-  )
-```
+#calculate landlord category score
 
 ```{r}
-#calculate landlord category score
 

### 2023-05-10 src/landlord.Rmd
added landlord policy data for all states

@@
 library(tidyverse)
-library(kableExtra)
 library(DT)
-library(ggplot2)
 library(readxl)
+library(xlsx)
 library(here)
@@
 
-MSA_landlord_data <- read_sheet("https://docs.google.com/spreadsheets/d/1zCePYRyJ5lb1Z5jvCujg17EtwqpxJN1E1_HwJL66uLM/edit#gid=122953884")
+MSA_landlord_data <- read_sheet("https://docs.google.com/spreadsheets/d/1zCePYRyJ5lb1Z5jvCujg17EtwqpxJN1E1_HwJL66uLM/edit#gid=122953884") %>% 
+  clean_names()
 
@@
 
+# Normalizing Landlord Metrics
+
 ```{r}
 #security_deposit_limit
-MSA_landlord_data <- MSA_landlord_data %>% 
+landlord_metrics <- MSA_landlord_data %>% 
   mutate(
@@
 #1 - normalized score (since more eviction notice is worse for landlords)
-  mutate(n_eviction_notice = 1- normalize(eviction_notice_days))
+  mutate(n_eviction_notice = 1- normalize(eviction_notice_days),
+         n_security_deposit = normalize(security_deposit_limit_num))
+
+# output into excel
+xlsx::write.xlsx2(landlord_metrics, here("data/intermediate/metric_data.xlsx"), sheetName = "landlord_metrics", append=TRUE)
 
@@
 #1 - normalized score (since more eviction notice is worse for landlords)
-  mutate(n_eviction_notice = 1- normalize(eviction_days))
+  mutate(n_eviction_notice = 1- normalize(eviction_days),
+         n_security_deposit = normalize(security_deposit_limit))
 
@@
 
-#calculate landlord category score
+# Calculate landlord category score
 
 ```{r}
-
 #input landlord weights for each indicator in the following order: eviction notice days, security deposit limit, notice for entry
@@
 #make landlord_score column
-MSA_landlord_data <- MSA_landlord_data %>% 
+wt_landlord_data <- landlord_metrics %>% 
+  select(msa, cbsa, state, c(starts_with("n_"))) %>% 
   mutate(
-    landlord_score = MSA #this is just a place holder
-  )
-size <- dim(MSA_landlord_data)
-for (i in 1:size[1]) {
-  MSA_x_data <- c(MSA_landlord_data$eviction_notice_days_normalized_inverted[i],
-                  MSA_landlord_data$security_depopsit_limit_num[i]
-                  )
-  MSA_landlord_data$landlord_score[i] = weighted.mean(MSA_x_data, landlord_weights)
-}
-```
+    wt_evict_notice = n_eviction_notice*landlord_weights[1],
+    wt_secur_deposit = n_security_deposit*landlord_weights[2]) %>% 
+  mutate(landlord_score = apply(.[, 6:7], 1, sum)) %>% 
+  select(msa, state, cbsa, 6:8) %>% 
+  mutate_if(is.numeric,round, digits = 5)
 
-```{r}
 #If there is rent control, immediately disqualify location (set score to -999)
-for (j in 1:size[1]) {
-  if(MSA_landlord_data$rent_control[j] == "Y"){
-    MSA_landlord_data$landlord_score[j] = -999
-  }
-}
+#for (j in 1:nrow(MSA_landlord_data)) {
+ # if(MSA_landlord_data$rent_control[j] == "Y"){
+  #  MSA_landlord_data$landlord_score[j] = -999
+  #}
+
+xlsx::write.xlsx2(wt_landlord_data, here("data/intermediate/weighted_data.xlsx"), sheetName = "wt_landlord_metrics", append=TRUE)
 
@@
 
+# reOpt Metric data wrangling and normalization
+
 ```{r}
-#write result to google sheet
-#gs4_create("MSA_landlord_score")
-sheet_write(MSA_landlord_data, ss = "https://docs.google.com/spreadsheets/d/1nLpDhM88a3rEBDVK5RZlk2B6QuwVIwvsfE5HeBz0mLo/edit?usp=sharing", sheet = "MSA_landlord_data")
+# read in reopt data (for STCOFIPS codes)
+reopt_metrics <- read_xlsx(here("data/intermediate/metric_data.xlsx"), sheet="reopt_metrics") %>% 
+  clean_names() %>% 
+  drop_na() %>% 
+  rename(
+  electricity_generation = annual_renewable_electricity_percent_of_total,
+  total_co2 = lifecycle_total_t_co2,
+  health_emissions_cost = lifecycle_costs_of_health_emissions_difference_from_bau) %>% 
+  select(!system_capacity_100_for_100_apartments) %>% 
+   mutate(stcofips = as.character(stcofips_5_digit))
+
+# merge reopt data with climate risk data
+n_reopt_metrics <- reopt_metrics %>% 
+mutate(across(electricity_generation:total_co2, normalize, .names = "n_{.col}")) %>% 
+  select(cbsa, city_msa, state, starts_with("n_"))
+
+xlsx::write.xlsx2(n_reopt_metrics, here("data/intermediate/metric_data.xlsx"), sheetName = "n_reopt_metrics", append=TRUE)
+
 ```
+
+# reOpt Metric data wrangling and normalization
+
+```{r}
+#read in National Risk Index (NRI) county data
+climate_risk_metrics <- read_csv(here("data/raw/NRI_Table_Counties.csv")) %>% 
+  clean_names() %>% 
+  select(state, stateabbrv, county, stcofips, risk_score) %>% 
+  right_join(reopt_metrics, by = join_by(stcofips==stcofips)) %>% 
+  select(1:"city_msa")
+
+# normalize climate risk score
+n_climate_risk <- climate_risk_metrics %>% 
+  mutate(n_climate_risk = 1- normalize(risk_score)) #invert normalized scores so that places with less climate risk have a better score
+
+xlsx::write.xlsx2(n_climate_risk, here("data/intermediate/metric_data.xlsx"), sheetName = "n_climate_risk_metrics", append=TRUE)
+```

### 2023-05-10 R/normalize.R
created normalization function for metric normalization

@@ -6,7 +6,7 @@
 #' @return  Normalized column
 
 normalize = function(col) {
-  normalized_col = (col-min(col))/(max(col)-min(col))
+  normalized_col = (col-min(col))/ (max(col)-min(col))
   return(normalized_col)
 }
                                     

### 2023-05-11 src/MSAtotal_scores.Rmd
calculating scores & cleaning up code

@@
 library(dplyr)
+library(here)
 ```
 
-```{r}
+```{r, eval=FALSE}
 # read in the normalized data
@@
 
+```
+
+```{r}
+real_estate_metrics <- read_xlsx(here("data/intermediate/metric_data.xlsx"), sheet="real_estate_metrics") %>% 
+  select(cbsa, anchor_city, msa, starts_with("n_")) %>% 
+  mutate(cbsa = factor(cbsa))
+
+landlord_metrics <- read_xlsx(here("data/intermediate/metric_data.xlsx"), sheet="landlord_metrics") %>% 
+  select(cbsa, starts_with("n_"))%>% 
+  mutate(cbsa = factor(cbsa))
+
+climate_risk_metrics <- read_xlsx(here("data/intermediate/metric_data.xlsx"), sheet="climate_risk_metrics") %>% 
+  mutate(cbsa = factor(cbsa)) %>% 
+  select(!risk_score)
+
+reopt_metrics <- read_xlsx(here("data/intermediate/metric_data.xlsx"), sheet="reopt_metrics") %>% 
+  select(cbsa, starts_with("n_")) %>% 
+  mutate(cbsa = factor(cbsa))
+
+```
+
+## Input Weights
+
+### Metric Inputs for Real Estate and Landlord Criteria
+
+```{r}
+
+real_estate_weights <- c(0.35, # pop change weight
+                         0.20, # employment change weight
+                         0.15, # occupancy weight
+                         0.10, # rent change weight
+                         0.10, # rent to income weight
+                         0.10) # income to homeprice weight
+
+# Weights must sum to 1
+landlord_weights <- c(0.8, # eviction notice weight
+                      0.2) # security deposit limit weight
+
+```
+
+# Weight Metrics
+
+```{r}
+# weight real estate metrics
+wt_real_estate_metrics <- real_estate_metrics %>% 
+  mutate(wt_pop_change = n_pop_growth* real_estate_weights[1],
+         wt_employ_change = n_employ_change* real_estate_weights[2],
+         wt_occ = n_occ_rate2021* real_estate_weights[3],
+         wt_rent_change = n_rent_change* real_estate_weights[4],
+         wt_rent_income = n_rent_to_income* real_estate_weights[5],
+         wt_income_homeprice = n_income_to_homeprice* real_estate_weights[6]) %>% 
+  select(!starts_with("n_")) %>% 
+  mutate(real_estate_score = apply(.[,4:9], 1, sum)) %>% 
+  mutate_if(is.numeric,round, digits = 4)
+
+# wt_landlord metrics
+wt_landlord_metrics <- landlord_metrics %>%
+      # multiply each criteria by respective weight
+      mutate(wt_eviction = n_eviction_notice*landlord_weights[1],
+             wt_deposit = n_security_deposit*landlord_weights[2]) %>% 
+      # for loop to sum weighted columns for each row
+      mutate(landlord_score = apply(.[, 4:5], 1, sum)) %>% 
+   mutate_if(is.numeric,round, digits = 4)  %>% 
+  select(!starts_with("n_")) 
+```
+
+# Weight Criteria
+
+```{r}
+  calculate_wt_criteria <- function(data, weights){
+    wt_criteria <- data %>%
+      # multiply each criteria by respective weight
+      mutate(wt_real_estate = real_estate_score * weights[1],
+             wt_landlord_policy = landlord_score * weights[2], 
+             wt_solar_irr = financial_score * weights[3],
+            wt_co2 = co2_score * weights[4], 
+             wt_health_impacts = health_impact_score * weights[5],
+             wt_electricity = electricity_score * weights[6],
+             wt_climate_risk = climate_risk_score * weights[7]) %>% 
+      # select weighted columns and rename them
+      select(!c(ends_with("_score"))) %>%
+      # create column for investment score
+      mutate(investment_score = 0) %>%
+      # for loop to sum weighted columns for each row
+      mutate(investment_score = apply(.[, 6:12], 1, sum)) %>% 
+      rename("Real Estate"="wt_real_estate",
+             "Landlord Policy"="wt_landlord_policy",
+             "Electricity Generation"="wt_electricity", 
+             "Avoided CO2 Emissions"="wt_co2",
+             "Climate Risk Avoided"="wt_climate_risk",
+             "Avoided Health Costs"="wt_health_impacts", 
+             "Solar Financials"="wt_solar_irr")
+  }
+
+```
+
+```{r}
+
+unwt_criteria <- wt_real_estate_metrics %>% 
+  inner_join(wt_landlord_metrics) %>% 
+  select(!starts_with("wt_")) %>% 
+  inner_join(reopt_metrics) %>% 
+  inner_join(climate_risk_metrics, by = join_by(cbsa==cbsa)) %>% 
+  select(!city_msa) %>% 
+    rename(financial_score = n_irr,
+         co2_score = n_total_co2,
+         health_impact_score = n_health_emissions_cost,
+         electricity_score = n_electricity_generation,
+         climate_risk_score = n_climate_risk)  %>% 
+  mutate_if(is.numeric,round, digits = 4) %>% 
+  select(cbsa, msa, anchor_city, state, stateabbrv, ends_with("score"))
+
+# weights for each criteria must be in this order
+equity_weights <- c(0.18, # real estate weight
+                    0.0, # landlord weight
+                    0.10, # financial weight
+                    0.19, # co2 emissions avoided weight
+                    0.17, # health impacts weight
+                    0.14, # electricity weight
+                    0.22) # climate risk weight
+
+client_weights <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)
+
+test <- calculate_wt_criteria(unwt_criteria, equity_weights)
+
+client_test <- calculate_wt_criteria(unwt_criteria, client_weights)
+```
+
+```{r}
+# data viz
+client_score_tidy <- client_test  %>% 
+  pivot_longer(cols = c(6:12), # selecting criteria cols
+               names_to = "criteria",
+               values_to = "wt_criteria_score") %>% 
+  unite("msa_state", anchor_city, stateabbrv, sep = ", ") %>% 
+ filter(cbsa != "38860")   # removing portland (negative number)
+
+ggplot(data = client_score_tidy, aes(y = reorder(msa_state,  wt_criteria_score, FUN=sum), x = wt_criteria_score)) +
+    geom_col(aes(fill = factor(criteria, levels = c("Climate Risk Avoided", "Avoided CO2 Emissions", "Electricity Generation", "Solar Financials", "Real Estate", "Landlord Policy", "Avoided Health Costs")))) +
+  theme_bw() +
+  labs(x = "Investment Favorabliity Score",
+       y = " ",
+       title = "") +
+  guides(fill=guide_legend(title="Criteria (weighted)")) +
+   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
+      breaks = c("Climate Risk Avoided", "Avoided CO2 Emissions", "Electricity Generation", "Solar Financials", "Real Estate", "Landlord Policy", "Avoided Health Costs")) +
+        scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1)) 
 ```
@@
 ```
-
-# error check so weights sum to 1
-# incorporate shiny??
 
@@
 ```{r client data viz}
-tiff(filename = "client_criteria_barchart.tiff",width = 8, height = 6, units = "in", res = 300, pointsize = 12)
+#tiff(filename = "client_criteria_barchart.tiff",width = 8, height = 6, units = "in", res = 300, pointsize = 12)
 ggplot(data = client_score_tidy, aes(y = reorder(msa_state,  wt_criteria_score, FUN=sum), x = wt_criteria_score)) +
@@
 #+ theme(legend.position = "none")
-dev.off()
+#dev.off()
 ```
@@
 
-```{r COMBINED BAR CHARTS}
+```{r COMBINED BAR CHARTS, eval=FALSE}
 library(patchwork)

### 2023-05-12 solar_fam_shiny/census_api_test.R
file organization

@@ -1,0 +1,84 @@
+
+library(tidyverse)
+library(censusapi)
+library(jsonlite)
+library(tidycensus)
+
+
+api_key <- "ebab921b3002df9b71881ad6c426f34281ce0e11"
+
+# Set API key
+Sys.setenv(CENSUS_KEY="ebab921b3002df9b71881ad6c426f34281ce0e11")
+
+# Define variables
+variables <- c(
+  tot_pop = "B01003"),
+  median_home_USD = "B25077_001")
+ # unemployment_rate = "DP03_0009E"
+
+
+# Get ACS data for metro areas
+acs_data <- get_acs(
+  geography = "B01003",
+  variables = variables,
+  year = 2021
+)
+
+# Rename columns
+colnames(acs_data) <- c("metro_area", "median_income", "population", "median_home_value")
+
+
+# Rename columns
+
+# Print the ACS data
+print(acs_data)
+
+
+############################
+
+# Add key to .Renviron
+Sys.setenv(CENSUS_KEY="ebab921b3002df9b71881ad6c426f34281ce0e11")
+# Reload .Renviron
+readRenviron("~/.Renviron")
+# Check to see that the expected key is output in your R console
+Sys.getenv("CENSUS_KEY")
+
+census_api_key <- "ebab921b3002df9b71881ad6c426f34281ce0e11"
[…]

### 2023-05-12 src/00_demographic_data_wrangling.Rmd
uploading instructional documentation for dryad

--- 2023-04-27 src/demo_data_wrangling.Rmd
@@
 # read in the data MSA population data from Census
-msa_pop_2021_data <- read_csv(here("data","raw", "msa_2021_pop_data.csv")) %>% 
+msa_pop_2021_data <- read_csv(here("data/raw/population_est2021.csv")) %>% 
   clean_names() %>% # data wrangling to make column headers in lower_snake_case
@@
 
-# create object with viable MSAs 
+# create object with viable MSAs for client 
 potential_MSAs <- msa_pop_2021_data %>% 
@@
 
+unfiltered_msas <- msa_pop_2021_data %>% 
+  filter(pop_percent_change>0.5) %>% 
+  select(1:3, pop_percent_change)
 ```
@@
 
-```{r population change visual, fig.height = 7, fig.align='center', eval = FALSE}
-# create plot of MSAs sorted grom highest population growth to lowest
-ggplot(data = positive_MSA_pop_change, aes(x = pop_percent_change20_21, y = reorder(MSA, pop_percent_change20_21))) + 
-  geom_point(size = 3, color = "darkgreen") + 
-  theme_minimal(base_size = 12.5)
-
-```
-
 ### Household Income
@@
 ```{r household income data wrangling}
-MSA_income_data <- read_csv(here("data","raw", "2021_income_data.csv")) %>% 
+MSA_income_data <- read_csv(here("data/raw/income_data2021.csv")) %>% 
   row_to_names(1) %>% 
@@
   mutate(cbsa = as.numeric(cbsa)) %>% # change column class from chr to dbl
-  inner_join(potential_MSAs)
+  select(cbsa, median_income)
 
 # join household median income to existing msa dataframe
-potential_MSAs <- potential_MSAs %>% add_column(med_income = MSA_income_data$median_income)
+potential_MSAs <- potential_MSAs %>% 
+  inner_join(MSA_income_data)
 
+unfiltered_msas <- unfiltered_msas %>% inner_join(MSA_income_data)
 ```
@@
 ```{r, employment data wrangling}
-unemployment_data <- read_csv(here("data", "raw", "civil_labor_force_BLS.csv")) %>% #rename col headers
+unemployment_data <- read_csv(here("data/raw/unemployment_data.csv")) %>% #rename col headers
   clean_names() %>% 
@@
          unemployment = x9,
-         unemployment_rate = x10)%>% 
-  mutate(employment_rate = (employment/civilian_labor_force)*100) # calculate
+         unemployment_rate = x10) %>% 
+  drop_na() %>% 
+  mutate(employment_rate = (employment/civilian_labor_force)*100) %>%  # calculate
+  select(cbsa, area, year, month, employment_rate)
 
@@
   mutate(cbsa = as.numeric(cbsa)) %>% 
-  right_join(potential_MSAs) %>% 
-  group_by(MSA, year) %>% 
+  group_by(area, cbsa, year) %>% 
+ #  right_join(potential_MSAs) %>% 
   summarize(avg_employment_rate = mean(employment_rate)) %>% 
   mutate(avg_emp_rate = (avg_employment_rate-lag(avg_employment_rate))/lag(avg_employment_rate)*100, 
-          avg_emp_rate = round(avg_emp_rate, digits = 4)) %>% 
-  drop_na() %>% 
-  mutate(emp_pos = case_when( 
-    avg_emp_rate > 0 ~ "TRUE",
-    avg_emp_rate < 1 ~ "FALSE"))
+          avg_emp_change = round(avg_emp_rate, digits = 4)) %>% 
+  filter(year=="2021") %>% 
+  select(cbsa, area, avg_emp_change)
 
-MSA_employ_rate <- MSA_employment %>% 
-  filter(year == "2021")
+client_MSAs <-potential_MSAs %>% 
+  left_join(MSA_employment)
 
-potential_MSAs$employ_change <- MSA_employ_rate$avg_emp_rate
+unfiltered_msas <- unfiltered_msas %>% 
+  left_join(MSA_employment)
 
@@
 # data wrangling 2021 median rent
-MSA_median_rent2021_data <- read_csv(here("data", "raw", "2021MSA_medianrent.csv")) %>% 
+median_rent2021_data <- read_csv(here("data/raw/median_rent2021.csv")) %>% 
   separate(col = cbsasub21,
@@
         median_rent2021 = as.numeric(rent50_1)) %>% # select median rent for 1 bedroom apartments
-  select(cbsa, median_rent2021) %>% 
-  right_join(potential_MSAs)
+  select(cbsa, median_rent2021)
 
 # 2022 median rent
-MSA_median_rent2022_data <- read_csv(here("data", "raw", "2022MSA_median_rent.csv")) %>% 
+median_rent2022_data <- read_csv(here("data", "raw", "median_rent2022.csv")) %>% 
   separate(col = cbsasub22,
@@
          median_rent2022 = as.numeric(rent_50_1)) %>% 
-  select(cbsa, median_rent2022) %>% 
-  right_join(potential_MSAs)
+  select(cbsa, areaname22, median_rent2022) %>% 
+  left_join(median_rent2021_data)
 
-potential_MSAs <- potential_MSAs %>% 
-  add_column(med_rent2021 = MSA_median_rent2021_data$median_rent2021, # add rent data
-             med_rent_2022 = MSA_median_rent2022_data$median_rent2022)
+client_msas <-client_MSAs %>% 
+  left_join(median_rent2022_data)
+
+unfiltered_msas <- unfiltered_msas %>% 
+  left_join(median_rent2022_data)
 
@@
 ```{r house price data}
-metro_house_cost_data <- read_csv(here("data","raw", "metro_home_costs", "metro-home-prices-2022-11-10.csv")) %>% 
+metro_house_cost_data <- read_csv(here("data/raw/metro-home-prices-2022.csv")) %>% 
   row_to_names(3) %>% # make sure column headers are correct
@@
         house_price_2021 = x2021 * 1000) %>% # data is proivded as thousands, so multiply to get correct ratio
-  right_join(potential_MSAs) %>% 
   mutate(house_price_2021 = ifelse(cbsa == "12260", "284900", house_price_2021), 
-        house_price_2021 = ifelse(cbsa == "32580", "259000", house_price_2021))
+        house_price_2021 = ifelse(cbsa == "32580", "259000", house_price_2021)) %>% 
+  select(cbsa, house_price_2021)
 
-potential_MSAs$house_price_2021 <- metro_house_cost_data$house_price_2021
+potential_MSAs <- client_msas %>% 
+  left_join(metro_house_cost_data)
+
+unfiltered_msas <- unfiltered_msas %>% 
+  left_join(metro_house_cost_data)
 
@@
 potential_MSAs <- potential_MSAs %>% 
-  rename(pop_growth = pop_percent_change20_21)
+  select(-area, -areaname22)
+
+unfiltered_msas <- unfiltered_msas %>% select(-area, -areaname22)
 
@@
 
-write.xlsx(potential_MSAs, here("data/intermediate/demographic.xlsx"), sheetName = "Sheet1")
+#write_csv(potential_MSAs, here("data/intermediate/demographic_data.csv"))
 
-#gs4_create("real_estate_inputs")
-#output data into google sheet
-#sheet_write(potential_MSAs, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "MSA_demographics")
 ```

### 2023-05-12 src/01_metric_normalization.Rmd
uploading instructional documentation for dryad

---
title: "landlord"
author: "Virginia Pan"
date: "10/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(googlesheets4)
library(janitor)
library(tidyverse)
library(DT)
library(readxl)
library(xlsx)
library(here)
source(here("R/normalize.R"))

```

# Calculate and Normalize Real Estate Metrics

```{r}
## read in real estate and demographic data 
city_occupancy_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990") %>% 
  clean_names() %>% 
  rename(occ_rate2021 = percent_occupancy_2021)

anchor_city <- read.xlsx(here("data/intermediate/demographic_data.xlsx"), sheetName="anchor_city") %>% 
  clean_names() %>% 
   inner_join(city_occupancy_data, by = join_by(cbsa== cbsa, anchor_city==city)) %>% 
  select(cbsa, anchor_city, occ_rate2021)

MSA_demographic_data <- read_csv(here("data/intermediate/demographic_data.csv")) %>% 
  clean_names() %>% 
  left_join(anchor_city)

# calculate real estate metrics
real_estate_metrics<- MSA_demographic_data %>% 
    # calculate rent change
  mutate(cbsa = factor(cbsa),
         rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100,
         house_price_2021 = as.numeric(house_price_2021),
         income_to_homeprice = median_income/house_price_2021,
         rent_to_income = (median_rent2021*12)/median_income) %>%  # multiply monthly rent x 12 for annual rent!
  select(!c(na, popestimate2021, median_rent2021, median_rent2022, house_price_2021, median_income)) %>% 
  select(cbsa, anchor_city, msa, state, pop_growth, avg_emp_change, 7:10)

## normalization
n_real_estate_metrics <- real_estate_metrics %>%
  # normalize cols and rename columns to start with n
  mutate(across(5:10, normalize, .names = "n_{.col}")) %>% 
  mutate(n_income_to_homeprice = 1- n_income_to_homeprice, # subtracted from 1 so higher values are more favorable
         n_rent_to_income = 1 - n_rent_to_income) %>% 
  select(cbsa, msa, anchor_city, state, starts_with("n_"))

```

# Normalizing Landlord Metrics

```{r}
MSA_landlord_data <- read_csv(here("data/intermediate/landlord_data.csv")) %>% 
  clean_names()

#security_deposit_limit
landlord_metrics <- MSA_landlord_data %>% 
  mutate(
    security_deposit_limit_num = case_when(
      security_deposit_limit == "N" ~ 1.0, 
      security_deposit_limit == "1_month_rent" ~ 0.0,
      security_deposit_limit == "2_months_rent" ~ 0.5)) %>% 
    #"invert" normalized score for eviction notice days
#1 - normalized score (since more eviction notice is worse for landlords)
  mutate(n_eviction_notice = 1- normalize(eviction_notice_days),
         n_security_deposit = normalize(security_deposit_limit_num))

# assign numeric values to security deposit limit for all states
#state_landlord_policy <- state_landlord_policy %>% 
 # mutate(security_deposit_limit = case_when(
  #    deposit_limit == "N" ~ 1.0, 
   #   deposit_limit == "1_months_rent" ~ 0.0,
    #  deposit_limit == "1_month_rent" ~ 0.0,
    #  deposit_limit == "1.5_months_rent" ~ 0.25,
    #  deposit_limit == "2_months_rent" ~ 0.5,
    #  deposit_limit == "2.5_months_rent" ~ 0.75,
    #  deposit_limit == "3_months_rent" ~ 0.75)) %>% 
   #"invert" normalized score for eviction notice days
#1 - normalized score (since more eviction notice is worse for landlords)
  # mutate(n_eviction_notice = 1- normalize(eviction_days),
  #       n_security_deposit = normalize(security_deposit_limit))
```

# reOpt Metric data wrangling and normalization

```{r}
# read in reopt data (for STCOFIPS codes)
reopt_metrics <- read_xlsx(here("data/intermediate/reopt_data_outputs.xlsx"), sheet="reopt_msa_outputs") %>% 
  clean_names() %>% 
  drop_na() %>% 
  rename(
  electricity_generation = annual_renewable_electricity_percent_of_total,
  total_co2 = lifecycle_total_t_co2,
  health_emissions_cost = lifecycle_costs_of_health_emissions_difference_from_bau) %>% 
  select(1:5, electricity_generation, total_co2, health_emissions_cost, irr)

# normalize crit
n_reopt_metrics <- reopt_metrics %>% 
mutate(across(electricity_generation:irr, normalize, .names = "n_{.col}"),
       n_health_emissions_cost = 1-n_health_emissions_cost,
       n_total_co2 = 1 - n_total_co2)

#write_csv(n_reopt_metrics, here("data/intermediate/reopt_metrics.csv"))

```

# reOpt Metric data wrangling and normalization

```{r}
#read in National Risk Index (NRI) county data
climate_risk_metrics <- read_csv(here("data/raw/NRI_Table_Counties.csv")) %>% 
  clean_names() %>% 
  select(state, stateabbrv, county, countytype, risk_score) %>% 
  inner_join(reopt_metrics, by = join_by(stateabbrv==state, county==county)) %>% 
  filter(countytype == "County") # to remove richmond city

# normalize climate risk score
n_climate_risk <- climate_risk_metrics %>% 
  mutate(n_climate_risk = 1- normalize(risk_score)) %>%  #invert normalized scores so that places with less climate risk have a better score
  select(cbsa, city_msa, state, stateabbrv, risk_score, n_climate_risk)

 #write_csv(n_climate_risk, here("data/intermediate/climate_risk_metrics.csv"))
```

### 2023-05-12 src/reopt_api.Rmd
uploading instructional documentation for dryad

@@
 library(jsonlite)
+library(openxlsx)
+library(writexl)
+library(here)
+library(tidyverse)
+library(janitor)
 ```
 
-### Obtain MSA zip code information
-```{r}
+### Obtain MSA city information
+Only used once to generate table and output results into data/intermediate folder to decrease runtime for the APi
+
+```{r, eval= FALSE}
 token <- "eyJ0eXAiOiJKV1QiLCJhbGciOiJSUzI1NiIsImp0aSI6ImNmMmU0Yzg1ZWZlMDNiYTkxNTA4YzdlNWJiYjE1NjVkNzIyNjg4MzgwYTE4OWNjNmJkOWRjNjFiZWUyZjIwYzg0ZmJjNmZkOTc0MDg4YzRmIn0.eyJhdWQiOiI2IiwianRpIjoiY2YyZTRjODVlZmUwM2JhOTE1MDhjN2U1YmJiMTU2NWQ3MjI2ODgzODBhMTg5Y2M2YmQ5ZGM2MWJlZTJmMjBjODRmYmM2ZmQ5NzQwODhjNGYiLCJpYXQiOjE2ODI0ODQyOTYsIm5iZiI6MTY4MjQ4NDI5NiwiZXhwIjoxOTk4MTAzNDk2LCJzdWIiOiI0ODUyMyIsInNjb3BlcyI6W119.G7NL2Khpo602p_fzk6EiAZ0-KdXsVW-IO9mUqwHem0YFLaTQZezotmLqXmbcTi7BxRzc8UHNYBw415CCzKo81w"
@@
 # zip-cbsa : type3
-zip_url <- "https://www.huduser.gov/hudapi/public/usps?type=3&query=All"
+#zip_url <- "https://www.huduser.gov/hudapi/public/usps?type=3&query=All"
 
@@
 # make data frame
-cbsa_city_data <- data.frame("geoid"= response_cbsa$data$results$geoid,
-                             "cbsa"= response_cbsa$data$results$cbsa,
-                            "city"= response_cbsa$data$results$city)
+cbsa_city_data <- data.frame("cbsa"= response_cbsa$data$results$cbsa,
+                            "city"= response_cbsa$data$results$city,
+                            "state"= response_cbsa$data$results$state)  
+# output into excel
+#write.xlsx(cbsa_city_data, here("data/intermediate/cbsa_city_data.xlsx"), sheetName = "Sheet1")
+```
 
-zip_response <- GET(url = zip_url ,add_headers("Authorization" = paste("Bearer", token)))
-# Extract the response content as JSON and convert it to a list
-response_zip_content <- content(zip_response, as="text")
-response_list <- fromJSON(response_zip_content)
+# read in US city with longitude and latitude data from https://simplemaps.com/data/us-cities
+```{r}
+uscities <- read_csv(here("data/raw/uscities.csv")) %>% 
+  mutate(city = tolower(city)) %>% 
+  select(state_id, city, lat, lng, population, density, zips) %>% 
+  group_by(state_id, city)
 
-zip_data <- data.frame("geoid"= response_list$data$results$geoid,
-                            "zipcode"= response_list$data$results$zip,
-                             "state"= response_list$data$results$state)
-# combine data sets
-zip_cbsa_data <- cbsa_city_data %>% 
-  right_join(zip_data)
+city_cbsa_api_data <- read.xlsx(here("data/intermediate/cbsa_city_data.xlsx")) %>% 
+   mutate(city = tolower(city)) %>% 
+  group_by(state, city)
 
@@
 
+# create cbsa and city data frame that was extracted using hud API
+```{r}
+cbsa_city_lat_long <- city_cbsa_api_data %>% 
+    filter(cbsa != "99999") %>% 
+  inner_join(uscities, by = join_by(state == state_id, city == city)) %>%    # removes cities without a cbsa
+  distinct()
+
+one_city_msa_data <- cbsa_city_lat_long %>% 
+  group_by(cbsa) %>% 
+  summarize(max_pop = max(population)) %>% 
+  left_join(cbsa_city_lat_long, by=join_by(cbsa== cbsa, max_pop==population), keep=FALSE) %>% 
+  # select one zipcode to run in open ei api to get utility data
+  separate(col = zips,
+           sep = " ",
+           into = "zips",
+           extra = "drop")
+
+```
+
+# get population estimates for each Metro area in 2021  using census api
 ```{r polulation estimates}
@@
 
-msa_zip_pop_est_data <- pop_list %>% 
-  left_join(cbsa_city_data) %>% 
-  inner_join(zip_data)
-
-```
-
-# reopt inputs
-```{r}
-# create list of default inputs for each city
-inputs <- list(
-  offtaker_tax_pct = 0, # host effective tax rate
-  co2_cost_us_dollars_per_tonne = 185,
-  module_type = 1,
-  dc_ac_ratio = 1.35,
-  losses = 0.05,
-  federal_itc_pct = 0.3,
-  location = "roof", # restrict projects to roof
-  can_net_meter = TRUE
-
-)
-```
-
-# Set the payload for the job submission request for reopt
-```{r}
-# Set the payload for the job submission request
-payload <- list(
-  "Scenario" = list(
-    "Site" = list(
-      "longitude" = -118.1164613,
-      "latitude" = 34.5794343,
-      "roof_squarefeet" = 5000.0,
-      "PV" = list(
-        "installed_cost_us_dollars_per_kw" = 2000.0,
-        "federal_rebate_us_dollars_per_kw" = 350.0,
-        # changes from default
-        "module_type" = 1,
-        "dc_ac_ratio" = 1.35, # changed default
-        "losses" = 0.05,   # changed default
-        "federal_itc_pct" = 0.3,  # changed default
-        "location" = "roof", # restrict projects to roof
-        "can_net_meter" = TRUE
-      ),
-      "LoadProfile" = list(
-        "doe_reference_name" = "MidriseApartment",
-        "annual_kwh" = 10000000.0,
-        "city" = "LosAngeles"
-      ),
-      "ElectricTariff" = list(
-        "urdb_label" = "5ed6c1a15457a3367add15ae",
-        "add_tou_energy_rates_to_urdb_rate" = TRUE
-      ),
-      "Financial" = list(
-        "escalation_pct" = 0.026,
-        "offtaker_discount_pct" = 0.081,
-        "owner_discount_pct" = 0.081,
-        "analysis_years" = 20,
-        "offtaker_tax_pct" = 0.4,
-        "owner_tax_pct" = 0.4,
-        "om_cost_escalation_pct" = 0.025
-      )
-    )
-  )
-)
+MSA_w_lat_lng <- pop_list %>% 
+  inner_join(cbsa_city_lat_long, by = join_by(cbsa==cbsa))
 
@@
 
-startdates <- EIutility_df$start_date
-start_dates_rev <- lapply(startdates, function(x) as.Date(x/86400, origin="1970-01-01"))
-
 ```
@@
 ```{r}
-## function to select the most recent utiilty rate information for a given area code
-get_zip_utility <- function(zipcode){
-  openEI_start_url <- "https://api.openei.org/utility_rates?version=latest&format=json&sector=Residential&detail=full&approved=true&address="
-openEI_api_end <- "&api_key=rU6MEqhK9OJ7obAnmcbzYhpFOenfBM5Q8GF0yJFA"
-zipcode=zipcode
+# Define a vector of zip codes
+zipcodes <- msa_zip_pop_est_data
+# break up zipcodes to run in openEI  
+zipcodes1 <- zipcodes[1:200,]
 
-  openEI_url <- paste0(openEI_start_url, zipcode, openEI_api_end)
-  response <- GET(openEI_url)
-  # extract content and parse as JSON
-  content <- content(EIresponse, "text")
-  json <- fromJSON(content)
+c(zipcodes1, zipcodes2, zipcodes3, zipcodes4, zipcodes5, zipcodes6)
 
-  utility_df <- data.frame(
-  # repeat zipcode in each row for each utility observation
-  "zipcode" = rep(zipcode, each=nrow(json$items$label)),
-  "start_date" = utility_data$items$startdate,
-  "label" = json$items$label,
-    "utility" = json$items$utility,
-    "rate_name" = json$items$name)
-
-  # Find index of row with most recent start date value
-  max_index <- which.max(utility_df$start_date)
-
-# Extract row with greatest seconds value
-  max_row <- utility_df[max_index, ]
-
-}
+OpenEI_zip_result <- lapply(zipcodes1$zipcode, get_zip_utility)
 
@@
 
+openei_url <- paste0("https://api.openei.org/geocoding?version=latest&format=json&postalcode=", zipcode, "&countrycode=US", "&api_key=U6MEqhK9OJ7obAnmcbzYhpFOenfBM5Q8GF0yJFA")
+response <- GET(openei_url)
+
+# Extract the content and parse as JSON
+content <- content(response, "text")
+json <- fromJSON(content)
+
 ```
 
+# Set the payload for the job submission request for reopt
 ```{r}
-# Define a vector of zip codes
-zipcodes <- c("90210", "10001", "60601", "90245", "90001", "90250", "90260", "90278", "90501", "90502")
 
-# Initialize an empty list to store the results
-utility_list <- list()
+zip_utility_info <- get_zip_utility(test$zips)
+zip_EI_label <- zip_utility_info$label
 
-# Loop through each zip code and call the get_zip_utility function
-for (zipcode in zipcodes) {
-  utility_data <- get_zip_utility(zipcode)
-  utility_list[[zipcode]] <- utility_data
-}
+test <- one_city_msa_data %>% filter(city == "des moines")
 
-# Combine the results into a single data frame
-utility_df <- do.call(rbind, utility_list)
+payload <- list(
+  "Scenario" = list(
+    "Site" = list(
+      "address"= test$zips,
+      "longitude"= test$lng,
+      "latitude" = test$lat,
+
+      "PV" = list(# changes from default
+    "module_type" = 1, # premium
+    "dc_ac_ratio" = 1.35, # changed default
+    "losses" = 0.05,   # changed default
+    "federal_itc_pct" = 0.3,  # changed default
+     "location" = "roof", # restrict projects to roof
+     "can_net_meter" = TRUE
+      ),
+      "LoadProfile" = list(
+        "doe_reference_name" = "MidriseApartment",
+        "city" = test$zips
+      ),
+      "ElectricTariff" = list(
+        "urdb_label" = zip_utility_info$label,
+        "net_metering_limit_kw" = 1000
+      ),
+      "Financial" = list(
+        "offtaker_tax_pct" = 0, # host effective tax rate
+        "owner_tax_pct" = 0.0,
+        "co2_cost_us_dollars_per_tonne" = 185
+      ))
+  )
+)
 
@@
 ```{r STUFF TO WORK ON}
+net_metering_limit_kw # by state
 
-net_metering_limit_kw # by state
 ```
@@
 
-job_url <- paste0(url, "?api_key=", api_key)
+job_url <- paste0(url, api_key)
 ```
@@
 output_df <- data.frame(
-  IRR = response_json$outputs$Scenario$Site$Financial$irr_pct,
+  #IRR = response_json$outputs$Scenario$Site$Financial$irr_pct,
   Health_Emissions_Cost = response_json$outputs$Scenario$Site$lifecycle_emissions_cost_Health_bau,
@@
 
+host/stable/annual_kwh
+doe_response <- GET(url = results_url, query = list(api_key = api_key), timeout(30))
 ```
@@
 ```{r}
-#PV & wind space available:  ❌ Land only ✅ Roofspace only ❌ Land & roofspace
-
-#PV System Characteristics
-ModuleType= Premium
-DCACratio = 1.35
-System_losses = 5
-FederalIncentive= 30
-
-```
-
-```{r}
 # test scenario post

### 2023-05-12 src/weight_inputs.Rmd
uploading instructional documentation for dryad

---
title: "Inputing Weights"
author: "Grace Bianchi"
date: "2023-05-04"
output: 
  html_document:
    code_folding: hide
runtime: shiny
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(xlsx)
library(readxl)
```

```{r echo=FALSE}

unfiltered_msas <- 
real_estate_metrics <- read_xlsx(here("data/intermediate/metric_data.xlsx"), sheet="real_estate_metrics") %>% 
  select(cbsa, anchor_city, msa, starts_with("n_")) %>% 
  mutate(cbsa = factor(cbsa))

landlord_metrics <- read_xlsx(here("data/intermediate/metric_data.xlsx"), sheet="landlord_metrics") %>% 
  select(cbsa, starts_with("n_"))%>% 
  mutate(cbsa = factor(cbsa))

climate_risk_metrics <- read_csv(here("data/intermediate/climate_risk_metrics.csv")) %>% 
  mutate(cbsa = factor(cbsa)) %>% 
  select(!risk_score)

reopt_metrics <- read_xlsx(here("data/intermediate/metric_data.xlsx"), sheet="reopt_metrics") %>% 
  select(cbsa, starts_with("n_")) %>% 
  mutate(cbsa = factor(cbsa))

```

## Input Weights

### Metric Inputs for Real Estate and Landlord Criteria

```{r real estate and landlord metric weights, echo=FALSE}

real_estate_weights <- c(0.35, # pop change weight
                         0.20, # employment change weight
                         0.15, # occupancy weight
                         0.10, # rent change weight
                         0.10, # rent to income weight
                         0.10) # income to homeprice weight

#input landlord weights for each indicator in the following order: eviction notice days, security deposit limit
# Weights must sum to 1
landlord_weights <- c(0.8, # eviction notice weight
                      0.2) # security deposit limit weight

```

```{r criteria weights, echo=FALSE}
# weights for each criteria must be in this order
equity_weights <- c(0.18, # real estate weight
                    0.0, # landlord weight
                    0.10, # financial weight
                    0.19, # co2 emissions avoided weight
                    0.17, # health impacts weight
                    0.14, # electricity weight
                    0.22) # climate risk weight

client_weights <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)

```

# Weight Metrics

```{r echo=FALSE}
# weight real estate metrics
wt_real_estate_metrics <- real_estate_metrics %>% 
  mutate(wt_pop_change = n_pop_growth* real_estate_weights[1],
         wt_employ_change = n_avg_emp_change* real_estate_weights[2],
         wt_occ = n_occ_rate2021* real_estate_weights[3],
         wt_rent_change = n_rent_change* real_estate_weights[4],
         wt_rent_income = n_rent_to_income* real_estate_weights[5],
         wt_income_homeprice = n_income_to_homeprice* real_estate_weights[6]) %>% 
  select(!starts_with("n_")) %>% 
  mutate(real_estate_score = apply(.[,4:9], 1, sum)) %>% 
  mutate_if(is.numeric,round, digits = 4)

# wt_landlord metrics
wt_landlord_metrics <- landlord_metrics %>%
      # multiply each criteria by respective weight
      mutate(wt_eviction = n_eviction_notice*landlord_weights[1],
             wt_deposit = n_security_deposit*landlord_weights[2]) %>% 
      # for loop to sum weighted columns for each row
      mutate(landlord_score = apply(.[, 4:5], 1, sum)) %>% 
   mutate_if(is.numeric,round, digits = 4)  %>% 
  select(!starts_with("n_"))

```

# Weight Criteria

```{r weighted score function, echo=FALSE}
  calculate_wt_criteria <- function(data, weights){
    wt_criteria <- data %>%
      # multiply each criteria by respective weight
      mutate(wt_real_estate = real_estate_score * weights[1],
             wt_landlord_policy = landlord_score * weights[2], 
             wt_solar_irr = financial_score * weights[3],
            wt_co2 = co2_score * weights[4], 
             wt_health_impacts = health_impact_score * weights[5],
             wt_electricity = electricity_score * weights[6],
             wt_climate_risk = climate_risk_score * weights[7]) %>% 
      # select weighted columns and rename them
      select(!c(ends_with("_score"))) %>%
      # create column for investment score
      mutate(investment_score = 0) %>%
      # for loop to sum weighted columns for each row
      mutate(investment_score = apply(.[, 6:12], 1, sum)) %>% 
      rename("Real Estate"="wt_real_estate",
             "Landlord Policy"="wt_landlord_policy",
             "Electricity Generation"="wt_electricity", 
             "Avoided CO2 Emissions"="wt_co2",
             "Climate Risk Avoided"="wt_climate_risk",
             "Avoided Health Costs"="wt_health_impacts", 
             "Solar Financials"="wt_solar_irr")
  }

```

```{r echo=FALSE}
unwt_criteria <- wt_real_estate_metrics %>% 
  inner_join(wt_landlord_metrics) %>% 
  select(!starts_with("wt_")) %>% 
  inner_join(reopt_metrics) %>% 
  inner_join(climate_risk_metrics, by = join_by(cbsa==cbsa)) %>% 
  select(!city_msa) %>% 
    rename(financial_score = n_irr,
         co2_score = n_total_co2,
         health_impact_score = n_health_emissions_cost,
         electricity_score = n_electricity_generation,
         climate_risk_score = n_climate_risk)  %>% 
  mutate_if(is.numeric,round, digits = 4) %>% 
  select(cbsa, msa, anchor_city, state, stateabbrv, ends_with("score"))

equity_test <- calculate_wt_criteria(unwt_criteria, equity_weights)

client_test <- calculate_wt_criteria(unwt_criteria, client_weights)
```

# Visualize results
```{r echo=FALSE}
# data viz
client_score_tidy <- client_test  %>% 
  pivot_longer(cols = c(6:12), # selecting criteria cols
               names_to = "criteria",
               values_to = "wt_criteria_score") %>% 
  unite("msa_state", anchor_city, stateabbrv, sep = ", ") %>% 
 filter(cbsa != "38860")   # removing portland (negative number)

ggplot(data = client_score_tidy, aes(y = reorder(msa_state,  wt_criteria_score, FUN=sum), x = wt_criteria_score)) +
    geom_col(aes(fill = factor(criteria, levels = c("Climate Risk Avoided", "Avoided CO2 Emissions", "Electricity Generation", "Solar Financials", "Real Estate", "Landlord Policy", "Avoided Health Costs")))) +
  theme_bw() +
  labs(x = "Investment Favorabliity Score",
       y = " ",
       title = "") +
  guides(fill=guide_legend(title="Criteria (weighted)")) +
   scale_fill_manual(values = c("palevioletred1","orange",  "gold", "#44AA99", "dodgerblue4", "#BC80BD", "#6495ED"),
      breaks = c("Climate Risk Avoided", "Avoided CO2 Emissions", "Electricity Generation", "Solar Financials", "Real Estate", "Landlord Policy", "Avoided Health Costs")) +
        scale_x_continuous(limits = c(0, 0.8), expand = c(0,0), breaks=seq(0, 0.8, 0.1)) 
```

## Inputs and Outputs

## Embedded Application

```{r tabsets, echo=FALSE}
shinyAppFile(
  here("solar_fam_shiny/app.R"),
  options = list(
    width = "100%", height = 550
  )
)
```

### 2023-05-12 src/03_input_weights.Rmd
added psuedocode

--- 2023-05-12 src/weight_inputs.Rmd
@@
 ```{r echo=FALSE}
-
-unfiltered_msas <- 
-real_estate_metrics <- read_xlsx(here("data/intermediate/metric_data.xlsx"), sheet="real_estate_metrics") %>% 
+# list of MSAs without filtering population size
+unfiltered_msas <- read_csv(here("data/intermediate/unfilter_msas.csv"))
+real_estate_metrics <- read_csv(here("data/intermediate/demographic_data.csv")) %>% 
   select(cbsa, anchor_city, msa, starts_with("n_")) %>% 
@@
 
-landlord_metrics <- read_xlsx(here("data/intermediate/metric_data.xlsx"), sheet="landlord_metrics") %>% 
+landlord_metrics <- read_csv(here("data/intermediate/landlord_data.csv")) %>% 
   select(cbsa, starts_with("n_"))%>% 
@@
 
-reopt_metrics <- read_xlsx(here("data/intermediate/metric_data.xlsx"), sheet="reopt_metrics") %>% 
+reopt_metrics <- read_csv(here("data/intermediate/reopt_metrics")) %>% 
   select(cbsa, starts_with("n_")) %>% 
@@
                     0.22) # climate risk weight
-
-client_weights <- c(0.18, 0.07, 0.11, 0.07, 0.0, 0.07, 0.50)
 
@@
 
-client_test <- calculate_wt_criteria(unwt_criteria, client_weights)
 ```
@@
 ```
-
-## Inputs and Outputs
-
-## Embedded Application
-
-```{r tabsets, echo=FALSE}
-shinyAppFile(
-  here("solar_fam_shiny/app.R"),
-  options = list(
-    width = "100%", height = 550
-  )
-)
-```

### 2023-05-18 src/03_input_weights.Rmd
updating data folder structures

@@
 unfiltered_msas <- read_csv(here("data/intermediate/unfilter_msas.csv"))
-real_estate_metrics <- read_csv(here("data/intermediate/demographic_data.csv")) %>% 
+
+real_estate_metrics <- read_csv(here("data/intermediate/n_real_estate_metrics.csv")) %>% 
   select(cbsa, anchor_city, msa, starts_with("n_")) %>% 

### 2023-06-07 solar_fam_shiny/app.R
creating shiny to input weights

@@ -1,40 +1,39 @@
-
 library(shiny)
 library(ggplot2)
 library(tidyverse)
 library(bslib) # shiny theme
 library(tidycensus)
 library(googlesheets4)
-library(tigris) # for shape files
-options(tigris_use_cache = TRUE)
-library(tmap) # interactive map
-tmap_options(check.and.fix = TRUE)
 library(sf) # map states and US
-library(shinydashboard)
+library(readxl)
+library(here)
+library(tmap)
+library(tigris)
+library(tmaptools)
+
+cbsa_geom <- core_based_statistical_areas(cb = TRUE) %>% 
+  mutate(cbsa = factor(CBSAFP))
 
 
-real_estate_metrics2021 <- read_sheet("https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=519968233") %>% 
-  rename(pop_estimate = popestimate2021)
-criteria_unweighted <- read_sheet("https://docs.google.com/spreadsheets/d/1yqjhJvXUcEiC3qiYWkKNlF6NNpYXWg15zZHwmwpS5Q0/edit#gid=1307699202", sheet = "unwt_criteria_scores") %>% 
-  mutate(cbsa = as.character(cbsa)) %>% 
-  filter(city_msa != "Portland") 
+states_geom <- tidycensus::state_laea    #read in the dataset
 
-#census_api_key("ebab921b3002df9b71881ad6c426f34281ce0e11", overwrite = "TRUE")
 
-# tidycensus package metro area shape files
-msa_shp <- core_based_statistical_areas() %>% 
-  janitor::clean_names() %>% 
-  select(!c(csafp, lsad, cbsafp)) #4269
+# state and U.S. map coordinates
+states <- map_data("state") %>% 
+  st_as_sf(coords = c("long", "lat"))
 
-states_4269 <- st_transform(state_laea, st_crs(msa_shp))
+usa<- map_data("usa") %>% 
+  st_as_sf(coords = c("long", "lat"))
 
-# state shape files from sf packages
-# state_shp <- states(class = "sf", progress_bar = TRUE, shift_geo = TRUE) %>% 
-#   st_transform(st_crs(., msa_shp))
+real_estate_metrics <- read_csv(here("data/intermediate/n_real_estate_metrics.csv")) %>% 
+  select(cbsa, anchor_city, msa, starts_with("n_")) %>% 
[…]

### 2023-06-07 R/calculate_wt_scores.R
create function to weight criteria

@@ -1,0 +1,33 @@
+#' Compute Weighted Metrics or Criteria Score
+#'
+#' THis function normalizes values using the min and max of a given vector 
+#' @param   data df of columns to be weighted
+#' @param   weights vector of weights, should sum to one, to weight each column. Number of weights should match number of columns
+#'
+#' @return  Normalized column
+
+calculate_wt_scores <- function(data, weights){
+  
+    # multiply each criteria by respective weight
+    mutate(wt_real_estate = real_estate_score * weights[1],
+           wt_landlord_policy = landlord_score * weights[2], 
+           wt_solar_irr = financial_score * weights[3],
+           wt_co2 = co2_score * weights[4], 
+           wt_health_impacts = health_impact_score * weights[5],
+           wt_electricity = electricity_score * weights[6],
+           wt_climate_risk = climate_risk_score * weights[7]) %>% 
+    # select weighted columns and rename them
+    select(!c(ends_with("_score"))) %>%
+    # create column for investment score
+    mutate(investment_score = 0) %>%
+    # for loop to sum weighted columns for each row
+    mutate(investment_score = apply(.[, 4:10], 1, sum)) %>% 
+    rename("Real Estate"="wt_real_estate",
+           "Landlord Policy"="wt_landlord_policy",
+           "Electricity Generation"="wt_electricity", 
+           "Avoided CO2 Emissions"="wt_co2",
+           "Climate Risk Avoided"="wt_climate_risk",
+           "Avoided Health Costs"="wt_health_impacts", 
+           "Solar Financials"="wt_solar_irr") %>%  
+    unite("msa_state", city_msa:state, sep = ", ")
+}

### 2023-06-07 src/01_metric_normalization.Rmd
read csv instead of xlsx file

@@
 
-anchor_city <- read.xlsx(here("data/intermediate/demographic_data.xlsx"), sheetName="anchor_city") %>% 
+anchor_city <- read_csv(here("data/intermediate/demographic_data.csv")) %>% 
   clean_names() %>% 

### 2023-06-08 solar_fam_shiny/app.R
added input boxes for landlord metrics

@@ -2,42 +2,51 @@ library(shiny)
 library(ggplot2)
 library(tidyverse)
 library(bslib) # shiny theme
-library(tidycensus)
-library(googlesheets4)
 library(sf) # map states and US
-library(readxl)
 library(here)
 library(tmap)
 library(tigris)
 library(tmaptools)
+source(here("R/normalize.R"))
+library(rsconnect) # to publish 
+library(shinydashboard)
+library(shinyWidgets)
 
+
+
+#### READ IN SHP DATA ----------------------------------------------
 cbsa_geom <- core_based_statistical_areas(cb = TRUE) %>% 
   mutate(cbsa = factor(CBSAFP))
-
 
 states_geom <- tidycensus::state_laea    #read in the dataset
 
 
-# state and U.S. map coordinates
-states <- map_data("state") %>% 
-  st_as_sf(coords = c("long", "lat"))
-
-usa<- map_data("usa") %>% 
-  st_as_sf(coords = c("long", "lat"))
-
+#### READ IN Metric and criteria DATA ----------------------------------------------
 real_estate_metrics <- read_csv(here("data/intermediate/n_real_estate_metrics.csv")) %>% 
   select(cbsa, anchor_city, msa, starts_with("n_")) %>% 
   mutate(cbsa = factor(cbsa)) %>% 
   rename("n_employ_change" = n_avg_emp_change)
+
+landlord_data <- read_csv(here("data/intermediate/landlord_data.csv")) %>% 
+  mutate(
+    security_deposit_limit_num = case_when(
+      security_deposit_limit == "N" ~ 1.0, 
+      security_deposit_limit == "1_month_rent" ~ 0.0,
+      security_deposit_limit == "2_months_rent" ~ 0.5)) %>% 
+  mutate(n_eviction_notice = 1- normalize(eviction_notice_days),
+         n_security_deposit = normalize(security_deposit_limit_num)) %>% 
+  select(cbsa, rent_control, n_eviction_notice, n_security_deposit)
 
 criteria_unweighted <- read_csv(here("data/intermediate/unwt_criteria_scores.csv")) %>% 
   mutate(cbsa = factor(cbsa),
          real_estate_score = 0) 
 
 
+#### start of shiny app----------------------------------------------
 ui <- fluidPage(
+  
   # Application title
-  titlePanel("title"),
+  titlePanel("Identifying favorable Areas for investing in Rooftop Solar on Multifamily Housing"),
   navbarPage("solar fam",
[…]

### 2023-06-08 solar_fam_shiny/app.R
adding tab icons and instructions for using shiny app

@@ -11,12 +11,13 @@ source(here("R/normalize.R"))
 library(rsconnect) # to publish 
 library(shinydashboard)
 library(shinyWidgets)
+library(tidycensus)
 
 
 
 #### READ IN SHP DATA ----------------------------------------------
 cbsa_geom <- core_based_statistical_areas(cb = TRUE) %>% 
-  mutate(cbsa = factor(CBSAFP))
+  mutate(cbsa = as.character(CBSAFP))
 
 states_geom <- tidycensus::state_laea    #read in the dataset
 
@@ -24,7 +25,7 @@ states_geom <- tidycensus::state_laea    #read in the dataset
 #### READ IN Metric and criteria DATA ----------------------------------------------
 real_estate_metrics <- read_csv(here("data/intermediate/n_real_estate_metrics.csv")) %>% 
   select(cbsa, anchor_city, msa, starts_with("n_")) %>% 
-  mutate(cbsa = factor(cbsa)) %>% 
+  mutate(cbsa = as.character(cbsa)) %>% 
   rename("n_employ_change" = n_avg_emp_change)
 
 landlord_data <- read_csv(here("data/intermediate/landlord_data.csv")) %>% 
@@ -34,83 +35,126 @@ landlord_data <- read_csv(here("data/intermediate/landlord_data.csv")) %>%
       security_deposit_limit == "1_month_rent" ~ 0.0,
       security_deposit_limit == "2_months_rent" ~ 0.5)) %>% 
   mutate(n_eviction_notice = 1- normalize(eviction_notice_days),
-         n_security_deposit = normalize(security_deposit_limit_num)) %>% 
+         n_security_deposit = normalize(security_deposit_limit_num),
+         cbsa = as.character(cbsa)) %>% 
   select(cbsa, rent_control, n_eviction_notice, n_security_deposit)
 
 criteria_unweighted <- read_csv(here("data/intermediate/unwt_criteria_scores.csv")) %>% 
-  mutate(cbsa = factor(cbsa),
-         real_estate_score = 0) 
+  mutate(cbsa = as.character(cbsa)) 
+
+####### Informatin
+RE_input_text <- ""
+abstract <- "As the renewable energy transition accelerates, housing, due to its high energy demand, 
+can play a critical role in the clean energy shift. Specifically, multifamily housing provides a unique 
+opportunity for solar photovoltaic (PV) system adoption, given the existing competing interests between 
+landlords and tenants which has historically slowed this transition. Landlords are less incentivized to
[…]

### 2023-06-09 solar_fam_shiny/app.R
fixing barchart output

@@ -12,6 +12,7 @@ library(rsconnect) # to publish
 library(shinydashboard)
 library(shinyWidgets)
 library(tidycensus)
+library(shiny.router)
 
 
 
@@ -35,9 +36,9 @@ landlord_data <- read_csv(here("data/intermediate/landlord_data.csv")) %>%
       security_deposit_limit == "1_month_rent" ~ 0.0,
       security_deposit_limit == "2_months_rent" ~ 0.5)) %>% 
   mutate(n_eviction_notice = 1- normalize(eviction_notice_days),
-         n_security_deposit = normalize(security_deposit_limit_num),
+         n_security_deposit = security_deposit_limit_num,
          cbsa = as.character(cbsa)) %>% 
-  select(cbsa, rent_control, n_eviction_notice, n_security_deposit)
+  select(cbsa, rent_control, n_eviction_notice, n_security_deposit) 
 
 criteria_unweighted <- read_csv(here("data/intermediate/unwt_criteria_scores.csv")) %>% 
   mutate(cbsa = as.character(cbsa)) 
@@ -66,36 +67,41 @@ Future studies can utilize the reproducible code to inform decisions on where to
 ui <- fluidPage(
   
   # Application title
-  titlePanel(div("Identifying areas to invest in Rooftop Solar on Multifamily Housing", 
-                 tags$img(src = "bren-leaf-logo.jpg", align="right", height="40px", width="40px")),
+  titlePanel(div(h4("Identifying areas to invest in Rooftop Solar on Multifamily Housing", 
+                 tags$img(src = "bren-leaf-logo.jpg", align="right", height="30px", width="30px"))),
              windowTitle="solar family"),
   navbarPage("solar fam",
              theme = bs_theme(bootswatch = "cerulean", primary = "#8aab57"),
              ### tab 1 - background
-             tabPanel("Landing Page", icon=icon("home", lib="glyphicon"), 
+             tabPanel("Landing Page", icon=icon("home", lib="glyphicon"), id="home", 
                       mainPanel(
-                        h5("How to use the shiny app"),
+                        h5("How to navigate the shiny app"),
                         p("This web app was created to show how several different criteria can impact the
                           investment feasability of rooftop solar on multifamily housing across the country.
                           The model includes 7 criteria that factor into the investment favorability score, in which
                           stakeholders can adjust the weight of each based on the relative importance the criteria given their priorities when
                           investing in multifamily housing."),
                         br(),
-                        p("To start, click on the Use the Model tab, and it will prompt you through each step. After each step,
[…]
